### 2022-09-15 Image_Analysis_Test_Anabena.Rmd
Initial uploads to git repository

---
title: "Image-Analysis-Test_Anabena"
author: "Tyler E. Harman"
date: "8/6/2021"
output: word_document
---

```{r}
library(EBImage)
library(tiff)
library(pixmap)
```

The following code is a collection of for loops that is based off of the Pokrzywinski et al., 2019 technical report (https://erdc-library.erdc.dren.mil/jspui/handle/11681/33870) utilizing R to automate cell counts of fluorescent microscopy images. This code uses [EBImage](https://www.bioconductor.org/packages/release/bioc/html/EBImage.html), [tiff](https://cran.r-project.org/web/packages/tiff/index.html), and [pixmap](https://cran.r-project.org/web/packages/pixmap/index.html) packages.

## Cell Counting Script (fluorescent channel 1)

The following for loops are taken from the 2019 technical report, but split into two different collections for two different fluorescent channels. Ultimately, this will be beneficial for researchers who want to detect differences in live/dead cells or different counts of fluorescent markers.

First, we need to convert the microscopy images into 8-bit greyscale images in order for the code to correctly analyze these images for the counting process:
```{r}
for(i in 1:1){
  F1_path<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test F1/")
  grey_F1_savdir<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F1_Grey/")
  F1_files<-list.files(F1_path,pattern="tif",full.name=F)
}
```

This takes the images from a pre-determined path and saves these converted files into a new save directory

The following code is the for-loop that converts the images to greyscale for all images within the pre-determined path

```{r}
for(i in 1:length(F1_files)){
  grey_image <- paste0(sub(".tif",replacement="",x=F1_files[i]),"_grey.tiff")
  image <- readTIFF(paste0(F1_path,F1_files[i]))
  image <- 2*image
  bar <- image[,,1]+image[,,2]+image[,,3]
  bar <- bar/max(bar)
  ImgF1=normalize(bar,inputRange=c(0.2,0.75))
  writeTIFF(ImgF1,paste0(grey_F1_savdir,"/",grey_image))
}
display(ImgF1)
```

This code corrects the contrast from the original image using the "image <- 2*image" code, while the "bar" variables convert the image to greyscale, and finnaly the 'normalize' function is used to reduce background noise below the 0.2 threshold.

Next, the images go through the following code to process the newly converted greyscale images to be mapped for further processing:
```{r}
for(i in 1:1){
  map_pathF1<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F1_Grey/")
  mapF1_savdir<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F1_Map/")
  map_F1files<-list.files(map_pathF1,pattern="tif",full.name=F)
}
for(i in 1:length(map_F1files)){
  mapped_image <- paste0(sub(".tiff", replacement="", x=map_F1files[i]),"_mapped.tiff")
  map1 <- readTIFF(paste0(map_pathF1,map_F1files[i]))
  map2 <- as.matrix(map1)
  map2[map2<0.3] <- 0
  writeTIFF(map2,paste0(mapF1_savdir,"/",mapped_image))
}
```

Once these images are processed, the last step is to generate the needed files paths for pulling/saving images, as well as creating a file path for saving a generated .csv file for the resulting data, as well as generating a data frame for the data to be loaded into:
```{r}
for(i in 1:1){
  image_pathF1<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F1_Map/")
  image_F1savdir<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F1_Analysis/")
  image_F1files<-list.files(image_pathF1,pattern="tiff",full.name=F)
  csv_F1path<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F1_Matrices/")
  resultsF1 <- data.frame(image=rep(NA, length(image_F1files)), count = rep(NA,length(image_F1files)))
}
```

Once these paths are made, the following code goes through each image within the pre-determined file path with mapped images and processes them for counting:
```{r}
for(i in 1:length(image_F1files)){
  analyzed_image <- paste0(sub(".tiff", replacement = "", x = image_F1files[i]),"_analyzed.tiff")
  image_matrix <- paste0(sub(".tiff", replacement = "", x = image_F1files[i]), "_matrix.csv")
  img1 <- readTIFF(paste0(image_pathF1,image_F1files[i]))
  img2 <- thresh(img1,w=17,h=17,offset=0.001)
  display(img2,method="raster")
  nmaskt <- watershed(distmap(img2),0.5)
  nmaskt <- fillHull(nmaskt)
  img3 <- colorLabels(nmaskt,normalize=T)
  display(img3, method="raster",all=F)
  count <- max(nmaskt)
  dims=dim(nmaskt)
  border=c(nmaskt[1:dims[1],1],nmaskt[1:dims[1],dims[2]], nmaskt[1,1:dims[2]],nmaskt[dims[1],1:dims[2]])
  ids=unique(border[which(border !=0)])
  inner=rmObjects(nmaskt,ids)
  count2 <- max(inner)
  img4 <- colorLabels(inner,normalize=T)
  display(img4, method="raster")
  writeImage(img4, files=paste0(image_F1savdir,"/",analyzed_image))
  write.csv(inner,file=paste0(csv_F1path,"/",image_matrix))
  resultsF1$image[i] <- analyzed_image
  resultsF1$count[i] <- count
  resultsF1$edge[i] <- length(ids)
  resultsF1$inner[i] <- count2
}
display(img4)
```

This code highlights individual cells with a unique color for the loop to eventually count as accurately as possible, as you can see with the image from above. The last amount of code is used to write the data from "resultsF1" to a .csv file:
```{r}
write.csv(resultsF1,paste0(image_F1savdir,"/F1_counts.csv"))
```

## Fluorescent channel 2 counts

The following code is similar to the fluorescent channel 1 code, but optimized for images within separate file paths:
```{r}
for(i in 1:1){
  #Path/directory/list of greyscale imaging
  F2_path<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test F2/")
  grey_F2_savdir<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F2_Grey/")
  F2_files<-list.files(F2_path,pattern="tif",full.name=F)
}

#create loop for batch converting F2 images to greyscale
for(f in 1:length(F2_files)){
  grey_image <- paste0(sub(".tif",replacement="",x=F2_files[f]),"_grey.tiff")
  image <- readTIFF(paste0(F2_path,F2_files[f]))
  image <- 2*image
  bar<- image[,,1]+image[,,2]+image[,,3]
  bar <- bar/max(bar)
  ImgF2=normalize(bar,inputRange=c(0.2,0.75))
  writeTIFF(ImgF2,paste0(grey_F2_savdir,"/",grey_image))
}

for(i in 1:1){
  #Path/directory/list of pixel maps
  map_pathF2<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F2_Grey/")
  mapF2_savdir<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F2_Map/")
  map_F2files<-list.files(map_pathF2,pattern="tif",full.name=F)
}

#create loop for batch image processing/creating mapped images
for(i in 1:length(map_F2files)){
  mapped_image <- paste0(sub(".tiff", replacement="", x=map_F2files[i]),"_mapped.tiff")
  map1 <- readTIFF(paste0(map_pathF2,map_F2files[i]))
  map2 <- as.matrix(map1)
  map2[map2<0.3] <- 0
  writeTIFF(map2,paste0(mapF2_savdir,"/",mapped_image))
}

for(i in 1:1){
  #Path/directory/list for images for processing
  image_pathF2<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F2_Map/")
  image_F2savdir<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F2_Analysis/")
  image_F2files<-list.files(image_pathF2,pattern="tiff",full.name=F)
  #Path to save csv files from image matrix - visual reference
  csv_F2path<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F2_Matrices/")
  #create loop to process and count the new images
  resultsF2 <- data.frame(image=rep(NA, length(image_F2files)), count = rep(NA,length(image_F2files)))
}

#create loop to process and count the new images
for(i in 1:length(image_F2files)){
  analyzed_image <- paste0(sub(".tiff", replacement = "", x = image_F2files[i]),"_analyzed.tiff")
  image_matrix <- paste0(sub(".tiff", replacement = "", x = image_F2files[i]), "_matrix.csv")
  img1 <- readTIFF(paste0(image_pathF2,image_F2files[i]))
  img2 <- thresh(img1,w=17,h=17,offset=0.001)
  display(img2,method="raster")
  nmaskt <- watershed(distmap(img2),0.5)
  nmaskt <- fillHull(nmaskt)
  img3 <- colorLabels(nmaskt,normalize=T)
  display(img3, method="raster",all=F)
  count <- max(nmaskt)
  dims=dim(nmaskt)
  border=c(nmaskt[1:dims[1],1],nmaskt[1:dims[1],dims[2]], nmaskt[1,1:dims[2]],nmaskt[dims[1],1:dims[2]])
  ids=unique(border[which(border !=0)])
  inner=rmObjects(nmaskt,ids)
  count2 <- max(inner)
  img4 <- colorLabels(inner,normalize=T)
  display(img4, method="raster")
  writeImage(img4, files=paste0(image_F2savdir,"/",analyzed_image))
  write.csv(inner,file=paste0(csv_F1path,"/",image_matrix))
  resultsF2$image[i] <- analyzed_image
  resultsF2$count[i] <- count
  resultsF2$edge[i] <- length(ids)
  resultsF2$inner[i] <- count2
}

write.csv(resultsF2,paste0(image_F2savdir,"/F2_counts.csv"))
```

## Multi-layered imaging analysis

Once the following counting procedures have been completed, the following code will take the grey-scale images produced previously to create outlines of the cells using the 'bwlabel' function. This creates black and white images of the cells for the 'propagate' function to accurately generate cell outlines.

The following code loads the needed file paths:
```{r}
for(i in 1:1){
  grey_pathF1<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F1_Grey/")
  grey_pathF2<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F2_Grey/")
  multi_savdir<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_Multi/")
  grey_filesF1<-list.files(grey_pathF1,pattern="tif",full.name=F)
  grey_filesF2<-list.files(grey_pathF2,pattern="tif",full.name=F)
}
```

Next, the following nested-loop processes these grey-scale images one at a time, then ultimately deletes the greyscale images after the multi-layered imaging is completed.

First, the file paths are reloaded within the nested-loop. As the nested-loop continues, these file paths need to be reloaded in order for the loop to begin the next image that's in line. In addition, another for-loop in placed after the reloading of the file paths in order to load in the file name of the first available image and associate it to a variable, in this case, being 'analyzed image'. This writes '_multi.tiff' to the file name:
```{r}
for(i in 1:length(grey_filesF1)){
  grey_pathF1<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F1_Grey/")
  grey_pathF2<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_F2_Grey/")
  multi_savdir<-("D:/NOAA Work/Cell Count Project/Multi-Detection/Test 2/Test_Multi/")
  grey_filesF1<-list.files(grey_pathF1,pattern="tif",full.name=F)
  grey_filesF2<-list.files(grey_pathF2,pattern="tif",full.name=F)
  for(i in 1:length(grey_filesF1)){
    analyzed_image <- paste0(sub(".tiff", replacement = "", x = grey_filesF1[i]),"_multi.tiff")
    break
  }
}
```

The remaining code within the nested-loop creates these 'bwlabel' images for the first image in line for each file path for F1 and F2 images. Notice the 'break' within the for-loops; this stops the for-loop to only process one image at a time:
```{r}
  for(i in 1:length(grey_filesF1)){
    image_greyF1 <- readImage(paste0(grey_pathF1,grey_filesF1[i]))
    ImgF1=normalize(image_greyF1,inputRange=c(0,0.75))
    F1_lab<-thresh(ImgF1,w=15,h=15,offset=0.05)
    F1_lab<-opening(F1_lab,makeBrush(5,shape='disc'))
    F1_lab<-fillHull(F1_lab)
    F1_lab<-bwlabel(F1_lab)
    break
  }
  for(f in 1:length(grey_filesF2)){
    image_greyF2 <- readImage(paste0(grey_pathF2,grey_filesF2[f]))
    ImgF2=normalize(image_greyF2,inputRange=c(0,0.75))
    F2_lab<-thresh(ImgF2,w=15,h=15,offset=0.05)
    F2_lab<-opening(F2_lab,makeBrush(5,shape='disc'))
    F2_lab<-fillHull(F2_lab)
    F2_lab<-bwlabel(F2_lab)
    break
  }
display(F1_lab)
display(F2_lab)
```

Once these images are made, the following code creates a colored layered image, as well as taking the 'bwlabel' images (i.e., F1_lab and F2_lab) and creaties outlines using the 'propagate' function, and saves the file to a pre-determined save directory:
```{r}
  img=rgbImage(red=ImgF2,blue=1.5*ImgF1)
  F2_mask<-opening(F2_lab)
  F2_mask2<-propagate(F2_lab,seeds=F2_lab,mask=F2_mask)
  res=paintObjects(F2_mask2,img,col='#ff00ff')
  res=paintObjects(F1_lab,res,col='#f0fff0')
  writeImage(res, files=paste0(multi_savdir,"/",analyzed_image))
  display(res)
```

Once the image has been saved, we now need to delete the grey-scale images used in this one analysis, in order for the nested-loop to process the next image in line (i.e., number 2, number 3, etc.). The following code deletes the grey-scale images from each fluorescent channel file path; notice the 'break' function, as we only want to delete the first file in the respective file path:
```{r}
  for(i in 1:length(grey_filesF1)){
    if(file.exists(paste0(grey_pathF1,grey_filesF1[i])))
      file.remove(paste0(grey_pathF1,grey_filesF1[i]))
    break
  }
  for(f in 1:length(grey_filesF2)){
    if(file.exists(paste0(grey_pathF2,grey_filesF2[f])))
      file.remove(paste0(grey_pathF2,grey_filesF2[f]))
    break
  }
```

This nested loop will run automatically, doing the following code for the length of the files that are processed.

### 2022-09-15 cellcount_intro.Rmd
Initial uploads to git repository

---
title: "Introduction to *cellcount*"
author: "*T.E. Harman, B. Boyd, K. Pokrzywinski*"
date: "11/2/2021"
abstract: >
  The *cellcount* package provides wrapper functions for  [*EBImage*](http://bioconductor.org/packages/devel/bioc/vignettes/EBImage/inst/doc/EBImage-introduction.html) to simplfy quantifying bacteria and cyanobacteria via flourescence microscopy. Cyanobacteria enumeration is an important tool to incorporate data results into mathematical models to assess bloom dynamics. However, standard manual cyanobacteria enumeration is considered time-consuming and can impact overall random error due to different inaccuracies from researcher to researcher. Here, we outline a new open-source tool, *cellcount*, a package designed for the computing language R to assist with cyanobacteria enumeration. Fluorescent microscopy images are analyzed by this package to generate cyanobacteria counts to assess overall cell density from samples.
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting started:

*cellcount* is an R package distributed as part of the [Bioconductor](http://bioconductor.org) project. To install the package, start R and enter:

```{r, message=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
if (!require("cellcount")) {
  BiocManager::install("cellcount")
}
library(cellcount)
```

The following packages are required. Please install and bring them into the R
environment: 
```{r, message=FALSE}
if (!require("EBImage")) {
  BiocManager::install("EBImage")
}
library(EBImage)
library(tiff)
library(pixmap)
library(raster)
```

# Setting up directories/reading files:

Next, generate a data frame for count data to be added into using the following
code:
```{r}
Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
```

You'll need to generate a couple of directories; one for saving the unique
color label images and another for CSV files. In this example, we are using 
images of vegetative *Alexandrium catenella* cells. Please note that your 
directory addresses will be different. We recommend using a dedicated directory 
for *cellcount* results:

```{r}
savdir <- ("E:/cellcount_data/CSV_data/")
image_savdir <- ("E:/cellcount_data/Convert_Images/Convert_Alexandrium10X/")
```

Next, you'll need to read in the images you are interested in processing using
the following code. Once again, your directories will be different and must be
updated to the correct file address:

```{r}
images <- list.files("E:/cellcount_data/Alexandrium_10X/", pattern = "tif", full.name = T)
images_names <- list.files("E:/cellcount_data/Alexandrium_10X/", pattern = "tif", full.name = F)
```

In addition, you'll need to assign the file names to a vector along with adding
in characterization to each file for easy organization:

```{r}
imgNames <- paste0("Alexandrium_", images_names)
```

Now we will "read" in the images that we are interested in analyzing. The
following code outlines this step using the "readTIFF" function from the 
*tiff* package:

```{r, warning=FALSE}
read_images <- lapply(images, readTIFF)
names(images) <- imgNames
```

# Beginning the conversion process:

Now that our images are within the R environment, we can begin to convert these
images into useful data that R can use to process images. First, we begin by
converting 24-bit images into 8-bit images using the "greyscale" function (see 
photo below). This function also has a built-in numerical contrast adjustment 
that end-users can adjust. Note that if you already have 8-bit images, you can 
skip this step:

```{r}
grey_images <- lapply(read_images, greyscale, contrast = 2)
display(grey_images[[1]])
```

Next, we will use the "mapped" function to turn the 8-bit image we just
converted into a data matrix using the following code. This also includes a
pixel intensity threshold adjustment that end-users can change; this removes
any pixel that is under the threshold value to easily separate the background
from cells:

```{r}
imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
```

Before we analyze our images in bulk, set aside some code space to visually 
analyze your images results. This could include adjusting the threshold
from the "mapped" function, the contrast from the "greyscale" function 
(if used), or could involve making adjustments to the "image_convert" function,
such as the width and height of the rectangular window or the pixel area
threshold (i.e., removes cells or other objects under a certain pixel area).
Observe your initial results and make adjustments where necessary:

```{r}
imagesConverted <- image_convert(imagesMapped[[1]], w = 15, h = 25, offset = 0.01, areathresh = 150)
final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
display(final_img)
```

# Analysis of images for producing count data:

Once you have made necessary adjustments, use the following for-loop to
analyze your images in bulk, one at a time. The for-loop uses the 
"image_convert" function to produce unique color labels on cells and produce an
image showcasing these labels using the "countImages" function. Finally, 
"countCells" is used to count the number of unique color labels (i.e., cells).
We also have integrated code to add these count results to the data frame made
earlier, and also to save each unique color label image to the "image_savdir" 
directory:

```{r}
for (z in 1:length(images)) {
  imagesConverted <- image_convert(imagesMapped[[z]], w = 15, h = 25, offset = 0.01, areathresh = 150)
  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
  count <- countCells(imagesConverted)
  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
  # display single-cell images via image_covert
  analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
  writeImage(final_img, files = paste0(image_savdir, analyzed_image))
}
```

Data from the image analysis has been generated in our data frame "Cell.Count"
that we previously made. We can use the following code to view our results:

```{r}
Cell.Count
```

In order to calculate cell density of these images, we can use the
"cell.density" function that's built into the *cellcount* package. We first
need to generate the total amount of cells counted from these images and
add it to our existing data frame using the following code:

```{r}
cell.total <- as.numeric(Cell.Count$Cell_Count)
cell.total <- sum(cell.total)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total)
```

Calculating cell density, taken from Wetzel and Likens 1991, requires multiple
values. End-users will need the following information to precisely 
calculate this value:

- Total counts (generated previously)
- Height of image field of view (μm)
- Width of image field of view (μm)
- Median cell diameter (μm)
- Total number of images analyzed
- Filtration area (mm2)
- Volume filtered per image (mL)
- Total sample volume (mL)

The following code uses these variables to calculate cell density and adds 
the data into the "Cell.Count" data frame:

```{r}
cell.den <- cell_density(cell.total, height = 203.2, width = 166.4, diameter = 2.5, images = 15, filtration.area = 491, volume = 1, total.volume = 100)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den)
```

Finally, we write the following code to export our finalized .csv file to our
pre-determined save directory:

```{r}
write.csv(Cell.Count, paste0(savdir, "/Alexandrium counts.csv"))
```

Literature

### 2022-09-15 R/border.R

@@ -1,0 +1,20 @@
+#' Border conversion
+#'
+#' Takes images to predict cells along a predetermined border in order
+#' to count cells along the predetermined border. This will be used to help
+#' determine those cells to "remove" them from the count function "countCells".
+#'
+#' @param x An image file
+#' @param y Dimming variable (using the 'dim' function from EBImage)
+#' @return Analyzing the cells on the border of \code{x}
+#' @examples
+#' img0 <- list.files("data/Anabena 2021/UV", pattern = "tif", full.name = TRUE)
+#' img1 <- lapply(img0, readTIFF)
+#' img2 <- lapply(img1, greyscale)
+#' img3 <- lapply(img2, mapped)
+#' img4 <- image_convert(img3)
+#' dims <- dim(img4)
+#' border(img4, dims)
+border <- function(x, y) {
+  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
+}

### 2022-09-15 R/cell_density.R

@@ -1,0 +1,30 @@
+#' Cell density
+#'
+#' Takes count results from processed images to calculate cell density based
+#' on camera field of view settings (inputted by end-user) and amount of cells
+#' filtered that the image is visualizing.
+#'
+#' @param x count results from processed image
+#' @param height height of image field of view in μm (comes from camera settings)
+#' @param width width of image field of view in μm (comes from camera settings)
+#' @param diameter median cell diameter in μm
+#' @param images total number of fields
+#' @param filtration.area filtration area in mm2
+#' @param volume volume filtered per field in mL
+#' @param total.volume total sample volume in mL
+#'
+#' @examples
+#' img <- list.files("data/Shewanella", pattern = "tif", full.name = T)
+#' img1 <- lapply(img[[1]], readTIFF)
+#' img2 <- lapply(img1, greyscale, contrast = 2)
+#' img3 <- lapply(img2, mapped, threshold = 0.3)
+#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.01, areathresh = 150)
+#' cell.total <- countCells(img4)
+#' cell_density(cell.total, height = 203.2, width = 166.4, diameter = 2.5, images = 15, filtration.area = 491, volume = 1, total.volume = 100)
+cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
+  MCF <- filtration.area / FOV
+  N <- x / images
+  cyano <- (MCF * N) / volume
+  cyano_total <- cyano * total.volume
+  return(cyano_total)
+}

### 2022-09-15 R/countCells.R

@@ -1,0 +1,24 @@
+#' Image count
+#'
+#' Takes information from a converted image and counts cells within the image.
+#' This uses the "border" function to remove cells along the image border and
+#' only count cells within the image. This creates a more accurate cell count.
+#'
+#' @param x converted image variable
+#' @return Counting cells within \code{x}
+#'
+#' @examples
+#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
+#' img1 <- lapply(img[[1]], readTIFF)
+#' img2 <- lapply(img1, greyscale, contrast = 2)
+#' img3 <- lapply(img2, mapped, threshold = 0.3)
+#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.01, areathresh = 150)
+#' countCells(img4)
+countCells <- function(x) {
+  countAll <- max(x)
+  dims <- dim(x)
+  border1 <- border(x, dims)
+  ids <- unique(border1[which(border1 != 0)])
+  countInner <- countAll - length(ids)
+  return(countInner)
+}

### 2022-09-15 R/countImages.R

@@ -1,0 +1,30 @@
+#' Count Images
+#'
+#' Displays converted microscopy images of distinct cell counts subsequently from the image_convert function.
+#' Can use the "display" function from EBImage to automatically generate the
+#' produced image from this function.
+#'
+#' @param image A tif image
+#' @param normalize if TRUE normalizes the resulting color image
+#' @param removeEdgeCells automatically removes cells from along the image
+#' border
+#'
+#' @examples
+#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
+#' img1 <- lapply(img[[1]], readTIFF)
+#' img2 <- lapply(img1, greyscale, contrast = 2)
+#' img3 <- lapply(img2, mapped, threshold = 0.3)
+#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.01, areathresh = 150)
+#' final_img <- countImages(img4, normalize = T, removeEdgeCells = T)
+#' display(final_img)
+countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+  if (removeEdgeCells) {
+    dims <- dim(image)
+    border1 <- border(image, dims)
+    ids <- unique(border1[which(border1 != 0)])
+    inner <- rmObjects(image, ids)
+    EBImage::colorLabels(inner, normalize = normalize)
+  } else {
+    EBImage::colorLabels(image, normalize = normalize)
+  }
+}

### 2022-09-15 R/greyscale.R

@@ -1,0 +1,20 @@
+#' Grey-scale conversion
+#'
+#' Takes images to adjust the contrast and converts them to grey-scale/8-bit.
+#' This function is needed to use the "mapping" function to properly
+#' convert images to data-frames.
+#'
+#' @param x an image variable
+#' @param contrast a numerical value for contrast adjustment
+#'
+#' @examples
+#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
+#' img1 <- lapply(img[[1]], readTIFF)
+#' img2 <- lapply(img1, greyscale, contrast = 2)
+greyscale <- function(x, contrast = 2) {
+  x <- contrast * x
+  x <- x[, , 1] + x[, , 2] + x[, , 3]
+  x <- x / max(x)
+  x <- normalize(x, inputRange = c(0.1, 0.75))
+  return(x)
+}

### 2022-09-15 R/image_convert.R

@@ -1,0 +1,27 @@
+#' Image convert
+#'
+#' Takes a mapped image (i.e., data matrix) and counts individual cells within an image.
+#' This is to be used with images with a single image that contains a single
+#' fluorescent dimension of cells. Used for single-cell type images.
+#'
+#' @param x an image variable that has been mapped using the "mapped" function
+#' @param w width of the moving rectangular window (see EBImage)
+#' @param h heigth of the moving rectangular window (see EBImage)
+#' @param offset a thresholding offset
+#' @param areathresh a numerical value of individual pixel areas of cells - this sets a threshold for what items are removed
+#'
+#' @examples
+#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
+#' img1 <- lapply(img[[1]], readTIFF)
+#' img2 <- lapply(img1, greyscale, contrast = 2)
+#' img3 <- lapply(img2, mapped, threshold = 0.3)
+#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.001, areathresh = 150)
+image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- fillHull(image)
+  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}

### 2022-09-15 R/image_convert2.R

@@ -1,0 +1,29 @@
+#' Image convert_2
+#'
+#' Takes a mapped image (i.e., data matrix) and counts individual cells within an image.
+#' This is to be used with images with a single image that contains a single
+#' fluorescent dimension of cells. Used for filamentous/colonial-type cell
+#' images.
+#'
+#' @param x an image variable that has been mapped using the "mapped" function
+#' @param w width of the moving rectangular window (see EBImage)
+#' @param h heigth of the moving rectangular window (see EBImage)
+#' @param offset a thresholding offset
+#' @param areathresh a numerical value of individual pixel areas of cells - this sets a threshold for what items are removed
+#'
+#' @examples
+#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
+#' img1 <- lapply(img[[1]], readTIFF)
+#' img2 <- lapply(img1, greyscale, contrast = 2)
+#' img3 <- lapply(img2, mapped, threshold = 0.3)
+#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.001, areathresh = 150)
+
+image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
+  image2 <- fillHull(image1)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}

### 2022-09-15 R/mapped.R

@@ -1,0 +1,19 @@
+#' Matrix mapping conversion
+#'
+#' Takes a grey-scale/8-bit image to convert to a data matrix,
+#' where data under the numerical pixel intensity threshold is removed.
+#' This function is needed for the use of the "image_convert" function.
+#'
+#' @param x a matrix mapping conversion function
+#' @param threshold a numerical value of pixel intensity
+#'
+#' @examples
+#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
+#' img1 <- lapply(img[[1]], readTIFF)
+#' img2 <- lapply(img1, greyscale, contrast = 2)
+#' img3 <- lapply(img2, mapped, threshold = 0.3)
+mapped <- function(x, threshold = 0.3) {
+  x <- as.matrix(x)
+  x[x < threshold] <- 0
+  return(x)
+}

### 2022-09-15 R/outline.R

@@ -1,0 +1,16 @@
+#' Outline conversion
+#'
+#' Takes grey-scale images to convert to an outline format for the 'propagation'
+#' function. Needed for the "multi_image" function.
+#' @param x normalized image variable (see example)
+#' @examples
+#' image_greyF2 <- readImage(paste0(grey_pathF2, grey_filesF2[f]))
+#' ImgF2 <- normalize(image_greyF2, inputRange = c(0, 0.75))
+#' F2_lab <- outline(ImgF2)
+outline <- function(x) {
+  x <- thresh(x, w = 15, h = 15, offset = 0.05)
+  x <- opening(x, makeBrush(5, shape = "disc"))
+  x <- fillHull(x)
+  x <- bwlabel(x)
+  return(x)
+}

### 2022-09-15 sandbox_analysis_test.R

@@ -1,0 +1,71 @@
+library(EBImage)
+library(tiff)
+library(pixmap)
+library(raster)
+library(beepr)
+library(outliers)
+library(cellcount)
+
+# clear environment and free unused memory
+gc()
+
+Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
+
+file_name<-paste0("E.coli_") #Change the name of species analyzed here
+
+#Change directories here
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/CSV_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/Convert_Images/E.coli_100x_DAPI/1E/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/1E/"
+                     , pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/1E/"
+                           , pattern = "tif", full.name = F)
+
+imgNames <- paste0(file_name, images_names)
+read_images <- lapply(images, readTIFF)
+img_transposed <- lapply(read_images,aperm,c(2,1,3))
+names(images) <- imgNames
+
+grey_images <- lapply(img_transposed, greyscale, contrast = 1)
+display(grey_images[[5]]) #visualize contrast/brightness adjustment here
+
+imagesMapped <- lapply(grey_images, mapped, threshold = 0.3) #background intensity threshold adjustment
+
+
+#IMAGE TESTING - do you need to make variable adjustments?
+
+imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+#Change variables above depending on species analyzed
+final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+display(final_img)
+
+for (z in 1:length(images)) {
+  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
[…]

### 2022-09-15 stats_code/BA_Code.R
Initial uploads to git repository

@@ -1,0 +1,88 @@
+library(ggplot2)
+library(gridExtra)
+library(ggpubr)
+
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/stat_code/")
+
+BA<-read.csv("Boxplot_Analysis.csv",header=T)
+
+BA_F192<-subset(BA,Species=="F192")
+BA_F210<-subset(BA,Species=="F210")
+BA_Anabena<-subset(BA,Species=="Anabena")
+BA_Anabena$Concentration_f = factor(BA_Anabena$Concentration, levels=c('High','Medium','Low'))
+BA_Shewanella<-subset(BA,Species=="Shewanella")
+
+
+F192<-ggplot(data=BA_F192, aes(x=Concentration,y=Output,color=Method))+
+  geom_boxplot(size=1)+
+  facet_wrap(.~Concentration,nrow=1,ncol=5,scales="free")+
+  theme_classic()+
+  labs(y="Manual counts")+
+  labs(title="Microcystis F192")+
+  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))+
+  theme(plot.title = element_text(face="bold.italic",size=30,hjust=0.5),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=13,vjust=2))+
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_text(face = "bold",size = 12,angle=30),
+        axis.ticks.x = element_blank(),
+        plot.margin=unit(c(1,1,1,1),"cm"))
+F192  
+
+F210<-ggplot(data=BA_F210, aes(x=Concentration,y=Output,color=Method))+
+  geom_boxplot(size=1)+
+  facet_wrap(.~Concentration,nrow=1,ncol=5,scales="free")+
+  theme_classic()+
+  labs(y="Manual counts")+
+  labs(title="Microcystis F210")+
+  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))+
+  theme(plot.title = element_text(face="bold.italic",size=30,hjust=0.5),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=13,vjust=2))+
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_text(face = "bold",size = 12,angle=30),
+        axis.ticks.x = element_blank(),
+        plot.margin=unit(c(1,1,1,1),"cm"))
+F210
+
[…]

### 2022-09-15 stats_code/RA_Code.R

@@ -1,0 +1,218 @@
+library(ggplot2)
+library(gridExtra)
+library(ggpubr)
+
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/stat_code/")
+
+RA<-read.csv("Regression_Analysis.csv",header=T)
+
+RA_F192<-subset(RA,Species=="F192")
+RA_F210<-subset(RA,Species=="F210")
+RA_Anabena<-subset(RA,Species=="Anabena")
+RA_Shewanella<-subset(RA,Species=="Shewanella")
+
+RA_F192B<-RA_F192[-c(1:3),]
+RA_F210B<-RA_F210[-c(1:3),]
+RA_AnabenaB<-RA_Anabena[-c(1:3),]
+RA_ShewanellaB<-RA_Shewanella[-c(1:3),]
+
+
+data.lm1<-lm(Visual~Script,data=RA_F192)
+summary(data.lm1)
+
+F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
+  theme_classic()+
+  labs(x=expression(paste(italic("cellcount"), " results")))+
+  labs(y="Manual counts")+
+  labs(title="Microcystis F192")+
+  theme(plot.title = element_text(face="bold.italic",size=17,hjust=0.5),
+        axis.title.x = element_text(size=15),
+        axis.title.y = element_text(size=15,vjust=2))+
+  theme(axis.text.x = element_text(face = "bold",size = 12),
+        axis.text.y = element_text(face = "bold",size = 12,angle=30),
+        plot.margin=unit(c(1,1,1,1),"cm"))+
+  geom_abline(slope=1,intercept=0,color="red",size=2,alpha=0.4)+
+  geom_abline(slope=coef(data.lm1)[["Script"]],
+              intercept=coef(data.lm1)[["(Intercept)"]],
+              linetype=2,
+              size=2)+
+  geom_point(size=5,shape=21,colour="black",fill="green",stroke=2)
+F192
+
+data.lm2<-lm(Visual~Script,data=RA_F210)
+summary(data.lm2)
+
+F210<-ggplot(data=RA_F210, aes(x=Script,y=Visual))+
+  theme_classic()+
+  labs(x=expression(paste(italic("cellcount"), " results")))+
+  labs(y="Manual counts")+
+  labs(title="Microcystis F210")+
+  theme(plot.title = element_text(face="bold.italic",size=17,hjust=0.5),
[…]

### 2022-09-15 stats_code/RA_IND_Code.R

@@ -1,0 +1,121 @@
+library(ggplot2)
+library(gridExtra)
+library(ggpubr)
+
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/stat_code/")
+
+RA<-read.csv("Regression_Analysis_IND.csv",header=T)
+
+RA_F192<-subset(RA,Species=="F192")
+RA_F210<-subset(RA,Species=="F210")
+RA_Anabena<-subset(RA,Species=="Anabena")
+RA_Shewanella<-subset(RA,Species=="Shewanella")
+
+data.lm1<-lm(Visual~Script,data=RA_F192)
+summary(data.lm1)
+
+F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
+  theme_classic()+
+  labs(x=expression(paste(italic("cellcount"), " results")))+
+  labs(y="Manual counts")+
+  labs(title="Microcystis F192")+
+  theme(plot.title = element_text(face="bold.italic",size=17,hjust=0.5),
+        axis.title.x = element_text(size=15),
+        axis.title.y = element_text(size=15,vjust=2))+
+  theme(axis.text.x = element_text(face = "bold",size = 12),
+        axis.text.y = element_text(face = "bold",size = 12,angle=30),
+        plot.margin=unit(c(1,1,1,1),"cm"))+
+  geom_abline(slope=1,intercept=0,color="red",size=2,alpha=0.4)+
+  geom_abline(slope=coef(data.lm1)[["Script"]],
+              intercept=coef(data.lm1)[["(Intercept)"]],
+              linetype=2,
+              size=2)+
+  geom_point(size=5,shape=21,colour="black",fill="green",stroke=2)
+F192
+
+data.lm2<-lm(Visual~Script,data=RA_F210)
+summary(data.lm2)
+
+F210<-ggplot(data=RA_F210, aes(x=Script,y=Visual))+
+  theme_classic()+
+  labs(x=expression(paste(italic("cellcount"), " results")))+
+  labs(y="Manual counts")+
+  labs(title="Microcystis F210")+
+  theme(plot.title = element_text(face="bold.italic",size=17,hjust=0.5),
+        axis.title.x = element_text(size=15),
+        axis.title.y = element_text(size=15,vjust=2))+
+  theme(axis.text.x = element_text(face = "bold",size = 12),
+        axis.text.y = element_text(face = "bold",size = 12,angle=30),
+        plot.margin=unit(c(1,1,1,1),"cm"))+
[…]

### 2022-09-15 stats_code/RA_Residuals.R

@@ -1,0 +1,183 @@
+library(ggplot2)
+library(gridExtra)
+library(ggpubr)
+
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/stat_code/")
+
+RA<-read.csv("Regression_Analysis_IND.csv",header=T)
+
+RA_F192<-subset(RA,Species=="F192")
+RA_F210<-subset(RA,Species=="F210")
+RA_Anabena<-subset(RA,Species=="Anabena")
+RA_Shewanella<-subset(RA,Species=="Shewanella")
+
+data.lm1<-lm(Visual~Script,data=RA_F192)
+RA_F192$predicted<-predict(data.lm1)
+RA_F192$residuals<-residuals(data.lm1)
+
+resid_F192<-ggplot(RA_F192,aes(x=Script,y=Visual))+
+  geom_smooth(method="lm",se=FALSE,color="lightgrey")+
+  geom_segment(aes(xend=Script,yend=predicted),alpha=0.2)+
+  geom_point(aes(color=abs(residuals),size=abs(residuals)))+
+  scale_color_continuous(low="green",high="red")+
+  guides(color="none",size="none")+
+  geom_point(aes(y=predicted),shape=1)+
+  theme_classic()+
+  labs(title="Microcystis F192")+
+  theme(
+    plot.title = element_text(face="bold.italic",size=17,hjust=0.5),
+    axis.text.x = element_text(face = "bold", 
+                               size = 13),
+    axis.text.y = element_text(face = "bold", 
+                               size = 13),
+    axis.title.x = element_text(size=17,vjust=-1),
+    axis.title.y = element_text(size=17,vjust=3),
+    plot.margin=unit(c(1,1,1,1),"cm")
+  )
+resid_F192
+
+data.lm2<-lm(Visual~Script,data=RA_F210)
+RA_F210$predicted<-predict(data.lm2)
+RA_F210$residuals<-residuals(data.lm2)
+
+resid_F210<-ggplot(RA_F210,aes(x=Script,y=Visual))+
+  geom_smooth(method="lm",se=FALSE,color="lightgrey")+
+  geom_segment(aes(xend=Script,yend=predicted),alpha=0.2)+
+  geom_point(aes(color=abs(residuals),size=abs(residuals)))+
+  scale_color_continuous(low="green",high="red")+
+  guides(color="none",size="none")+
+  geom_point(aes(y=predicted),shape=1)+
[…]

### 2022-09-15 stats_code/RosnerTest.R

@@ -1,0 +1,75 @@
+library(outliers)
+library(EnvStats)
+
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/stat_code/")
+
+RA<-read.csv("Regression_Analysis_IND.csv",header=T)
+
+RA_F192<-subset(RA,Species=="F192")
+RA_F192A<-subset(RA_F192,Concentration=="A")
+RA_F192B<-subset(RA_F192,Concentration=="B")
+RA_F192C<-subset(RA_F192,Concentration=="C")
+RA_F192D<-subset(RA_F192,Concentration=="D")
+RA_F192E<-subset(RA_F192,Concentration=="E")
+
+rosnerTest(RA_F192A$Script)
+rosnerTest(RA_F192B$Script)
+rosnerTest(RA_F192C$Script)
+rosnerTest(RA_F192D$Script)
+rosnerTest(RA_F192E$Script)
+rosnerTest(RA_F192A$Visual)
+rosnerTest(RA_F192B$Visual)
+rosnerTest(RA_F192C$Visual)
+rosnerTest(RA_F192D$Visual)
+rosnerTest(RA_F192E$Visual)
+
+
+RA_F210<-subset(RA,Species=="F210")
+RA_F210A<-subset(RA_F210,Concentration=="A")
+RA_F210B<-subset(RA_F210,Concentration=="B")
+RA_F210C<-subset(RA_F210,Concentration=="C")
+RA_F210D<-subset(RA_F210,Concentration=="D")
+RA_F210E<-subset(RA_F210,Concentration=="E")
+
+rosnerTest(RA_F210A$Script)
+rosnerTest(RA_F210B$Script)
+rosnerTest(RA_F210C$Script)
+rosnerTest(RA_F210D$Script)
+rosnerTest(RA_F210E$Script)
+rosnerTest(RA_F210A$Visual)
+rosnerTest(RA_F210B$Visual)
+rosnerTest(RA_F210C$Visual)
+rosnerTest(RA_F210D$Visual)
+rosnerTest(RA_F210E$Visual)
+
+
+RA_Anabena<-subset(RA,Species=="Anabena")
+RA_AnabenaHigh<-subset(RA_Anabena,Concentration=="High")
+RA_AnabenaMed<-subset(RA_Anabena,Concentration=="Medium")
[…]

### 2022-09-15 stats_code/RA_Code.R
Updates to code

@@ -1,8 +1,9 @@
 library(ggplot2)
 library(gridExtra)
 library(ggpubr)
+library(cowplot)
 
-setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/stat_code/")
+setwd("~/cellcount_main/data")
 
 RA<-read.csv("Regression_Analysis.csv",header=T)
 
@@ -11,123 +12,133 @@ RA_F210<-subset(RA,Species=="F210")
 RA_Anabena<-subset(RA,Species=="Anabena")
 RA_Shewanella<-subset(RA,Species=="Shewanella")
 
-RA_F192B<-RA_F192[-c(1:3),]
-RA_F210B<-RA_F210[-c(1:3),]
-RA_AnabenaB<-RA_Anabena[-c(1:3),]
-RA_ShewanellaB<-RA_Shewanella[-c(1:3),]
-
+RA_F192$Script<-(RA_F192$Script/1000000)
+RA_F192$Visual<-(RA_F192$Visual/1000000)
+RA_F192$Microscope<-(RA_F192$Microscope/1000000)
 
 data.lm1<-lm(Visual~Script,data=RA_F192)
 summary(data.lm1)
 
 F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
-  theme_classic()+
-  labs(x=expression(paste(italic("cellcount"), " results")))+
-  labs(y="Manual counts")+
-  labs(title="Microcystis F192")+
-  theme(plot.title = element_text(face="bold.italic",size=17,hjust=0.5),
-        axis.title.x = element_text(size=15),
+  theme_bw()+
+  labs(x=expression(paste(italic("cellcount"), " results (",x10^{6},")")))+
+  labs(y=expression(paste("Manual counts ( ",x10^{6},")")))+
+  theme(axis.title.x = element_text(size=15),
         axis.title.y = element_text(size=15,vjust=2))+
-  theme(axis.text.x = element_text(face = "bold",size = 12),
-        axis.text.y = element_text(face = "bold",size = 12,angle=30),
-        plot.margin=unit(c(1,1,1,1),"cm"))+
-  geom_abline(slope=1,intercept=0,color="red",size=2,alpha=0.4)+
+  theme(axis.text.x = element_text(size = 15),
+        axis.text.y = element_text(size = 15),
+        plot.margin=unit(c(1,1,1,1),"cm"),
+        panel.grid.major = element_blank(),
[…]

### 2022-09-15 stats_code/RA_Code.R
Updates to code

@@ -27,9 +27,12 @@ F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
         axis.title.y = element_text(size=15,vjust=2))+
   theme(axis.text.x = element_text(size = 15),
         axis.text.y = element_text(size = 15),
+        axis.ticks = element_line(size=1.5),
+        axis.ticks.length = unit(0.25,"cm"),
         plot.margin=unit(c(1,1,1,1),"cm"),
         panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank())+
+        panel.grid.minor = element_blank(),
+        panel.border = element_rect(size=1.5))+
   scale_x_continuous(labels=function(x) format(x,scientific=FALSE))+
   scale_y_continuous(labels=function(x) format(x,scientific=FALSE))+
   xlim(0,5)+
@@ -58,9 +61,12 @@ F210<-ggplot(data=RA_F210, aes(x=Script,y=Visual))+
         axis.title.y = element_text(size=15,vjust=2))+
   theme(axis.text.x = element_text(size = 15),
         axis.text.y = element_text(size = 15),
+        axis.ticks = element_line(size=1.5),
+        axis.ticks.length = unit(0.25,"cm"),
         plot.margin=unit(c(1,1,1,1),"cm"),
         panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank())+
+        panel.grid.minor = element_blank(),
+        panel.border = element_rect(size=1.5))+
   scale_x_continuous(labels=function(x) format(x,scientific=FALSE))+
   scale_y_continuous(labels=function(x) format(x,scientific=FALSE))+
   xlim(0,5)+
@@ -89,9 +95,12 @@ Anabena<-ggplot(data=RA_Anabena, aes(x=Script,y=Visual))+
         axis.title.y = element_text(size=15,vjust=2))+
   theme(axis.text.x = element_text(size = 15),
         axis.text.y = element_text(size = 15),
+        axis.ticks = element_line(size=1.5),
+        axis.ticks.length = unit(0.25,"cm"),
         plot.margin=unit(c(1,1,1,1),"cm"),
         panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank())+
+        panel.grid.minor = element_blank(),
+        panel.border = element_rect(size=1.5))+
   scale_x_continuous(labels=function(x) format(x,scientific=FALSE))+
[…]

### 2022-09-15 sandbox_analysis_test.R
Updates

@@ -49,7 +49,16 @@ for (z in 1:length(images)) {
 }
 
 Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
-Cell.Count$outliers<-scores(Cell.Count$Cell_Count,type="z",prob=0.9)
+outlier<-function(){
+  mean<-median(Cell.Count$Cell_Count)
+  sd<-sd(Cell.Count$Cell_Count)*2
+  sd_neg<-(mean-sd)
+  sd_pos<-(mean+sd)
+  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
+    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
+  return(SD_range_detection)
+}
+Cell.Count$SD_range<-outlier()
 
 cell.total <- as.numeric(Cell.Count$Cell_Count)
 cell.total <- sum(cell.total)

### 2022-09-15 stats_code/BA_Code.R
Updates to code

@@ -1,8 +1,9 @@
 library(ggplot2)
 library(gridExtra)
 library(ggpubr)
+library(cowplot)
 
-setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/stat_code/")
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data")
 
 BA<-read.csv("Boxplot_Analysis.csv",header=T)
 
@@ -13,76 +14,100 @@ BA_Anabena$Concentration_f = factor(BA_Anabena$Concentration, levels=c('High','M
 BA_Shewanella<-subset(BA,Species=="Shewanella")
 
 
+BA_F192$Output<-(BA_F192$Output/1000000)
+BA_F192$Output<-(sqrt(BA_F192$Output))
+
 F192<-ggplot(data=BA_F192, aes(x=Concentration,y=Output,color=Method))+
-  geom_boxplot(size=1)+
-  facet_wrap(.~Concentration,nrow=1,ncol=5,scales="free")+
-  theme_classic()+
-  labs(y="Manual counts")+
-  labs(title="Microcystis F192")+
+  scale_colour_grey()+
+  stat_boxplot(geom='errorbar')+
+  geom_boxplot(size=0.5)+
+  theme_bw()+
+  labs(x="Concentration")+
+  labs(y=expression(paste("Manual counts ( ",x10^{6},") (sqrt transformed)")))+
   scale_y_continuous(labels = function(x) format(x, scientific = FALSE))+
-  theme(plot.title = element_text(face="bold.italic",size=30,hjust=0.5),
-        axis.title.x = element_blank(),
+  theme(axis.title.x = element_text(size=15),
         axis.title.y = element_text(size=13,vjust=2))+
-  theme(axis.text.x = element_blank(),
-        axis.text.y = element_text(face = "bold",size = 12,angle=30),
-        axis.ticks.x = element_blank(),
-        plot.margin=unit(c(1,1,1,1),"cm"))
+  theme(axis.text.x = element_text(size = 15),
+        axis.text.y = element_text(size = 15),
+        axis.ticks = element_line(size=1.5),
+        axis.ticks.length = unit(0.25,"cm"),
+        plot.margin=unit(c(1,1,1,1),"cm"),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.border = element_rect(size=1.5))
 F192  
 
+BA_F210$Output<-(BA_F210$Output/1000000)
[…]

### 2022-09-15 stats_code/RA_Code.R

@@ -3,7 +3,7 @@ library(gridExtra)
 library(ggpubr)
 library(cowplot)
 
-setwd("~/cellcount_main/data")
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data")
 
 RA<-read.csv("Regression_Analysis.csv",header=T)
 

### 2022-09-20 cellcount_intro.Rmd
Updates to markdown - NOT COMPLETED

@@
 title: "Introduction to *cellcount*"
-author: "*T.E. Harman, B. Boyd, K. Pokrzywinski*"
-date: "11/2/2021"
+author: "*T.E. Harman, M. Vandersea, D.E. Berthold, H.D. Laughinghouse, K. Pokrzywinski*"
+date: "9/__/2022"
 abstract: >
@@
 library(raster)
+library(beepr)
 ```
@@
 color label images and another for CSV files. In this example, we are using 
-images of vegetative *Alexandrium catenella* cells. Please note that your 
+images of *Microcystis* sp. cells. Please note that your 
 directory addresses will be different. We recommend using a dedicated directory 
@@
 ```{r}
-savdir <- ("E:/cellcount_data/CSV_data/")
-image_savdir <- ("E:/cellcount_data/Convert_Images/Convert_Alexandrium10X/")
+savdir <- ("H:/CSV_data/")
+image_savdir <- ("H:/Convert_Images/F192_Test/")
 ```
@@
 the following code. Once again, your directories will be different and must be
-updated to the correct file address:
+updated to the correct file address. However, in this example we will be using
+the provided test images from this package:
 
 ```{r}
-images <- list.files("E:/cellcount_data/Alexandrium_10X/", pattern = "tif", full.name = T)
-images_names <- list.files("E:/cellcount_data/Alexandrium_10X/", pattern = "tif", full.name = F)
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 ```
@@
 ```{r}
-imgNames <- paste0("Alexandrium_", images_names)
+file_name<-paste0("F192_")
 ```
@@
 following code outlines this step using the "readTIFF" function from the 
-*tiff* package:
+*tiff* package. In addition, we'll need to re-orient the image as it's uploaded
+into the environment. The "readTIFF" function uploads images in a portrait 
+format - the "aperm" function reorients the uploaded image to a landscape format:
 
 ```{r, warning=FALSE}
+imgNames <- paste0(file_name, images_names)
 read_images <- lapply(images, readTIFF)
+img_transposed <- lapply(read_images,aperm,c(2,1,3))
 names(images) <- imgNames
@@
 ```{r}
-grey_images <- lapply(read_images, greyscale, contrast = 2)
-display(grey_images[[1]])
+grey_images <- lapply(img_transposed, greyscale, contrast = 1)
+display(grey_images[[5]])
 ```
@@
 (if used), or could involve making adjustments to the "image_convert" function,
-such as the width and height of the rectangular window or the pixel area
-threshold (i.e., removes cells or other objects under a certain pixel area).
-Observe your initial results and make adjustments where necessary:
+such as the width and height of the rectangular window, the pixel area
+threshold (i.e., removes cells or other objects under a certain pixel area), 
+or the tolerance and 'ext' variables. Observe your initial results as indicated
+below and make adjustments where necessary:
 
 ```{r}
-imagesConverted <- image_convert(imagesMapped[[1]], w = 15, h = 25, offset = 0.01, areathresh = 150)
+imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
 final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
@@
 analyze your images in bulk, one at a time. The for-loop uses the 
-"image_convert" function to produce unique color labels on cells and produce an
-image showcasing these labels using the "countImages" function. Finally, 
-"countCells" is used to count the number of unique color labels (i.e., cells).
-We also have integrated code to add these count results to the data frame made
-earlier, and also to save each unique color label image to the "image_savdir" 
-directory:
+"image_convert" function to produce unique color labels on cells within
+an image and to produce a final image showcasing these labels using the 
+"countImages" function. Finally, "countCells" is used to count the number of 
+unique color labels (i.e., cells). We also have integrated code to add these 
+count results to the data frame made earlier, and also to save each unique 
+color label image to the "image_savdir" directory:
 
@@
 for (z in 1:length(images)) {
-  imagesConverted <- image_convert(imagesMapped[[z]], w = 15, h = 25, offset = 0.01, areathresh = 150)
+  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
   final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
@@
   Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
-  # display single-cell images via image_covert
   analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
-  writeImage(final_img, files = paste0(image_savdir, analyzed_image))
+  writeImage(final_img, files = paste0(image_savdir, analyzed_image),compression=c("LZW"))
 }
@@
 ```{r}
+Cell.Count
+```
+
+Some images that were initially taken may be too poor of quality or contain
+high background fluorescence, among other errors. The code below analyzes 
+outliers from the produced data, within a range of two standard deviations 
+above and below the median. The "outlier" function tags which images are
+out of this range and dictates a warning within our data frame, which is shown 
+below. This alerts the end-user to perhaps re-image a particular sample, or
+to investigate the error further.
+
+```{r}
+Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
+Cell.Count$SD_range<-outlier()
 Cell.Count
@@
 cell.total <- sum(cell.total)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total,"null")
 ```
+
+*Important note* - *the "null" data produced in the code above (and others that*
+*follow) fills in the blanks from the previous data produced by the* 
+*"outlier" function.*
 
@@
 - Total number of images analyzed
-- Filtration area (mm2)
+- Filtration area (mm^2)
 - Volume filtered per image (mL)
@@
 ```{r}
-cell.den <- cell_density(cell.total, height = 203.2, width = 166.4, diameter = 2.5, images = 15, filtration.area = 491, volume = 1, total.volume = 100)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den)
+cell.den <- cell_density(cell.total, height = 203.2, width = 166.4, diameter = 2.5, images = 15,
+                         filtration.area = 491, volume = 1, total.volume = 100)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
+```
+
+We can also calculate the average number of cells per image, as well as cells
+per mL using the following code:
+
+```{r}
+cell.av<-(cell.total/10)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")
+
+cell.mL<-(cell.den/2) #change this via total volume
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")
+```
+
+We can now view our data frame to double check for any calculation errors:
+
+```{r}
+Cell.Count
 ```

### 2022-09-27 sandbox_analysis_test.R
Updated

@@ -15,10 +15,10 @@ file_name<-paste0("E.coli_") #Change the name of species analyzed here
 
 #Change directories here
 savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/CSV_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/Convert_Images/E.coli_100x_DAPI/1E/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/1E/"
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/Convert_Images/E.coli_100x_DAPI/3C/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/3C/"
                      , pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/1E/"
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/3C/"
                            , pattern = "tif", full.name = F)
 
 imgNames <- paste0(file_name, images_names)
@@ -66,15 +66,15 @@ cell.total <- sum(cell.total)
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total,"null")
 
 #Change variables outlined below
-cell.den <- cell_density(cell.total, FOV = 0.00084, images = 10, filtration.area = 213.8, volume = 0.05, total.volume = 2)
+cell.den <- cell_density(cell.total, FOV = 0.00084, images = 10, filtration.area = 213.8, volume = 0.02, total.volume = 1)
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
 
 cell.av<-(cell.total/10)
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")
 
-cell.mL<-(cell.den/2) #change this via total volume
+cell.mL<-(cell.den/1) #change this via total volume
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")
 
[…]

### 2022-09-27 sandbox_analysis_test.R
Updates

@@ -11,14 +11,14 @@ gc()
 
 Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
 
-file_name<-paste0("E.coli_") #Change the name of species analyzed here
+file_name<-paste0("F192_") #Change the name of species analyzed here
 
 #Change directories here
-savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/CSV_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/Convert_Images/E.coli_100x_DAPI/3C/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/3C/"
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/"
                      , pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/3C/"
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/"
                            , pattern = "tif", full.name = F)
 
 imgNames <- paste0(file_name, images_names)
@@ -34,13 +34,13 @@ imagesMapped <- lapply(grey_images, mapped, threshold = 0.3) #background intensi
 
 #IMAGE TESTING - do you need to make variable adjustments?
 
-imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 1, ext = 1)
 #Change variables above depending on species analyzed
 final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
 display(final_img)
 
 for (z in 1:length(images)) {
[…]

### 2022-09-27 sandbox_analysis_test.R
Updates

@@ -11,14 +11,14 @@ gc()
 
 Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
 
-file_name<-paste0("F192_") #Change the name of species analyzed here
+file_name<-paste0("Anabena_") #Change the name of species analyzed here
 
 #Change directories here
 savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
 image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/"
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/"
                      , pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/"
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/"
                            , pattern = "tif", full.name = F)
 
 imgNames <- paste0(file_name, images_names)
@@ -34,13 +34,13 @@ imagesMapped <- lapply(grey_images, mapped, threshold = 0.3) #background intensi
 
 #IMAGE TESTING - do you need to make variable adjustments?
 
-imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 1, ext = 1)
+imagesConverted <- image_convert2(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 1, ext = 1)
 #Change variables above depending on species analyzed
 final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
 display(final_img)
 
 for (z in 1:length(images)) {
-  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 1, ext = 1)
[…]

### 2022-09-27 sandbox_analysis_test.R
Updates

@@ -75,6 +75,6 @@ Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")
 cell.mL<-(cell.den/15) #change this via total volume
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")
 
-write.csv(Cell.Count, paste0(savdir, "/F192_test counts.csv")) #Change this CSV file name
+write.csv(Cell.Count, paste0(savdir, "/Anabena_test counts.csv")) #Change this CSV file name
 
 beepr::beep(sound=2) #analysis complete

### 2022-09-27 cellcount_intro.Rmd
Updates

@@
 
-```{r, message=FALSE}
 if (!requireNamespace("BiocManager", quietly = TRUE)) {
@@
 library(cellcount)
-```
 
@@
 images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 ```
@@
 ```{r}
+greyscale <- function(x, contrast = 2) {
+  x <- contrast * x
+  x <- x[, , 1] + x[, , 2] + x[, , 3]
+  x <- x / max(x)
+  x <- normalize(x, inputRange = c(0.1, 0.75))
+  return(x)
+}
 grey_images <- lapply(img_transposed, greyscale, contrast = 1)
@@
 ```{r}
+mapped <- function(x, threshold = 0.3) {
+  x <- as.matrix(x)
+  x[x < threshold] <- 0
+  return(x)
+}
 imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
@@
 ```{r}
+border <- function(x, y) {
+  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
+}
+countCells <- function(x) {
+  countAll <- max(x)
+  dims <- dim(x)
+  border1 <- border(x, dims)
+  ids <- unique(border1[which(border1 != 0)])
+  countInner <- countAll - length(ids)
+  return(countInner)
+}
+countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+  if (removeEdgeCells) {
+    dims <- dim(image)
+    border1 <- border(image, dims)
+    ids <- unique(border1[which(border1 != 0)])
+    inner <- rmObjects(image, ids)
+    EBImage::colorLabels(inner, normalize = normalize)
+  } else {
+    EBImage::colorLabels(image, normalize = normalize)
+  }
+}
+image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- fillHull(image)
+  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
 imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
@@
 Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
+outlier<-function(){
+  mean<-median(Cell.Count$Cell_Count)
+  sd<-sd(Cell.Count$Cell_Count)*2
+  sd_neg<-(mean-sd)
+  sd_pos<-(mean+sd)
+  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
+    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
+  return(SD_range_detection)
+}
 Cell.Count$SD_range<-outlier()
@@
 
-The following code uses these variables to calculate cell density and adds 
-the data into the "Cell.Count" data frame:
+Field of view (FOV) is an important variable calculated from image height,
+image width, and cell diameter variables. The following equation is listed
+below:
+
+$FOV = HW - 2D(H+W)+4D^{2}$
+
+Where:
+
+$H = height$
+
+$W = width$
+
+$D = diameter$
+
+Image height and width are variables that are taken from dimensions 
+disseminated by your microscope camera's manufacturer. Please refer to their 
+documentation to calculate your FOV variable correctly.
+
+With the variables calculated, the following code uses these variables to 
+calculate cell density and adds the data into the "Cell.Count" data frame:
 
 ```{r}
-cell.den <- cell_density(cell.total, height = 203.2, width = 166.4, diameter = 2.5, images = 15,
-                         filtration.area = 491, volume = 1, total.volume = 100)
+cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
+  MCF <- filtration.area / FOV
+  N <- x / images
+  cyano <- (MCF * N) / volume
+  cyano_total <- cyano * total.volume
+  return(cyano_total)
+}
+cell.den <- cell_density(cell.total, FOV = 0.00084, images = 10, filtration.area = 213.8, volume = 0.05, total.volume = 2)
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
@@
 ```{r}
-write.csv(Cell.Count, paste0(savdir, "/Alexandrium counts.csv"))
+write.csv(Cell.Count, paste0(savdir, "/F192_Test counts.csv"))
 ```
 
+# Potential errors
+
+End-users should note that potential errors with image analysis using
+*cellcount* may occur. Here, we outline a few examples of what may occur, and
+how to combat these errors to ensure the highest image quality for accurate
+enumeration results.
+
+## Extreme detail from fluorescence
+
+While high image quality is applauded when using this package, sometimes the
+natural auto-fluorescence may be too detailed, revealing potential problems
+when analyzing images. End-users may want to try different objectives in order
+to reduce detail as much as possible while able to show distinction between
+cells, or use another fluorescent dye (i.e. DAPI) in order to the program to
+produce more accurate enumeration results.
+
+![A greyscale image (left) of *Alexandrium catenella* and a binary filter 
+image (right) demonstrating the error results of high-detail from cells,
+causing incorrect enumeration results.](C:\Users\Tyler.Harman\Desktop\cellcount_work\misc_\Error Images\Error1.png)
+
+![A color-label image of *Prorocentrum* sp. where high details were shown from
+its fluorescing outer membrane, which caused the program to divide and 
+enumerate distinct regions shown here.](C:\Users\Tyler.Harman\Desktop\cellcount_work\misc_\Error Images\3214435.png)
+
+## High background intensities
+
+If high background intensities is observed within images, errors could occur
+like the one shown below, where the threshold adjustment built into the
+"mapped" function cannot separate the high intensity background from the
+majority of cells within the image. End-users should re-take these images and
+reduce the background intensity as much as possible. Snap images as rapidly as
+you can to avoid fluorescent dyes from increasing in fluorescent output - one
+of the contributing factors that we have observed throughout the development
+of this package.
+
+![A color-label image of *Shewanella* IRI-160 resulting from a high
+background intensity error. This caused the program to produce an 
+inconclusive image, due to the lack of being able to separate the high
+background intensity from the cells within the image.](C:\Users\Tyler.Harman\Desktop\cellcount_work\misc_\Error Images\542365.png)
+
 Literature

### 2022-09-27 sandbox_analysis_test.R
Updates

@@ -11,14 +11,14 @@ gc()
 
 Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
 
-file_name<-paste0("Anabena_") #Change the name of species analyzed here
+file_name<-paste0("E.coli_") #Change the name of species analyzed here
 
 #Change directories here
-savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/"
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/CSV_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/Convert_Images/E.coli_100x_DAPI/1C/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/1C/"
                      , pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/"
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/quantitative_images/E.coli_100x_DAPI/1C/"
                            , pattern = "tif", full.name = F)
 
 imgNames <- paste0(file_name, images_names)
@@ -34,13 +34,13 @@ imagesMapped <- lapply(grey_images, mapped, threshold = 0.3) #background intensi
 
 #IMAGE TESTING - do you need to make variable adjustments?
 
-imagesConverted <- image_convert2(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 1, ext = 1)
+imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
 #Change variables above depending on species analyzed
 final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
 display(final_img)
 
 for (z in 1:length(images)) {
[…]

### 2022-10-03 cellcount_intro.Rmd
Updates

--- 2022-09-20 cellcount_intro.Rmd
@@
 title: "Introduction to *cellcount*"
-author: "*T.E. Harman, M. Vandersea, D.E. Berthold, H.D. Laughinghouse, K. Pokrzywinski*"
-date: "9/__/2022"
+author: "*T.E. Harman, M. Vandersea, D.E. Berthold, H.D. Laughinghouse, B. Boyd, K. Pokrzywinski*"
+date: "__/__/2022"
 abstract: >
@@
 
-```{r}
-savdir <- ("H:/CSV_data/")
-image_savdir <- ("H:/Convert_Images/F192_Test/")
+```{r,message=FALSE}
+savdir <- ("C:/file path for saving CSV data")
+image_savdir <- ("C:/file path for saving converted images")
+```
+```{r,include=FALSE}
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
 ```
@@
 
-```{r}
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+```{r,message=FALSE}
+images <- list.files("C:/file path of images to analyze", pattern = "tif", full.name = T)
+images_names <- list.files("C:/file path of images to analyze", pattern = "tif", full.name = F)
+```
+```{r,include=FALSE}
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 ```
@@
 photo below). This function also has a built-in numerical contrast adjustment 
-that end-users can adjust. Note that if you already have 8-bit images, you can 
-skip this step:
+that end-users can adjust:
 
+```{r,include=FALSE}
+greyscale <- function(x, contrast = 2) {
+  x <- contrast * x
+  x <- x[, , 1] + x[, , 2] + x[, , 3]
+  x <- x / max(x)
+  x <- normalize(x, inputRange = c(0.1, 0.75))
+  return(x)
+}
+```
 ```{r}
@@
 
+```{r,include=FALSE}
+mapped <- function(x, threshold = 0.3) {
+  x <- as.matrix(x)
+  x[x < threshold] <- 0
+  return(x)
+}
+```
 ```{r}
@@
 
+```{r,include=FALSE}
+border <- function(x, y) {
+  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
+}
+countCells <- function(x) {
+  countAll <- max(x)
+  dims <- dim(x)
+  border1 <- border(x, dims)
+  ids <- unique(border1[which(border1 != 0)])
+  countInner <- countAll - length(ids)
+  return(countInner)
+}
+countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+  if (removeEdgeCells) {
+    dims <- dim(image)
+    border1 <- border(image, dims)
+    ids <- unique(border1[which(border1 != 0)])
+    inner <- rmObjects(image, ids)
+    EBImage::colorLabels(inner, normalize = normalize)
+  } else {
+    EBImage::colorLabels(image, normalize = normalize)
+  }
+}
+image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- fillHull(image)
+  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+```
 ```{r}
-imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
 final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
@@
 for (z in 1:length(images)) {
-  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
   final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
@@
 
+```{r,include=FALSE}
+outlier<-function(){
+  mean<-median(Cell.Count$Cell_Count)
+  sd<-sd(Cell.Count$Cell_Count)*2
+  sd_neg<-(mean-sd)
+  sd_pos<-(mean+sd)
+  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
+    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
+  return(SD_range_detection)
+}
+```
 ```{r}
@@
 
-The following code uses these variables to calculate cell density and adds 
-the data into the "Cell.Count" data frame:
+Field of view (FOV) is an important variable calculated from image height,
+image width, and cell diameter variables. The following equation is listed
+below:
 
+$FOV = HW - 2D(H+W)+4D^{2}$
+
+Where:
+
+$H = height$
+
+$W = width$
+
+$D = diameter$
+
+Image height and width are variables that are taken from dimensions 
+disseminated by your microscope camera's manufacturer. Please refer to their 
+documentation to calculate your FOV variable correctly.
+
+With the variables calculated, the following code uses these variables to 
+calculate cell density and adds the data into the "Cell.Count" data frame:
+
+```{r,include=FALSE}
+cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
+  MCF <- filtration.area / FOV
+  N <- x / images
+  cyano <- (MCF * N) / volume
+  cyano_total <- cyano * total.volume
+  return(cyano_total)
+}
+```
 ```{r}
-cell.den <- cell_density(cell.total, height = 203.2, width = 166.4, diameter = 2.5, images = 15,
-                         filtration.area = 491, volume = 1, total.volume = 100)
+cell.den <- cell_density(cell.total, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5)
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
@@
 We can also calculate the average number of cells per image, as well as cells
-per mL using the following code:
+per mL using the following code. Note that 'cell.av' must be divided by the
+number of images the end-user has analyzed, and the 'cell.mL' must be divided
+by the total volume from the original sample:
 
 ```{r}
-cell.av<-(cell.total/10)
+cell.av<-(cell.total/10) #change this via total number of images
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")
 
-cell.mL<-(cell.den/2) #change this via total volume
+cell.mL<-(cell.den/5) #change this via total volume
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")
@@
 ```{r}
-write.csv(Cell.Count, paste0(savdir, "/Alexandrium counts.csv"))
+write.csv(Cell.Count, paste0(savdir, "/F192_Test counts.csv"))
 ```
 
-Literature
+# Output data from test images in the *cellcount* package
+
+```{r,echo=FALSE}
+display(final_img)
+Cell.Count
+```
+Test data results for *Microcystis* strain F-192
+
+```{r,include=FALSE}
+# clear environment and free unused memory
+rm(list=ls())
+gc()
+```
+
+```{r,include=FALSE}
+greyscale <- function(x, contrast = 2) {
+  x <- contrast * x
+  x <- x[, , 1] + x[, , 2] + x[, , 3]
+  x <- x / max(x)
+  x <- normalize(x, inputRange = c(0.1, 0.75))
+  return(x)
+}
+mapped <- function(x, threshold = 0.3) {
+  x <- as.matrix(x)
+  x[x < threshold] <- 0
+  return(x)
+}
+border <- function(x, y) {
+  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
+}
+countCells <- function(x) {
+  countAll <- max(x)
+  dims <- dim(x)
+  border1 <- border(x, dims)
+  ids <- unique(border1[which(border1 != 0)])
+  countInner <- countAll - length(ids)
+  return(countInner)
+}
+countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+  if (removeEdgeCells) {
+    dims <- dim(image)
+    border1 <- border(image, dims)
+    ids <- unique(border1[which(border1 != 0)])
+    inner <- rmObjects(image, ids)
+    EBImage::colorLabels(inner, normalize = normalize)
+  } else {
+    EBImage::colorLabels(image, normalize = normalize)
+  }
+}
+image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
+  image2 <- fillHull(image1)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
+  MCF <- filtration.area / FOV
+  N <- x / images
+  cyano <- (MCF * N) / volume
+  cyano_total <- cyano * total.volume
+  return(cyano_total)
+}
+```
+
+```{r,echo=FALSE}
+Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
+
+file_name<-paste0("img_")
+
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)
+
+imgNames <- paste0(file_name, images_names)
+read_images <- lapply(images, readTIFF)
+img_transposed <- lapply(read_images,aperm,c(2,1,3))
+names(images) <- imgNames
+
+grey_images <- lapply(img_transposed, greyscale, contrast = 1)
+
+imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
+
+imagesConverted <- image_convert2(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
+final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+display(final_img)
+```
+
+```{r,echo=FALSE}
+for (z in 1:length(images)) {
+  imagesConverted <- image_convert2(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
+  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+  count <- countCells(imagesConverted)
+  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
+  analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
+  writeImage(final_img, files = paste0(image_savdir, analyzed_image),compression=c("LZW"))
+}
+
+Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
+outlier<-function(){
+  mean<-median(Cell.Count$Cell_Count)
+  sd<-sd(Cell.Count$Cell_Count)*2
+  sd_neg<-(mean-sd)
+  sd_pos<-(mean+sd)
+  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
+    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
+  return(SD_range_detection)
+}
+Cell.Count$SD_range<-outlier()
+
+cell.total <- as.numeric(Cell.Count$Cell_Count)
+cell.total <- sum(cell.total)
+
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total,"null")
+
+cell.den <- cell_density(cell.total, FOV = 0.027770534, images = 10, filtration.area = 213.8, volume = 0.5, total.volume =15)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
+
+cell.av<-(cell.total/10)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")
+
+cell.mL<-(cell.den/15)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")
+
+Cell.Count
+```
+Test data results for *Anabena* sp.
+
+```{r,include=FALSE}
+# clear environment and free unused memory
+rm(list=ls())
+gc()
+```
+
+```{r,include=FALSE}
+greyscale <- function(x, contrast = 2) {
+  x <- contrast * x
+  x <- x[, , 1] + x[, , 2] + x[, , 3]
+  x <- x / max(x)
+  x <- normalize(x, inputRange = c(0.1, 0.75))
+  return(x)
+}
+mapped <- function(x, threshold = 0.3) {
+  x <- as.matrix(x)
+  x[x < threshold] <- 0
+  return(x)
+}
+border <- function(x, y) {
+  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
+}
+countCells <- function(x) {
+  countAll <- max(x)
+  dims <- dim(x)
+  border1 <- border(x, dims)
+  ids <- unique(border1[which(border1 != 0)])
+  countInner <- countAll - length(ids)
+  return(countInner)
+}
+countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+  if (removeEdgeCells) {
+    dims <- dim(image)
+    border1 <- border(image, dims)
+    ids <- unique(border1[which(border1 != 0)])
+    inner <- rmObjects(image, ids)
+    EBImage::colorLabels(inner, normalize = normalize)
+  } else {
+    EBImage::colorLabels(image, normalize = normalize)
+  }
+}
+image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- fillHull(image)
+  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
+  MCF <- filtration.area / FOV
+  N <- x / images
+  cyano <- (MCF * N) / volume
+  cyano_total <- cyano * total.volume
+  return(cyano_total)
+}
+```
+
+```{r,echo=FALSE}
+Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
+
+file_name<-paste0("img_")
+
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)
+
+imgNames <- paste0(file_name, images_names)
+read_images <- lapply(images, readTIFF)
+img_transposed <- lapply(read_images,aperm,c(2,1,3))
+names(images) <- imgNames
+
+grey_images <- lapply(img_transposed, greyscale, contrast = 1)
+
+imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
+
+imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+display(final_img)
+```
+
+```{r,echo=FALSE}
+for (z in 1:length(images)) {
+  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+  count <- countCells(imagesConverted)
+  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
+  analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
+  writeImage(final_img, files = paste0(image_savdir, analyzed_image),compression=c("LZW"))
+}
+
+Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
+outlier<-function(){
+  mean<-median(Cell.Count$Cell_Count)
+  sd<-sd(Cell.Count$Cell_Count)*2
+  sd_neg<-(mean-sd)
+  sd_pos<-(mean+sd)
+  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
+    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
+  return(SD_range_detection)
+}
+Cell.Count$SD_range<-outlier()
+
+cell.total <- as.numeric(Cell.Count$Cell_Count)
+cell.total <- sum(cell.total)
+
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total,"null")
+
+cell.den <- cell_density(cell.total, FOV = 0.00042, images = 10, filtration.area = 213.8, volume = 0.02, total.volume =2)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
+
+cell.av<-(cell.total/10)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")
+
+cell.mL<-(cell.den/2)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")
+
+Cell.Count
+```
+Test data results for *Shewanella* IRI-160
+
+# Literature cited:

### 2022-10-03 R/border.R
Updated - removal of examples in documentation

@@ -7,14 +7,7 @@
 #' @param x An image file
 #' @param y Dimming variable (using the 'dim' function from EBImage)
 #' @return Analyzing the cells on the border of \code{x}
-#' @examples
-#' img0 <- list.files("data/Anabena 2021/UV", pattern = "tif", full.name = TRUE)
-#' img1 <- lapply(img0, readTIFF)
-#' img2 <- lapply(img1, greyscale)
-#' img3 <- lapply(img2, mapped)
-#' img4 <- image_convert(img3)
-#' dims <- dim(img4)
-#' border(img4, dims)
+
 border <- function(x, y) {
   c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
 }

### 2022-10-03 R/cell_density.R

@@ -5,22 +5,14 @@
 #' filtered that the image is visualizing.
 #'
 #' @param x count results from processed image
-#' @param height height of image field of view in μm (comes from camera settings)
-#' @param width width of image field of view in μm (comes from camera settings)
-#' @param diameter median cell diameter in μm
+#' @param FOV field of view from camera specs in mm2
 #' @param images total number of fields
 #' @param filtration.area filtration area in mm2
 #' @param volume volume filtered per field in mL
 #' @param total.volume total sample volume in mL
 #'
 #' @examples
-#' img <- list.files("data/Shewanella", pattern = "tif", full.name = T)
-#' img1 <- lapply(img[[1]], readTIFF)
-#' img2 <- lapply(img1, greyscale, contrast = 2)
-#' img3 <- lapply(img2, mapped, threshold = 0.3)
-#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.01, areathresh = 150)
-#' cell.total <- countCells(img4)
-#' cell_density(cell.total, height = 203.2, width = 166.4, diameter = 2.5, images = 15, filtration.area = 491, volume = 1, total.volume = 100)
+
 cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
   MCF <- filtration.area / FOV
   N <- x / images

### 2022-10-03 R/countCells.R

@@ -8,12 +8,7 @@
 #' @return Counting cells within \code{x}
 #'
 #' @examples
-#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
-#' img1 <- lapply(img[[1]], readTIFF)
-#' img2 <- lapply(img1, greyscale, contrast = 2)
-#' img3 <- lapply(img2, mapped, threshold = 0.3)
-#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.01, areathresh = 150)
-#' countCells(img4)
+
 countCells <- function(x) {
   countAll <- max(x)
   dims <- dim(x)

### 2022-10-03 R/countImages.R

@@ -10,13 +10,7 @@
 #' border
 #'
 #' @examples
-#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
-#' img1 <- lapply(img[[1]], readTIFF)
-#' img2 <- lapply(img1, greyscale, contrast = 2)
-#' img3 <- lapply(img2, mapped, threshold = 0.3)
-#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.01, areathresh = 150)
-#' final_img <- countImages(img4, normalize = T, removeEdgeCells = T)
-#' display(final_img)
+
 countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
   if (removeEdgeCells) {
     dims <- dim(image)

### 2022-10-03 R/greyscale.R

@@ -8,9 +8,7 @@
 #' @param contrast a numerical value for contrast adjustment
 #'
 #' @examples
-#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
-#' img1 <- lapply(img[[1]], readTIFF)
-#' img2 <- lapply(img1, greyscale, contrast = 2)
+
 greyscale <- function(x, contrast = 2) {
   x <- contrast * x
   x <- x[, , 1] + x[, , 2] + x[, , 3]

### 2022-10-03 R/image_convert.R

@@ -2,20 +2,18 @@
 #'
 #' Takes a mapped image (i.e., data matrix) and counts individual cells within an image.
 #' This is to be used with images with a single image that contains a single
-#' fluorescent dimension of cells. Used for single-cell type images.
+#' fluorescent dimension of cells. Best used for single-cell type images.
 #'
 #' @param x an image variable that has been mapped using the "mapped" function
 #' @param w width of the moving rectangular window (see EBImage)
 #' @param h heigth of the moving rectangular window (see EBImage)
 #' @param offset a thresholding offset
 #' @param areathresh a numerical value of individual pixel areas of cells - this sets a threshold for what items are removed
+#' @param tolerance The minimum height of the object in the units of image intensity between its highest point and the point where it contacts another object (see EBImage)
+#' @param ext Radius of the neighborhood in pixels for the detection of neighboring objects. Higher value smooths out small objects (see EBImage)
 #'
 #' @examples
-#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
-#' img1 <- lapply(img[[1]], readTIFF)
-#' img2 <- lapply(img1, greyscale, contrast = 2)
-#' img3 <- lapply(img2, mapped, threshold = 0.3)
-#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.001, areathresh = 150)
+
 image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
   image <- thresh(x, w = w, h = h, offset = offset)
   image1 <- fillHull(image)

### 2022-10-03 R/image_convert2.R

@@ -2,7 +2,7 @@
 #'
 #' Takes a mapped image (i.e., data matrix) and counts individual cells within an image.
 #' This is to be used with images with a single image that contains a single
-#' fluorescent dimension of cells. Used for filamentous/colonial-type cell
+#' fluorescent dimension of cells. Best used for filamentous/colonial-type cells
 #' images.
 #'
 #' @param x an image variable that has been mapped using the "mapped" function
@@ -10,13 +10,11 @@
 #' @param h heigth of the moving rectangular window (see EBImage)
 #' @param offset a thresholding offset
 #' @param areathresh a numerical value of individual pixel areas of cells - this sets a threshold for what items are removed
+#' @param tolerance The minimum height of the object in the units of image intensity between its highest point and the point where it contacts another object (see EBImage)
+#' @param ext Radius of the neighborhood in pixels for the detection of neighboring objects. Higher value smooths out small objects (see EBImage)
 #'
 #' @examples
-#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
-#' img1 <- lapply(img[[1]], readTIFF)
-#' img2 <- lapply(img1, greyscale, contrast = 2)
-#' img3 <- lapply(img2, mapped, threshold = 0.3)
-#' img4 <- image_convert(img3, w = 15, h = 25, offset = 0.001, areathresh = 150)
+
 
 image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
   image <- thresh(x, w = w, h = h, offset = offset)

### 2022-10-03 R/mapped.R

@@ -8,10 +8,7 @@
 #' @param threshold a numerical value of pixel intensity
 #'
 #' @examples
-#' img <- list.files("data/Karlodinium", pattern = "tif", full.name = T)
-#' img1 <- lapply(img[[1]], readTIFF)
-#' img2 <- lapply(img1, greyscale, contrast = 2)
-#' img3 <- lapply(img2, mapped, threshold = 0.3)
+
 mapped <- function(x, threshold = 0.3) {
   x <- as.matrix(x)
   x[x < threshold] <- 0

### 2022-10-03 R/outline.R
No longer needed - former initial analysis function

@@ -1,16 +1,0 @@
-#' Outline conversion
-#'
-#' Takes grey-scale images to convert to an outline format for the 'propagation'
-#' function. Needed for the "multi_image" function.
-#' @param x normalized image variable (see example)
-#' @examples
-#' image_greyF2 <- readImage(paste0(grey_pathF2, grey_filesF2[f]))
-#' ImgF2 <- normalize(image_greyF2, inputRange = c(0, 0.75))
-#' F2_lab <- outline(ImgF2)
-outline <- function(x) {
-  x <- thresh(x, w = 15, h = 15, offset = 0.05)
-  x <- opening(x, makeBrush(5, shape = "disc"))
-  x <- fillHull(x)
-  x <- bwlabel(x)
-  return(x)
-}

### 2022-10-04 stats_code/BA_Code.R
Updated

@@ -12,26 +12,29 @@ BA_F210<-subset(BA,Species=="F210")
 BA_Anabena<-subset(BA,Species=="Anabena")
 BA_Anabena$Concentration_f = factor(BA_Anabena$Concentration, levels=c('High','Medium','Low'))
 BA_Shewanella<-subset(BA,Species=="Shewanella")
+BA_E.coli<-subset(BA,Species=="E.coli")
 
 
 BA_F192$Output<-(BA_F192$Output/1000000)
 BA_F192$Output<-(sqrt(BA_F192$Output))
 
-F192<-ggplot(data=BA_F192, aes(x=Concentration,y=Output,color=Method))+
-  scale_colour_grey()+
+F192<-ggplot(data=BA_F192, aes(x=Concentration,y=Output,fill=Method))+
+  scale_fill_grey(start=0.5,end=0.9)+
   stat_boxplot(geom='errorbar')+
   geom_boxplot(size=0.5)+
   theme_bw()+
   labs(x="Concentration")+
-  labs(y=expression(paste("Manual counts ( ",x10^{6},") (sqrt transformed)")))+
+  labs(y=expression(paste("Cells per mL ( ",x10^{6},") (sqrt transformed)")))+
   scale_y_continuous(labels = function(x) format(x, scientific = FALSE))+
+  theme(legend.text = element_text(size=13),
+        legend.title = element_text(size=13))+
   theme(axis.title.x = element_text(size=15),
         axis.title.y = element_text(size=13,vjust=2))+
   theme(axis.text.x = element_text(size = 15),
         axis.text.y = element_text(size = 15),
         axis.ticks = element_line(size=1.5),
         axis.ticks.length = unit(0.25,"cm"),
-        plot.margin=unit(c(1,1,1,1),"cm"),
+        plot.margin=unit(c(1,0.25,1,1),"cm"),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         panel.border = element_rect(size=1.5))
@@ -40,74 +43,106 @@ F192
 BA_F210$Output<-(BA_F210$Output/1000000)
 BA_F210$Output<-(sqrt(BA_F210$Output))
 
-F210<-ggplot(data=BA_F210, aes(x=Concentration,y=Output,color=Method))+
-  scale_colour_grey()+
+F210<-ggplot(data=BA_F210, aes(x=Concentration,y=Output,fill=Method))+
[…]

### 2022-10-04 stats_code/RA_Code.R

@@ -11,6 +11,7 @@ RA_F192<-subset(RA,Species=="F192")
 RA_F210<-subset(RA,Species=="F210")
 RA_Anabena<-subset(RA,Species=="Anabena")
 RA_Shewanella<-subset(RA,Species=="Shewanella")
+RA_E.coli<-subset(RA,Species=="E.coli")
 
 RA_F192$Script<-(RA_F192$Script/1000000)
 RA_F192$Visual<-(RA_F192$Visual/1000000)
@@ -22,7 +23,7 @@ summary(data.lm1)
 F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
   theme_bw()+
   labs(x=expression(paste(italic("cellcount"), " results (",x10^{6},")")))+
-  labs(y=expression(paste("Manual counts ( ",x10^{6},")")))+
+  labs(y=expression(paste("Visual counts ( ",x10^{6},")")))+
   theme(axis.title.x = element_text(size=15),
         axis.title.y = element_text(size=15,vjust=2))+
   theme(axis.text.x = element_text(size = 15),
@@ -32,17 +33,17 @@ F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
         plot.margin=unit(c(1,1,1,1),"cm"),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
-        panel.border = element_rect(size=1.5))+
+        panel.border = element_rect(size=2))+
   scale_x_continuous(labels=function(x) format(x,scientific=FALSE))+
   scale_y_continuous(labels=function(x) format(x,scientific=FALSE))+
   xlim(0,5)+
   ylim(0,5)+
-  geom_abline(slope=1,intercept=0,color="red",size=1,alpha=0.4)+
+  geom_abline(slope=1,intercept=0,color="grey40",size=1.5,alpha=0.4)+
   geom_abline(slope=coef(data.lm1)[["Script"]],
               intercept=coef(data.lm1)[["(Intercept)"]],
               linetype=2,
-              size=1)+
-  geom_point(size=3,shape=21,colour="black",fill="grey60",stroke=1.5)
+              size=2)+
+  geom_point(size=4,shape=21,colour="black",fill="grey60",stroke=1.5)
 F192
 
 
@@ -56,7 +57,7 @@ summary(data.lm2)
 F210<-ggplot(data=RA_F210, aes(x=Script,y=Visual))+
   theme_bw()+
[…]

### 2022-10-04 stats_code/RA_IND_Code.R

@@ -11,6 +11,7 @@ RA_F192<-subset(RA,Species=="F192")
 RA_F210<-subset(RA,Species=="F210")
 RA_Anabena<-subset(RA,Species=="Anabena")
 RA_Shewanella<-subset(RA,Species=="Shewanella")
+RA_E.coli<-subset(RA,Species=="E.coli")
 
 data.lm1<-lm(Visual~Script,data=RA_F192)
 summary(data.lm1)
@@ -18,7 +19,7 @@ summary(data.lm1)
 F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
   theme_bw()+
   labs(x=expression(paste(italic("cellcount"), " results")))+
-  labs(y=expression(paste("Manual counts")))+
+  labs(y=expression(paste("Visual counts")))+
   theme(axis.title.x = element_text(size=15),
         axis.title.y = element_text(size=15,vjust=2))+
   theme(axis.text.x = element_text(size = 15),
@@ -28,17 +29,18 @@ F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
         plot.margin=unit(c(1,1,1,1),"cm"),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
-        panel.border = element_rect(size=1.5))+
+        panel.border = element_rect(size=2))+
   scale_x_continuous(labels=function(x) format(x,scientific=FALSE))+
   scale_y_continuous(labels=function(x) format(x,scientific=FALSE))+
   xlim(0,500)+
   ylim(0,500)+
-  geom_abline(slope=1,intercept=0,color="red",size=1,alpha=0.4)+
+  geom_abline(slope=1,intercept=0,color="grey40",size=1.5,alpha=0.4)+
+  geom_smooth(method=lm,level=0.9,color=NA, fullrange=TRUE)+
   geom_abline(slope=coef(data.lm1)[["Script"]],
               intercept=coef(data.lm1)[["(Intercept)"]],
               linetype=2,
-              size=1)+
-  geom_point(size=3,shape=21,colour="black",fill="grey60",stroke=1.5)
+              size=1.5)+
+  geom_point(size=4,shape=21,colour="black",fill="grey60",stroke=1.5)
 F192
 
 data.lm2<-lm(Visual~Script,data=RA_F210)
@@ -47,7 +49,7 @@ summary(data.lm2)
 F210<-ggplot(data=RA_F210, aes(x=Script,y=Visual))+
   theme_bw()+
[…]

### 2022-10-04 stats_code/RosnerTest.R

@@ -1,7 +1,7 @@
 library(outliers)
 library(EnvStats)
 
-setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/stat_code/")
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data/")
 
 RA<-read.csv("Regression_Analysis_IND.csv",header=T)
 
@@ -73,3 +73,22 @@ rosnerTest(RA_ShewanellaB$Visual)
 rosnerTest(RA_ShewanellaC$Visual)
 rosnerTest(RA_ShewanellaD$Visual)
 rosnerTest(RA_ShewanellaE$Visual)
+
+
+RA_E.coli<-subset(RA,Species=="E.coli")
+RA_E.coliA<-subset(RA_E.coli,Concentration=="A")
+RA_E.coliB<-subset(RA_E.coli,Concentration=="B")
+RA_E.coliC<-subset(RA_E.coli,Concentration=="C")
+RA_E.coliD<-subset(RA_E.coli,Concentration=="D")
+RA_E.coliE<-subset(RA_E.coli,Concentration=="E")
+
+rosnerTest(RA_E.coliA$Script)
+rosnerTest(RA_E.coliB$Script)
+rosnerTest(RA_E.coliC$Script)
+rosnerTest(RA_E.coliD$Script)
+rosnerTest(RA_E.coliE$Script)
+rosnerTest(RA_E.coliA$Visual)
+rosnerTest(RA_E.coliB$Visual)
+rosnerTest(RA_E.coliC$Visual)
+rosnerTest(RA_E.coliD$Visual)
+rosnerTest(RA_E.coliE$Visual)

### 2022-10-05 EBImage_Markdown.Rmd
Initial upload
Updated

--- 2022-09-20 cellcount_intro.Rmd
@@
 
+# Overview:
+
+Climate change and anthropogenic impacts have exacerbated the presence of 
+harmful algal blooms (HABs) within marine and freshwater environments. 
+Rising temperatures, stratification strength, reduction of vertical mixing, 
+and increases in nutrients attribute to these increases in blooms. Routine 
+monitoring of HABs has become important in order to understand ecosystem and 
+public health risk and establish mitigation practices.
+
+Standard methods involve the enumeration of HAB species using microscopy 
+methods (i.e. hemocytometer), which is considered the gold standard, as well 
+as other methods such as qPCR and flow cytometry. These methods can be time 
+consuming and require significant knowledge related to these technologies. 
+Technology has advanced enumeration needs to decrease time consumption while 
+also maintaining accuracy. Instruments such as the FlowCam&trade; and Imaging 
+Flow Cytobot (IFCB) have been developed for these purposes; however, these 
+instruments are expensive to purchase, operate, and maintain. The need for 
+open-source technologies is important for accessibility reasons, but also for 
+end-user customization needs.
+
+Our project sought to solve the two mentioned issues on developing enumeration 
+technology for open-source capabilities: accessibility and customization. A 
+prior report released by Pokrzywinski et al. (2019) had outlined the 
+development of an automated cyanobacteria enumeration script using the 
+open-source programming language, R, along with several functions from the 
+R package *EBImage*. However, the previous presentation of 
+this script can be considered extremely difficult to grasp for inexperienced 
+R users. We utilized this script as a template for the development of an R 
+package, involving the reduction of code into simple functions, expanding 
+the ability for end-user modifications, development of a vignette for 
+inexperienced users, and promoting open-source availability. Along with these 
+goals, we aim to establish a rapid and accurate enumeration method for 
+researchers, where significantly less time is consumed compared to the 
+standard enumeration methods available.
+
 ```{r setup, include=FALSE}
@@
 
-*cellcount* is an R package distributed as part of the [Bioconductor](http://bioconductor.org) project. To install the package, start R and enter:
-
+The following packages are required. Please install and bring them into the R
+environment: 
 ```{r, message=FALSE}
@@
 }
-if (!require("cellcount")) {
-  BiocManager::install("cellcount")
-}
-library(cellcount)
-```
-
-The following packages are required. Please install and bring them into the R
-environment: 
-```{r, message=FALSE}
 if (!require("EBImage")) {
@@
 library(raster)
-library(beepr)
 ```
@@
 color label images and another for CSV files. In this example, we are using 
-images of *Microcystis* sp. cells. Please note that your 
-directory addresses will be different. We recommend using a dedicated directory 
-for *cellcount* results:
+images of *Microcystis* sp. cells that are contained within the package as test
+data. Please note that your directory addresses will be different. We recommend 
+using a dedicated directory for *cellcount* results:
 
-```{r}
-savdir <- ("H:/CSV_data/")
-image_savdir <- ("H:/Convert_Images/F192_Test/")
+```{r,message=FALSE}
+savdir <- ("file path for saving CSV data")
+image_savdir <- ("file path for saving converted images")
+```
+```{r,include=FALSE}
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
 ```
@@
 
-```{r}
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+```{r,message=FALSE}
+images <- list.files("file path of images to analyze", pattern = "tif", full.name = T)
+images_names <- list.files("file path of images to analyze", pattern = "tif", full.name = F)
+```
+```{r,include=FALSE}
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 ```
@@
 photo below). This function also has a built-in numerical contrast adjustment 
-that end-users can adjust. Note that if you already have 8-bit images, you can 
-skip this step:
+that end-users can adjust:
 
+```{r}
+greyscale <- function(x, contrast = 2) {
+  x <- contrast * x
+  x <- x[, , 1] + x[, , 2] + x[, , 3]
+  x <- x / max(x)
+  x <- normalize(x, inputRange = c(0.1, 0.75))
+  return(x)
+}
+```
 ```{r}
@@
 Next, we will use the "mapped" function to turn the 8-bit image we just
-converted into a data matrix using the following code. This also includes a
-pixel intensity threshold adjustment that end-users can change; this removes
-any pixel that is under the threshold value to easily separate the background
-from cells:
+converted into a data matrix (using 'as.matrix') using the following code. 
+This also includes a pixel intensity threshold adjustment that end-users can 
+change; this removes any pixel that is under the threshold value to easily 
+separate the background from cells:
 
+```{r}
+mapped <- function(x, threshold = 0.3) {
+  x <- as.matrix(x)
+  x[x < threshold] <- 0
+  return(x)
+}
+```
 ```{r}
@@
 
-Before we analyze our images in bulk, set aside some code space to visually 
+Before we analyze our images in bulk, we introduce here several functions
+that are required to take matrix data to count cells within images and to
+produce 'color-label' images. These include the following:
+
+**"countImages" function:** Displays converted microscopy images of distinct 
+cell counts subsequently from the image_convert function. Can use the "display" 
+function from EBImage to automatically generate the produced image from this 
+function.
+
+```{r}
+countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+  if (removeEdgeCells) {
+    dims <- dim(image)
+    border1 <- border(image, dims)
+    ids <- unique(border1[which(border1 != 0)])
+    inner <- rmObjects(image, ids)
+    EBImage::colorLabels(inner, normalize = normalize)
+  } else {
+    EBImage::colorLabels(image, normalize = normalize)
+  }
+}
+```
+
+**"border" function:** Takes images to predict cells along a predetermined 
+border in order to count cells within the predetermined border. Cells along the
+predetermined border are not counted, which is outlined in the "countCells"
+function.
+
+```{r}
+border <- function(x, y) {
+  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
+}
+```
+
+**"countCells" function:** Takes information from a converted image and counts 
+cells within the image. This uses the "border" function to remove cells along 
+the image border and only count cells within the image.
+
+```{r}
+countCells <- function(x) {
+  countAll <- max(x)
+  dims <- dim(x)
+  border1 <- border(x, dims)
+  ids <- unique(border1[which(border1 != 0)])
+  countInner <- countAll - length(ids)
+  return(countInner)
+}
+```
+
+**"image_convert" function:** Takes a mapped image (i.e., data matrix) and 
+counts individual cells within an image. This is to be used with a single image 
+with cells that contains a single fluorescent dimension 
+(i.e. auto-fluorescence, DAPI, etc.). This is useful for analyzing images
+containing single-cell cyanobacteria.
+
+```{r}
+image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- fillHull(image)
+  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+```
+
+We also include a separate, yet similar function to "image_convert" called 
+"image_convert2". This function is ideally meant for filamentous or 
+colonial-type cyanobacteria that are harder to discern with the standard 
+"image_convert" function. The difference between these functions are that 
+the "watershed" function from the EBImage package comes first in the outlined 
+function compared to the "image_convert" function.
+
+```{r}
+image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
+  image2 <- fillHull(image1)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+```
+
+In addition, set aside some code space to visually 
 analyze your images results. This could include adjusting the threshold
@@
 ```{r}
-imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, 
+                                 tolerance = 0.8, ext = 1)
 final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
@@
 for (z in 1:length(images)) {
-  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, 
+                                   tolerance = 0.8, ext = 1)
   final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
@@
 
+```{r}
+outlier<-function(){
+  mean<-median(Cell.Count$Cell_Count)
+  sd<-sd(Cell.Count$Cell_Count)*2
+  sd_neg<-(mean-sd)
+  sd_pos<-(mean+sd)
+  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
+    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
+  return(SD_range_detection)
+}
+```
 ```{r}
@@
 
-The following code uses these variables to calculate cell density and adds 
-the data into the "Cell.Count" data frame:
+Field of view (FOV) is an important variable calculated from image height,
+image width, and cell diameter variables. The following equation is listed
+below:
+
+$FOV = HW - 2D(H+W)+4D^{2}$
+
+Where:
+
+$H = height$
+
+$W = width$
+
+$D = diameter$
+
+Image height and width are variables that are taken from dimensions 
+disseminated by your microscope camera's manufacturer. Please refer to their 
+documentation to calculate your FOV variable correctly.
+
+With the variables calculated, the following "cell_density" uses these 
+variables to calculate cell density and adds the data into the "Cell.Count" 
+data frame:
 
 ```{r}
-cell.den <- cell_density(cell.total, height = 203.2, width = 166.4, diameter = 2.5, images = 15,
-                         filtration.area = 491, volume = 1, total.volume = 100)
+cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, 
+                         volume = 1, total.volume = 5) {
+  MCF <- filtration.area / FOV
+  N <- x / images
+  cyano <- (MCF * N) / volume
+  cyano_total <- cyano * total.volume
+  return(cyano_total)
+}
+```
+```{r}
+cell.den <- cell_density(cell.total, FOV = 0.0726, images = 10, filtration.area = 213.8, 
+                         volume = 1, total.volume = 5)
 Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
@@
 ```{r}
-write.csv(Cell.Count, paste0(savdir, "/Alexandrium counts.csv"))
+write.csv(Cell.Count, paste0(savdir, "/F192_Test counts.csv"))
 ```
 
-Literature
+#Literature cited:

### 2022-12-02 EBImage_Markdown.Rmd
Updated

@@
 abstract: >
-  The *cellcount* package provides wrapper functions for  [*EBImage*](http://bioconductor.org/packages/devel/bioc/vignettes/EBImage/inst/doc/EBImage-introduction.html) to simplfy quantifying bacteria and cyanobacteria via flourescence microscopy.
+  The ***cellcount*** package provides wrapper functions for  [***EBImage***](http://bioconductor.org/packages/devel/bioc/vignettes/EBImage/inst/doc/EBImage-introduction.html) to simplfy quantifying bacteria and cyanobacteria via flourescence microscopy.
 Cyanobacteria enumeration is an important tool to incorporate data results into mathematical models to assess bloom dynamics.
@@
 open-source programming language, R, along with several functions from the 
-R package *EBImage*. However, the previous presentation of 
+R package ***EBImage***. However, the previous presentation of 
 this script can be considered extremely difficult to grasp for inexperienced 
@@
 
-Next, generate a data frame for count data to be added into using the following
-code:
+To begin, we generate a data frame for the proceeding count data to be added 
+into using the following code:
 ```{r}
@@
 
-You'll need to generate a couple of directories; one for saving the unique
-color label images and another for CSV files. In this example, we are using 
-images of *Microcystis* sp. cells that are contained within the package as test
-data. Please note that your directory addresses will be different. We recommend 
-using a dedicated directory for *cellcount* results:
+Next, end-users will need to generate a couple of directories; one for saving the unique color label images and another for CSV files.
+In this example, we are using images of *Microcystis* sp.
+cells that are contained within the package as test data.
+Please note that your directory addresses will be different.
+We recommend using a dedicated directory for ***cellcount*** results:
 
@@
 
-Next, you'll need to read in the images you are interested in processing using
-the following code. Once again, your directories will be different and must be
-updated to the correct file address. However, in this example we will be using
+Next, end-users will need to read in the images you are interested in processing using the following code.
+Once again, your directories will be different and must be updated to the correct file address.
+However, in this example we will be using
 the provided test images from this package:
@@
 
-In addition, you'll need to assign the file names to a vector along with adding
-in characterization to each file for easy organization:
+In addition, end-users will need to assign the file names to a vector; this provides characterization to each file for easy organization:
 
@@
 Now we will "read" in the images that we are interested in analyzing. The
-following code outlines this step using the "readTIFF" function from the 
-*tiff* package. In addition, we'll need to re-orient the image as it's uploaded
-into the environment. The "readTIFF" function uploads images in a portrait 
-format - the "aperm" function reorients the uploaded image to a landscape format:
+following code outlines this step using the "*readTIFF*" function from the 
+***tiff*** package.
+In addition, we'll need to re-orient the image as it's uploaded into the environment.
+The "*readTIFF*" function uploads images in a portrait format - the "*aperm*" function reorients the uploaded image to a landscape format:
 
@@
 images into useful data that R can use to process images. First, we begin by
-converting 24-bit images into 8-bit images using the "greyscale" function (see 
+converting 24-bit images into 8-bit images using the "*greyscale*" function (see 
 photo below). This function also has a built-in numerical contrast adjustment 
@@
 
-Next, we will use the "mapped" function to turn the 8-bit image we just
-converted into a data matrix (using 'as.matrix') using the following code. 
+Next, we will use the "*mapped*" function to turn the 8-bit image we just
+converted into a data matrix (using "*as.matrix*") using the following code. 
 This also includes a pixel intensity threshold adjustment that end-users can 
@@
 that are required to take matrix data to count cells within images and to
-produce 'color-label' images. These include the following:
+produce "*color-label*" images. These include the following:
 
-**"countImages" function:** Displays converted microscopy images of distinct 
-cell counts subsequently from the image_convert function. Can use the "display" 
-function from EBImage to automatically generate the produced image from this 
-function.
+**"*countImages*" function:** Displays converted microscopy images of distinct 
+cell counts subsequently from the "*image_convert function*". End-users can use the "*display*" function from ***EBImage*** to visually generate the produced image from this function.
 
@@
 
-**"border" function:** Takes images to predict cells along a predetermined 
+**"*border*" function:** Takes images to predict cells along a predetermined 
 border in order to count cells within the predetermined border. Cells along the
-predetermined border are not counted, which is outlined in the "countCells"
+predetermined border are not counted, which is outlined in the "*countCells*"
 function.
@@
 
-**"countCells" function:** Takes information from a converted image and counts 
-cells within the image. This uses the "border" function to remove cells along 
+**"*countCells*" function:** Takes information from a converted image and counts 
+cells within the image. This uses the "*border*" function to remove cells along 
 the image border and only count cells within the image.
@@
 
-**"image_convert" function:** Takes a mapped image (i.e., data matrix) and 
+**"*image_convert*" function:** Takes a mapped image (i.e., data matrix) and 
 counts individual cells within an image. This is to be used with a single image 
@@
 
-We also include a separate, yet similar function to "image_convert" called 
-"image_convert2". This function is ideally meant for filamentous or 
+We also include a separate, yet similar function to "*image_convert*" called 
+"*image_convert2*". This function is ideally meant for filamentous or 
 colonial-type cyanobacteria that are harder to discern with the standard 
-"image_convert" function. The difference between these functions are that 
-the "watershed" function from the EBImage package comes first in the outlined 
-function compared to the "image_convert" function.
+"*image_convert*" function. The difference between these functions are that 
+the "*watershed*" function from the ***EBImage*** package comes first in the outlined function compared to the "*image_convert*" function.
 
@@
 In addition, set aside some code space to visually 
-analyze your images results. This could include adjusting the threshold
-from the "mapped" function, the contrast from the "greyscale" function 
-(if used), or could involve making adjustments to the "image_convert" function,
+analyze image results. This could include adjusting the threshold
+from the "*mapped*" function, the contrast from the "*greyscale*" function 
+(if used), or could involve making adjustments to the "*image_convert*" function,
 such as the width and height of the rectangular window, the pixel area
 threshold (i.e., removes cells or other objects under a certain pixel area), 
-or the tolerance and 'ext' variables. Observe your initial results as indicated
+or the tolerance and 'ext' variables. Observe the initial results as indicated
 below and make adjustments where necessary:
@@
 
-Once you have made necessary adjustments, use the following for-loop to
-analyze your images in bulk, one at a time. The for-loop uses the 
-"image_convert" function to produce unique color labels on cells within
+Once end-users have made necessary adjustments, use the following for-loop to
+analyze images in bulk from a directory, yet one at a time. The for-loop uses the "*image_convert*" function to produce unique color labels on cells within
 an image and to produce a final image showcasing these labels using the 
-"countImages" function. Finally, "countCells" is used to count the number of 
+"*countImages*" function. Finally, "*countCells*" is used to count the number of 
 unique color labels (i.e., cells). We also have integrated code to add these 
@@
 Data from the image analysis has been generated in our data frame "Cell.Count"
-that we previously made. We can use the following code to view our results:
+that we previously made. We can use the following code to view the current results:
 
@@
 outliers from the produced data, within a range of two standard deviations 
-above and below the median. The "outlier" function tags which images are
-out of this range and dictates a warning within our data frame, which is shown 
-below. This alerts the end-user to perhaps re-image a particular sample, or
-to investigate the error further.
+above and below the median. The "*outlier*" function tags which images are
+out of this range and dictates a warning within our data frame, "Cell.Count", which is shown below.
+This function's primary objective is to alert end-users of potential imaging/processing issues.
+This alerts the end-user to perhaps re-image a particular sample, or to investigate the error further, if any.
 
@@
 In order to calculate cell density of these images, we can use the
-"cell.density" function that's built into the *cellcount* package. We first
+"*cell.density*" function that's built into the ***cellcount*** package. We first
 need to generate the total amount of cells counted from these images and
@@
 - Total number of images analyzed
-- Filtration area (mm^2)
+- Filtration area (mm<sup>2</sup>)
 - Volume filtered per image (mL)
@@
 Image height and width are variables that are taken from dimensions 
-disseminated by your microscope camera's manufacturer. Please refer to their 
-documentation to calculate your FOV variable correctly.
+disseminated by the microscope camera's manufacturer. End-users must refer to 
+their camera documentation to calculate the FOV variable correctly.
 
-With the variables calculated, the following "cell_density" uses these 
+With the variables calculated, the following "*cell_density*" uses these 
 variables to calculate cell density and adds the data into the "Cell.Count" 

### 2022-12-02 stats_code/RA_Code.R

@@ -13,9 +13,17 @@ RA_Anabena<-subset(RA,Species=="Anabena")
 RA_Shewanella<-subset(RA,Species=="Shewanella")
 RA_E.coli<-subset(RA,Species=="E.coli")
 
+RA_F192[ nrow(RA_F192) + 1:6,] <- NA
+RA_F192$bot<-c(0,0,0,800000,800000,800000,1600000,1600000,1600000,2400000,2400000,2400000,3200000,3200000,3200000,4000000,4000000,4000000,4800000,4800000,4800000)
+RA_F192$top<-c(0,0,0,1200000,1200000,1200000,2400000,2400000,2400000,3600000,3600000,3600000,4800000,4800000,4800000,6000000,6000000,6000000,7200000,7200000,7200000)
 RA_F192$Script<-(RA_F192$Script/1000000)
 RA_F192$Visual<-(RA_F192$Visual/1000000)
 RA_F192$Microscope<-(RA_F192$Microscope/1000000)
+RA_F192$bot<-(RA_F192$bot/1000000)
+RA_F192$top<-(RA_F192$top/1000000)
+RA_F192$func1<-(sapply(RA_F192$bot, FUN = function(x){(x)}))
+RA_F192$func2<-(sapply(RA_F192$top, FUN = function(x){(x)}))
+RA_F192$num<-c(0,0,0,1,1,1,2,2,2,3,3,3,4,4,4,5,5,5,6,6,6)
 
 data.lm1<-lm(Visual~Script,data=RA_F192)
 summary(data.lm1)
@@ -210,9 +218,11 @@ F192_M<-ggplot(data=RA_F192, aes(x=Script,y=Microscope))+
         panel.border = element_rect(size=2))+
   scale_x_continuous(labels=function(x) format(x,scientific=FALSE))+
   scale_y_continuous(labels=function(x) format(x,scientific=FALSE))+
-  xlim(0,5)+
-  ylim(0,5)+
-  geom_abline(slope=1,intercept=0,color="grey40",size=1.5,alpha=0.4)+
+  geom_ribbon(aes(x = num,
+                  ymin = func1,
+                  ymax = func2),
+              fill="grey80")+
+  geom_abline(slope=1,intercept=0,size=1.5,alpha=0.4)+
   geom_abline(slope=coef(data.lm6)[["Script"]],
               intercept=coef(data.lm6)[["(Intercept)"]],
               linetype=2,

### 2022-12-15 cellcount_intro.Rmd
Updates
Updated

@@
 ---
-title: "Introduction to *cellcount*"
+title: "Introduction to *cellcount* - Draft vignette"
 author: "*T.E. Harman, M. Vandersea, D.E. Berthold, H.D. Laughinghouse, B. Boyd, K. Pokrzywinski*"
-date: "__/__/2022"
+date: "12/15/2022"
 abstract: >
@@
 
+write.csv(Cell.Count, paste0(savdir, "/Anabena_test counts.csv"))
+
 Cell.Count
@@
 
+write.csv(Cell.Count, paste0(savdir, "/Shewanella_test counts.csv"))
+
 Cell.Count
@@
 # Literature cited:
+
+Wetzel, RG and Likens, GE. 1991. Limnological Analyses, Second Edition. Springer-
+Verlag. 391 pp.
+
+Gregoire, P, Fuchs, F, Sklyar, O, Boutros, M, and Huber, W.
+2010. EBImage - an R package for image processing with applications to cellular
+phenotypes. Bioinformatics, 26(7), pp. 979-981, 10.1093/bioinformatics/btq046
+
+Pokrzywinski, K, Boyd, B, Smith, J. 2019. A High-throughput Method for Counting 
+Cyanobacteria Using Fluorescence Microscopy. ERDC TR-19-21.

### 2022-12-15 sandbox_analysis_test.R

@@ -11,14 +11,14 @@ gc()
 
 Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
 
-file_name<-paste0("D0_Treatment_") #Change the name of species analyzed here
+file_name<-paste0("D6_Treatment_") #Change the name of species analyzed here
 
 #Change directories here
-savdir <- ("C:/Users/Tyler.Harman/Desktop/DinoSHIELD/Total_Bacteria/CSV_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/DinoSHIELD/Total_Bacteria/Convert_Images/Day_0/Treatment/4_mesoJ/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/DinoSHIELD/Total_Bacteria/Quantitative_Images/Day_0/Treatment/4_mesoJ/"
+savdir <- ("C:/Users/Tyler.Harman/Desktop/DinoSHIELD/Total_Bacteria/CSV_data/Day_6/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/DinoSHIELD/Total_Bacteria/Convert_Images/Day_6/Treatment/4_mesoJ/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/DinoSHIELD/Total_Bacteria/Quantitative_Images/Day_6/Treatment/4_mesoJ/"
                      , pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/DinoSHIELD/Total_Bacteria/Quantitative_Images/Day_0/Treatment/4_mesoJ/"
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/DinoSHIELD/Total_Bacteria/Quantitative_Images/Day_6/Treatment/4_mesoJ/"
                            , pattern = "tif", full.name = F)
 
 imgNames <- paste0(file_name, images_names)
@@ -58,7 +58,7 @@ outlier<-function(){
     ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
   return(SD_range_detection)
 }
-Cell.Count$SD_range<-outlier()
+#Cell.Count$SD_range<-outlier()
 
 cell.total <- as.numeric(Cell.Count$Cell_Count)
 cell.total <- sum(cell.total)
@@ -66,15 +66,15 @@ cell.total <- sum(cell.total)
[…]

### 2023-02-16 draft_manuscript.Rmd
Manuscript file - initial upload

---
title: "cellcount: an open-source and automated approach to cyanobacteria enumeration"
description: |
  blank text.
author:
  - name: Tyler E. Harman 
    affiliation: Consolidated Safety Services, Fairfax, VA 22031, USA
    affiliation_email: tyler.harman@noaa.gov
    orcid_id: 0000-0001-5007-8170
  - name: Mark W. Vandersea
    affiliation: National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
  - name: David E. Berthold
    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
  - name: H. Dail Laughinghouse
    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
    orcid_id: 0000-0003-1018-6948
  - name: Brandon Boyd
    affiliation: U.S. Army Corps of Engineers, Engineer Research and Development Center, Vicksburg, MS 39180, USA
  - name: Kaytee Pokrzywinski
    affiliation: National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.

### 2023-02-23 draft_manuscript.Rmd
Initial upload

@@
 ---
-title: "cellcount: an open-source and automated approach to cyanobacteria enumeration"
-description: |
-  blank text.
+title: "*cellcount*: an open-source and automated approach to cyanobacteria enumeration"
 author:
@@
     affiliation: Consolidated Safety Services, Fairfax, VA 22031, USA
+      National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
     affiliation_email: tyler.harman@noaa.gov
     orcid_id: 0000-0001-5007-8170
+
   - name: Mark W. Vandersea
     affiliation: National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
+
   - name: David E. Berthold
     affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
+
   - name: H. Dail Laughinghouse
@@
     orcid_id: 0000-0003-1018-6948
+
   - name: Brandon Boyd
     affiliation: U.S. Army Corps of Engineers, Engineer Research and Development Center, Vicksburg, MS 39180, USA
+
   - name: Kaytee Pokrzywinski
@@
 ---
+# Summary
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+The ***cellcount*** package is a continuation of the report outlined in [@Pokrzywinski2019] and simplfies the quantification of bacteria and cyanobacteria *via* flourescence microscopy, using the computing language [R](https://www.r-project.org/).
+Functions within this package deconstruct images to assess individual cells, generate cell counts, and assess overall cell density from samples.
+In addition, additional tools are presented in this package using [R Shiny](https://cran.r-project.org/package=shiny) [@Chang2022] interfaces to demonstrate ease-of-use for researchers not familiar with the R environment.
+All materials outlined in this article, along with tutorials, are found in [GitHub.](https://github.com/teharman/cellcount_main)
+
+# Statement of need
+
+Harmful algal blooms (HABs) can exist within freshwater, estuarine, and marine environments. These can occur naturally within the 
+environment and are stabilized in terms of population by means of nutrient availability, temperature, light, and competition with other
+phytoplankton [@Catherine2013].
+However, in recent years, HABs have become increasingly common throughout the world, where climate change and anthropogenic impacts exacerbate the commonality of HABs.
+These blooms can cause harm to the aquatic environment, reducing phytoplankton assemblages, causing anoxic environments, and toxin production impacting higher trophic level organisms [@Kurmayer2003], [@Stumpf2012], [@Catherine2013], [@Borges2019], and [@Wood2020].
+From a public health aspect, the production of toxins can cause health issues or even death, but also cause impact from an economic standpoint such as reductions in property value and tourism in coastal areas [@Carey2012].
+
+Several methods have been established to understand HAB dynamics, primarily to develop monitoring protocols and establish mitigation 
+practices.
+Microscopic enumeration has been considered as the gold standard for several decades as the method is widely available and simple to use [@Jin2018].
+Although it is the standard method, microscopic enumeration can be time consuming, can be limited to small sample quantities, and requires significant knowledge of species identification [@Li2014] and [@Jin2018].
+Technology has become significant in advancing scientific methods, and it is no less significant in its role in advancing enumeration processes.
+Several commercially available products such as the FlowCam (Fluid Imaging Technologies Inc.) and Imaging FlowCytobot (McLane Research Laboratories Inc.) can be used to simplify enumeration methods (REF).
+These technologies can be accurate, demonstrate time-savings, and simplify data collection (REF).
+However, several limitations can be observed such as particle size, misidentification of cells, and overall grouping of particles within samples, which can require assistance from researchers (REF).
+In addition, these technologies are expensive to purchase and maintain, reducing the nature of these technologies being widely available [@Stauffer2019].
+A few studies have explored the process of automated cell enumeration methods by means of technology and software development, but promoting widespread availability.
+ImageJ is an open-source imaging analysis software and has available software for particle/cell enumeration [@Schulze2011].
+Other software algorithms have been developed to accomplish automated cellular enumeration [@Jin2018].
+These methods can still be considered time-consuming (ImageJ) and require extensive knowledge of mathematical modeling [@Jin2018].
+
+The methods mentioned previously can be viable options for researchers regarding monitoring purposes, but can be expensive and require 
+extensive training and materials.
+This leads to the suggestion of the need for accurate, low-cost, and widely available methods for rapid and simplistic automated enumeration of cells.
+A report released by [@Pokrzywinski2019] outlines an automated cyanobacteria enumeration script using the open-source programming language, R.
+[@Pokrzywinski2019] shows that while using fluorescent microscopy images, this enumeration script performs considerably well compared to standard microscopic enumeration methods.
+However, the previous presentation of this script can be considered extremely difficult to grasp for inexperienced R users.
+We utilized this script as a template for the development of an R package, involving the reduction of code into simple functions, expanding the ability for end-user modifications, development of a vignette for inexperienced users, and promoting open-source availability.
+
+# Overview of functions
+
+The ***cellcount*** package is an open-source tool that can be accessed by researchers for use in the computing language R.
+Readers should note that the Bioconductor package ***EBImage*** [@Pau2010], ***tiff*** [@Urbanek2022], ***raster*** (REF), and ***pixmap*** [@Bivand2021] packages are needed for use of this toolset for accurate data collection.
+
+The ***cellcount*** package can be installed locally and used within the RStudio environment.
+The ***cellcount*** functions are to be used in sequential order, as outlined in the *sandbox_analysis_test.R* script embedded within the ***cellcount*** package.
+Once image files and save directories have been identified, users can begin to upload their images using ***tiff*** package functions (see *readTIFF*).
+From here, images within the R environment must be converted into formats that R can use to process images.
+The *greyscale* function converts 24-bit images (RGB) into 8-bit images, and contains a built-in numerical contrast adjustment.
+
+```{r,include=FALSE}
+library(EBImage)
+library(tiff)
+library(pixmap)
+library(raster)
+library(beepr)
+library(knitr)
+Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+file_name<-paste0("F192_")
+imgNames <- paste0(file_name, images_names)
+read_images <- lapply(images, readTIFF)
+img_transposed <- lapply(read_images,aperm,c(2,1,3))
+names(images) <- imgNames
 ```
 
-Distill is a publication format for scientific and technical writing, native to the web.
+```{r,message=FALSE}
+greyscale <- function(x, contrast = 2) {
+  x <- contrast * x
+  x <- x[, , 1] + x[, , 2] + x[, , 3]
+  x <- x / max(x)
+  x <- normalize(x, inputRange = c(0.1, 0.75))
+  return(x)
+}
+```
+```{r,echo=FALSE,fig.cap="\\label Converted greyscale image of *Microcystis* sp."}
+grey_images <- lapply(img_transposed, greyscale, contrast = 1)
+display(grey_images[[5]])
+```
 
-Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.
+The *mapped* function is used to turn 8-bit images into a data matrix from the dimensions of the initial image, using the "*as.matrix*" function (see the ***data.table*** package), and includes a pixel intensity threshold adjustment.
+This removes any pixel that is under the threshold value to easilyseparate the background from cells.
+
+```{r, message=FALSE}
+mapped <- function(x, threshold = 0.3) {
+  x <- as.matrix(x)
+  x[x < threshold] <- 0
+  return(x)
+}
+```
+```{r,include=FALSE}
+imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
+```
+
+The *border* function is built into several other functions within ***cellcount*** (i.e.
+*countImages*, *image_convert*, etc.), where its primarily goal is to predict cells along a predetermined border and to remove these cells from the final cell count estimate from each image.
+(WHY IS THAT IMPORTANT-OUTLINE IT HERE).
+
+```{r, message=FALSE}
+border <- function(x, y) {
+  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
+}
+```
+
+The *image_convert* functions take the data matrices corresponding to each image and counts individual cells within an image, omitting the cells indentified from the *border* function.
+Fluorescence microscopy images analyzed by this function are of a single fluorescent dimension (i.e.
+auto-fluorescence, DAPI, etc.).
+End-users should note that this function is useful only for analyzing images containing single-cell cyanobacteria.
+*fillHull* is placed first in this function before *watershed* in order to disseminate single-cells with more accuracy.
+A parallel function in ***cellcount***, *image_convert2*, is similarly organized but meant for filamentous or colonial-type cyanobacteria.
+The difference with this is largely where *watershed* is placed before *fillHull*, opposite of *image_convert*, which improves accuracy.
+Within these two functions, several variables can be adjusted by end-users, such as the width and height of the rectangular window (see ***EBImage***), the pixel area threshold (i.e., removes cells or other objects under a certain pixel area), and the tolerance and ‘ext’ variables (see ***EBImage***).
+
+```{r}
+image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- fillHull(image)
+  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+```
+```{r}
+image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
+  image2 <- fillHull(image1)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+```
+
+Once the conversion occurs, the *countImages* function generates a color-label image, where each separated cell is given a unique color-label.
+This function includes removing cells along a predetermined border using the *border* function in order to improve count accuracy.
+End-users may view these color-label images using the *display* function (see the ***arm*** package).
+
+```{r}
+countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+  if (removeEdgeCells) {
+    dims <- dim(image)
+    border1 <- border(image, dims)
+    ids <- unique(border1[which(border1 != 0)])
+    inner <- rmObjects(image, ids)
+    EBImage::colorLabels(inner, normalize = normalize)
+  } else {
+    EBImage::colorLabels(image, normalize = normalize)
+  }
+}
+```
+```{r,echo=FALSE,fig.cap="A converted color-label image of *Microcystis* sp."}
+imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
+final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+display(final_img)
+```
+
+The final function, *countCells*, counts the number of cells resulting from the image generated by the *countImages* function.
+Once again, the *border* function is integrated here to omit counting cells within its predetermined border.
+Other items within the *sandbox_analysis_test.R* script enables users to analyze images in bulk, generate tables of results, calculate cell densities using equations from [@Wetzel2000], and export data to .csv files for easy organization.
+
+```{r}
+countCells <- function(x) {
+  countAll <- max(x)
+  dims <- dim(x)
+  border1 <- border(x, dims)
+  ids <- unique(border1[which(border1 != 0)])
+  countInner <- countAll - length(ids)
+  return(countInner)
+}
+```
+
+```{r,echo=FALSE,fig.pos="h"}
+for (z in 1:length(images)) {
+  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
+  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+  count <- countCells(imagesConverted)
+  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
+}
+cell.total <- as.numeric(Cell.Count$Cell_Count)
+cell.total <- sum(cell.total)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c(" ", " ")
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total)
+cell.av<-(cell.total/10) #change this via total number of images
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av)
+cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
+  MCF <- filtration.area / FOV
+  N <- x / images
+  cyano <- (MCF * N) / volume
+  cyano_total <- cyano * total.volume
+  return(cyano_total)
+}
+cell.den <- cell_density(cell.total, FOV = 0.078125, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5)
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den)
+cell.mL<-(cell.den/5) #change this via total volume
+Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cells per mL",cell.mL)
+kable(Cell.Count, caption = "Analyzed results from test images of *Microcystis* sp.")
+```
+
+# Graphical user interface using Shiny apps
+
+Informal discussions with researchers ignited the development of a graphical user interface (GUI) using Shiny apps, paired with functions from the ***cellcount*** package.
+This results in the use of two functions, *input_data* and *analyze images*.
+These require the ***shiny*** [@Chang2022], ***shinyalert*** [@1Attali2021], ***shinyFiles*** [@Pedersen2022], ***shinyjs*** [@2Attali2021], and ***beepr*** [@Baath2018] libraries.
+
+*input_data* uses a Shiny app UI for end-users to easily fill out the needed information in order to correctly label analyzed images, provide numerical inputs for variables, and easily select directories containing images and saving data.
+
+---
+references:
+  - type: article-journal
+    id: Catherine2013
+    author:
+      - family: Catherine
+        given: Q.
+      - family: Susanna
+        given: W.
+      - family: Isidora
+        given: E.S.
+      - family: Mark
+        given: H.
+      - family: Aurelie
+        given: V.
+      - family: Jean-Francois
+        given: H.
+    issued:
+      date-parts:
+      - - 2013
+        - 10
+        - 1
+    title: 'A review of current knowledge on toxic benthic freshwater cyanobacteria - Ecology, toxin production and risk management'
+    container-title: Water Research
+    volume: 47
+    issue: 15
+    page: 5464-5479
+    DOI: 10.1016/j.watres.2013.06.042
+    URL: https://www.sciencedirect.com/science/article/pii/S0043135413005332
+
+  - type: article-journal
+    id: Carey2012
+    author:
+      - family: Carey
+        given: C.C.
+      - family: Ibelings
+        given: B.W.
+      - family: Hoffmann
+        given: E.P.
+      - family: Hamilton
+        given: D.P.
+      - family: Brookes
+        given: J.D.
+    issued:
+      date-parts:
+      - - 2012
+        - 4
+        - 1
+    title: 'Eco-physiological adaptations that favour freshwater cyanobacteria in a changing climate'
+    container-title: Water Research
+    volume: 46
+    issue: 5
+    page: 1394-1407
+    DOI: 10.1016/j.watres.2011.12.016
+    URL: https://www.sciencedirect.com/science/article/pii/S0043135411007883
+
+  - type: article-journal
+    id: Kurmayer2003
+    author:
+      - family: Kurmayer
+        given: R.
+      - family: Christiansen
+        given: G.
+      - family: Chorus
+        given: I.
+    issued:
+      date-parts:
+      - - 2003
+        - 2
+    title: 'The Abundance of Microcystin-Producing Genotypes Correlates Positively with Colony Size in Microcystis sp. and 
+      Determines Its Microcystin Net Production in Lake Wannsee'
+    container-title: Applied and Evnrionmental Microbiology
+    volume: 69
+    issue: 2
+    page: 787-795
+    DOI: 10.1128/AEM.69.2.787-795.2003
+    URL: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC143648/
+
+  - type: article-journal
+    id: Stumpf2012
+    author:
+      - family: Stumpf
+        given: R.P.
+      - family: Wynne
+        given: T.T.
+      - family: Baker
+        given: D.B.
+      - family: Fahnenstiel
+        given: G.L.
+    issued:
+      date-parts:
+      - - 2012
+        - 8
+        - 1
+    title: 'Interannual Variability of Cyanobacterial Blooms in Lake Erie'
+    container-title: PLoS One
+    volume: 7
+    issue: 8
+    page: e42444
+    DOI: 10.1371/journal.pone.0042444
+    URL: https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0042444
+
+  - type: article-journal
+    id: Borges2019
+    author:
+      - family: Borges
+        given: H.D.
+      - family: Cicerelli
+        given: R.E.
+      - family: de Almeida
+        given: T.
+      - family: Roig
+        given: H.L.
+      - family: Olivetti
+        given: D.
+    issued:
+      date-parts:
+      - - 2019
+        - 5
+        - 7
+    title: 'Monitoring cyanobacteria occurrence in freshwater reservoirs using semi-analytical algorithms and orbital remote sensing'
+    container-title: Marine and Freshwater Research
+    volume: 71
+    issue: 5
+    page: 569-578
+    DOI: 10.1071/MF18377
+    URL: https://www.publish.csiro.au/MF/MF18377
+
+  - type: article-journal
+    id: Wood2020
+    author:
+      - family: Wood
+        given: S.A.
+      - family: Kelly
+        given: L.
+      - family: Bouma-Gregson
+        given: K.
+      - family: Humbert
+        given: J.F.
+      - family: Laughinghouse
+        given: H.D.
+      - family: Lazorchak
+        given: J.
+      - family: McAllister
+        given: T.       
+      - family: McQueen
+        given: A.
+      - family: Pokrzywinski
+        given: K.
+      - family: Puddick
+        given: J.
+      - family: Quiblier
+        given: C.
+      - family: Reitz
+        given: L.A.
+      - family: Ryan
+        given: K.
+      - family: Vadeboncoeur
+        given: Y.
+      - family: Zastepa
+        given: A.
+      - family: Davis
+        given: T.W.
+    issued:
+      date-parts:
+      - - 2020
+        - 10
+        - 1
+    title: 'Toxic benthic freshwater cyanobacterial proliferations: Challenges and solutions for enhancing knowledge and improving 
+      monitoring and mitigation'
+    container-title: Freshwater Biology
+    volume: 65
+    issue: 10
+    page: 1824-1842
+    DOI: 10.1111/fwb.13532
+    URL: https://onlinelibrary.wiley.com/doi/10.1111/fwb.13532
+
+  - type: article-journal
+    id: Jin2018
+    author:
+      - family: Jin
+        given: C.
+      - family: Mesquita
+        given: M.M.F.
+      - family: Deglint
+        given: J.L.
+      - family: Emelko
+        given: M.B.
+      - family: Wong
+        given: A.
+    issued:
+      date-parts:
+      - - 2018
+        - 6
+        - 13
+    title: 'Quantification of cyanobacterial cells via a novel imaging-driven technique with an integrated fluorescence signature'
+    container-title: Scientific Reports
+    volume: 8
+    DOI: 10.1038/s41598-018-27406-0
+    URL: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5998128/
+
+  - type: article-journal
+    id: Li2014
+    author:
+      - family: Li
+        given: M.
+      - family: Zhu
+        given: W.
+      - family: Gao
+        given: L.
+    issued:
+      date-parts:
+      - - 2014
+        - 2
+        - 26
+    title: 'Analysis of Cell Concentration, Volume Concentration, and Colony Size of Microcystis Via Laser Particle Analyzer'
+    container-title: Environmental Management
+    volume: 53
+    page: 947-958
+    DOI: 10.1007/s00267-014-0252-8
+    URL: https://link.springer.com/article/10.1007/s00267-014-0252-8
+
+  - type: article-journal
+    id: Stauffer2019
+    author:
+      - family: Stauffer
+        given: B.A.
+      - family: Bowers
+        given: H.A.
+      - family: Buckley
+        given: E.
+      - family: Davis
+        given: T.W.
+      - family: Johengen
+        given: T.H.
+      - family: Kudela
+        given: R.
+      - family: McManus
+        given: M.A. 
+      - family: Purcell
+        given: H.
+      - family: Smith
+        given: G.J.
+      - family: Vander Woude
+        given: A.
+      - family: Tamburri
+        given: M.N.
+    issued:
+      date-parts:
+      - - 2019
+        - 7
+        - 16
+    title: 'Considerations in Harmful Algal Bloom Research and Monitoring: Perspectives From a Consensus-Building Workshop and Technology 
+      Testing'
+    container-title: Frontiers in Marine Science
+    volume: 6
+    issue: 399
+    DOI: 10.3389/fmars.2019.00399
+    URL: https://www.frontiersin.org/articles/10.3389/fmars.2019.00399/full
+
+  - type: technical-report
+    id: Pokrzywinski2019
+    author:
+      - family: Pokrzywinski
+        given: K.
+      - family: Boyd
+        given: B.M.
+      - family: Smith
+        given: S.J.
+    issued:
+      date-parts:
+      - - 2019
+        - 8
+    title: 'A high-throughput method for counting cyanobacteria using fluorescence microscopy'
+    DOI: 10.21079/11681/33870
+    URL: https://erdc-library.erdc.dren.mil/jspui/handle/11681/33870
+
+  - type: article-journal
+    id: Schulze2011
+    author:
+      - family: Schulze
+        given: K.
+      - family: Lopez
+        given: D.A.
+      - family: Tillich
+        given: U.M.
+      - family: Frohme
+        given: M.
+    issued:
+      date-parts:
+      - - 2011
+        - 11
+        - 30
+    title: 'A simple viability analysis for unicellular cyanobacteria using a new autofluorescence assay, automated microscopy, and ImageJ'
+    container-title: BMC Biotechnology
+    volume: 11
+    issue: 118
+    DOI: 10.1186/1472-6750-11-118
+    URL: https://link.springer.com/article/10.1186/1472-6750-11-118
+
+  - type: article-journal
+    id: Pau2010
+    author:
+      - family: Pau
+        given: G.
+      - family: Fuchs
+        given: F.
+      - family: Skylar
+        given: O.
+      - family: Boutros
+        given: M.
+      - family: Huber
+        given: W.
+    issued:
+      date-parts:
+      - - 2010
+        - 3
+        - 23
+    title: 'EBImage—an R package for image processing with applications to cellular phenotypes'
+    container-title: Bioinformatics
+    volume: 26
+    issue: 7
+    page: 979-981
+    DOI: 10.1093/bioinformatics/btq046
+    URL: https://academic.oup.com/bioinformatics/article/26/7/979/211835
+
+  - type: package
+    id: Urbanek2022
+    author:
+      - family: Urbanek
+        given: S.
+      - family: Johnson
+        given: K.
+    issued:
+      date-parts:
+      - - 2022
+        - 1
+        - 31
+    title: 'tiff: Read and Write TIFF Images. R package version 0.1-11.'
+    URL: https://www.rforge.net/tiff/
+
+  - type: package
+    id: Bivand2021
+    author:
+      - family: Bivand
+        given: R.
+      - family: Leisch
+        given: F.
+      - family: Maechler
+        given: Martin
+    issued:
+      date-parts:
+      - - 2021
+        - 1
+        - 29
+    title: 'Bitmap Images/Pixel Maps, R Package Version 0.4.12.'
+    URL: https://cran.r-project.org/package=pixmap
+
+  - type: package
+    id: Chang2022
+    author:
+      - family: Chang
+        given: W.
+      - family: Cheng
+        given: J.
+      - family: Allaire
+        given: J.
+      - family: Boutros
+        given: M.
+      - family: Huber
+        given: W.
+    issued:
+      date-parts:
+      - - 2022
+        - 12
+        - 15
+    title: 'Shiny: Web application framework for R'
+    URL: https://cran.r-project.org/web/packages/shiny/index.html
+
+  - type: book
+    id: Wetzel2000
+    author:
+      - family: Wetzel
+        given: R.G.
+      - family: Likens
+        given: G.E.
+    issued:
+      date-parts:
+      - - 2000
+        - 3
+        - 3
+    title: 'Limnological Analyses - 3rd Edition'
+    page: 271-274
+    DOI: 10.1007/978-1-4757-3250-4
+    URL: https://link.springer.com/book/10.1007/978-1-4757-3250-4
+
+  - type: package
+    id: Pedersen2022
+    author:
+      - family: Pedersen
+        given: T.L.
+      - family: Nijs
+        given: V.
+      - family: Schaffner
+        given: T.
+      - family: Nantz
+        given: E.
+    issued:
+      date-parts:
+      - - 2022
+        - 8
+        - 19
+    title: 'shinyFiles: A Server-Side File System Viewer for Shiny'
+    URL: https://cran.r-project.org/web/packages/shinyFiles/index.html
+
+  - type: package
+    id: 1Attali2021
+    author:
+      - family: Attali
+        given: D.
+      - family: Edwards
+        given: T.
+      - family: Wang
+        given: Z.
+    issued:
+      date-parts:
+      - - 2021
+        - 12
+        - 20
+    title: 'shinyalert: Easily Create Pretty Popup Messages (Modals) in Shiny'
+    URL: https://cran.r-project.org/web/packages/shinyalert/index.html
+
+  - type: package
+    id: 2Attali2021
+    author:
+      - family: Attali
+        given: D.
+    issued:
+      date-parts:
+      - - 2021
+        - 12
+        - 23
+    title: 'shinyjs: Easily Improve the User Experience of Your Shiny Apps in Seconds'
+    URL: https://cran.r-project.org/web/packages/shinyjs/index.html
+
+  - type: package
+    id: Baath2018
+    author:
+      - family: Baath
+        given: R.
+      - family: Dobbyn
+        given: A.
+    issued:
+      date-parts:
+      - - 2018
+        - 6
+        - 4
+    title: 'beepr: Easily Play Notification Sounds on any Platform'
+    URL: https://cran.r-project.org/web/packages/beepr/index.html
+...

### 2023-02-24 R/analyze_images.R
Initial upload from GUI project files

@@ -1,0 +1,37 @@
+#' Analyze Images
+#'
+#'
+
+analyze_images<-function(){
+  library(shiny)
+  library(shinyalert)
+  library(shinyFiles)
+  library(shinyjs)
+  
+  jscode <- "shinyjs.closeWindow = function() { window.close(); }"
+  
+  ui = fluidPage(
+    useShinyjs(),
+    h1("  "),
+    fluidRow(
+      column(4, offset=2,actionButton("buttonId", "Run Image Analysis",class = "btn-success",style='height:130px;width:375px;font-size:120%',icon=icon("wand-magic-sparkles")))),
+    fluidRow(
+      column(4, offset=2,actionButton("close", "Close window",class = "btn-danger",style='height:130px;width:375px;font-size:120%',icon=icon("check"))))
+  )
+  
+  
+  server = function(input, output, session) {
+    observeEvent(input$buttonId, {
+      message("running analyze_img.R")
+      source("analyze_img.R")
+    })
+    output$table <- renderDataTable(Cell.Count)
+    observeEvent(input$close, {
+      js$closeWindow()
+      stopApp()
+    })
+  }
+  runGadget(ui, server, viewer = dialogViewer("cellcount Image Analysis Interface",
+                                              width = 400, height = 100))
+  
+}

### 2023-02-24 R/input_data.R

@@ -1,0 +1,314 @@
+input_data<-function(){
+  library(shiny)
+  library(shinyalert)
+  library(shinyFiles)
+  library(shinyjs)
+  
+  jscode <- "shinyjs.closeWindow = function() { window.close(); }"
+  
+  image_names1<<-textConnection('image_names2','wr',local=FALSE)
+  thresh_adj1<<-textConnection('thresh_adj2','wr',local=FALSE)
+  FOV1<<-textConnection('FOV2','wr',local=FALSE)
+  image_num1<<-textConnection('image_num2','wr',local=FALSE)
+  filter_vol1<<-textConnection('filter_vol2','wr',local=FALSE)
+  total_vol1<<-textConnection('total_vol2','wr',local=FALSE)
+  contrast_adj1<<-textConnection('contrast_adj2','wr',local=FALSE)
+  filter_area1<<-textConnection('filter_area2','wr',local=FALSE)
+  csv_file1<<-textConnection('csv_file2','wr',local=FALSE)
+  tolerance1<<-textConnection('tolerance2','wr',local=FALSE)
+  ext1<<-textConnection('ext2','wr',local=FALSE)
+  areathresh1<<-textConnection('areathresh2','wr',local=FALSE)
+  
+  ui1 = fixedPage(
+    useShinyjs(),
+    extendShinyjs(text = jscode, functions = c("closeWindow")),
+    div(navlistPanel(
+      "Cellcount - Image Analysis UI",
+      " ",
+      tabPanel("Directories",
+               h3("   "),
+               fixedRow(
+                 column(2,offset=2,shinyDirButton('path1','CSV save directory','Please select a directory to save CSV files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
+               ),
+               fixedRow(
+                 column(2,offset=2,shinyDirButton('path2','Image save directory','Please select a directory to save Image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
+               ),
+               fixedRow(
+                 column(2,offset=2,shinyDirButton('path3','Image analysis directory','Please select a directory containing image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
+               )
+      ),
+      tabPanel("Analysis Type",
+               h3("   "),
+               fixedRow(
+                 column(8,offset=2,selectInput("selectData",
[…]

### 2023-02-24 analyze_img.R

@@ -1,0 +1,81 @@
+library(EBImage)
+library(tiff)
+library(pixmap)
+library(raster)
+library(beepr)
+
+border <- function(x, y) {
+  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
+}
+cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
+  MCF <- filtration.area / FOV
+  N <- x / images
+  cyano <- (MCF * N) / volume
+  cyano_total <- cyano * total.volume
+  return(cyano_total)
+}
+countCells <- function(x) {
+  countAll <- max(x)
+  dims <- dim(x)
+  border1 <- border(x, dims)
+  ids <- unique(border1[which(border1 != 0)])
+  countInner <- countAll - length(ids)
+  return(countInner)
+}
+countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+  if (removeEdgeCells) {
+    dims <- dim(image)
+    border1 <- border(image, dims)
+    ids <- unique(border1[which(border1 != 0)])
+    inner <- rmObjects(image, ids)
+    EBImage::colorLabels(inner, normalize = normalize)
+  } else {
+    EBImage::colorLabels(image, normalize = normalize)
+  }
+}
+greyscale <- function(x, contrast = 2) {
+  x <- contrast * x
+  x <- x[, , 1] + x[, , 2] + x[, , 3]
+  x <- x / max(x)
+  x <- normalize(x, inputRange = c(0.1, 0.75))
+  return(x)
+}
+image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- fillHull(image)
+  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+mapped <- function(x, threshold = 0.3) {
+  x <- as.matrix(x)
+  x[x < threshold] <- 0
+  return(x)
+}
+
+
+read_images <- lapply(images, readTIFF)
[…]

### 2023-02-24 cellcount_UI_test.R

@@ -1,0 +1,348 @@
+input_data<-function(){
+    library(shiny)
+    library(shinyalert)
+    library(shinyFiles)
+    library(shinyjs)
+    
+    jscode <- "shinyjs.closeWindow = function() { window.close(); }"
+    
+    image_names1<<-textConnection('image_names2','wr',local=FALSE)
+    thresh_adj1<<-textConnection('thresh_adj2','wr',local=FALSE)
+    FOV1<<-textConnection('FOV2','wr',local=FALSE)
+    image_num1<<-textConnection('image_num2','wr',local=FALSE)
+    filter_vol1<<-textConnection('filter_vol2','wr',local=FALSE)
+    total_vol1<<-textConnection('total_vol2','wr',local=FALSE)
+    contrast_adj1<<-textConnection('contrast_adj2','wr',local=FALSE)
+    filter_area1<<-textConnection('filter_area2','wr',local=FALSE)
+    csv_file1<<-textConnection('csv_file2','wr',local=FALSE)
+    tolerance1<<-textConnection('tolerance2','wr',local=FALSE)
+    ext1<<-textConnection('ext2','wr',local=FALSE)
+    areathresh1<<-textConnection('areathresh2','wr',local=FALSE)
+    
+    ui1 = fixedPage(
+      useShinyjs(),
+      extendShinyjs(text = jscode, functions = c("closeWindow")),
+      div(navlistPanel(
+        "Cellcount - Image Analysis UI",
+        " ",
+        tabPanel("Directories",
+                 h3("   "),
+                 fixedRow(
+                   column(2,offset=2,shinyDirButton('path1','CSV save directory','Please select a directory to save CSV files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
+                 ),
+                 fixedRow(
+                   column(2,offset=2,shinyDirButton('path2','Image save directory','Please select a directory to save Image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
+                 ),
+                 fixedRow(
+                   column(2,offset=2,shinyDirButton('path3','Image analysis directory','Please select a directory containing image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
+                 )
+        ),
+        tabPanel("Analysis Type",
+                 h3("   "),
+                 fixedRow(
+                   column(8,offset=2,selectInput("selectData",
[…]

### 2023-02-24 cellcount_GUI_intro.Rmd
Add files via upload

---
title: "cellcount - GUI format"
author: "T.E. Harman"
date: "2023-01-19"
abstract: >
  *cellcount*, a package available within the R environment, is an enumeration tool used to detect concentrations of cells within fluroescent microscopy images (see Harman et al. ____). This graphical user interface (GUI) takes from functions outlined in the *cellcount* package and applies them to a *shiny* interface for ease of use for end users, regardless of technological knowledge in R. This vignette demonstrates how to operate this package and provides examples for end-users to follow along.
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2023-02-24 draft_manuscript.Rmd
Updated.

@@
 The ***cellcount*** package is an open-source tool that can be accessed by researchers for use in the computing language R.
-Readers should note that the Bioconductor package ***EBImage*** [@Pau2010], ***tiff*** [@Urbanek2022], ***raster*** (REF), and ***pixmap*** [@Bivand2021] packages are needed for use of this toolset for accurate data collection.
+Readers should note that the Bioconductor package ***EBImage*** [@Pau2010], ***tiff*** [@Urbanek2022], ***raster*** [@Hijmans2023], and ***pixmap*** [@Bivand2021] packages are needed for use of this toolset for accurate data collection.
 
-The ***cellcount*** package can be installed locally and used within the RStudio environment.
+The ***cellcount*** package can be installed locally and used within the [RStudio](https://posit.co/download/rstudio-desktop/) environment.
 The ***cellcount*** functions are to be used in sequential order, as outlined in the *sandbox_analysis_test.R* script embedded within the ***cellcount*** package.
-Once image files and save directories have been identified, users can begin to upload their images using ***tiff*** package functions (see *readTIFF*).
+Once image files have been identified, users can begin to upload their images using ***tiff*** package functions (see *readTIFF*).
 From here, images within the R environment must be converted into formats that R can use to process images.
@@
 The *mapped* function is used to turn 8-bit images into a data matrix from the dimensions of the initial image, using the "*as.matrix*" function (see the ***data.table*** package), and includes a pixel intensity threshold adjustment.
-This removes any pixel that is under the threshold value to easilyseparate the background from cells.
+This removes any pixel that is under the threshold value to easily separate the background from cells.
 
@@
 Informal discussions with researchers ignited the development of a graphical user interface (GUI) using Shiny apps, paired with functions from the ***cellcount*** package.
-This results in the use of two functions, *input_data* and *analyze images*.
-These require the ***shiny*** [@Chang2022], ***shinyalert*** [@1Attali2021], ***shinyFiles*** [@Pedersen2022], ***shinyjs*** [@2Attali2021], and ***beepr*** [@Baath2018] libraries.
+This results in the use of two functions, *input_data* and *analyze images*, to run the entire analysis process as outlined in the *sandbox_analysis_test.R* script.
+In addition to packages required by ***cellcount***, these GUIs require the ***shiny*** [@Chang2022], ***shinyalert*** [@1Attali2021], ***shinyFiles*** [@Pedersen2022], ***shinyjs*** [@2Attali2021], and ***beepr*** [@Baath2018] libraries.
 
-*input_data* uses a Shiny app UI for end-users to easily fill out the needed information in order to correctly label analyzed images, provide numerical inputs for variables, and easily select directories containing images and saving data.
+*input_data* uses a Shiny app UI for end-users to easily fill out the needed information in order to correctly label analyzed images, provide numerical inputs for variables, select image analysis type, and easily select directories containing images and saving data.
+The UI format is a simple approach, with tabs outlined on the left-hand side of the interface for its respective purpose (i.e., directory selection, analysis type selection, etc.).
+Each tab contains interactive buttons or drop-down menus for users to enter in the respective information or make appropriate selections for their analysis needs.
+Shiny directory buttons provide users a simple and familiar interface to select appropriate folders containing images or to set up save directories.
+Alert modals (see ***shinyalert***) are used for users to interact and provide information in text format, with helpful dialog to keep users organized and informed of the information needed.
+Once all information is provided, users select the "Completion" tab and click the "submit data entries" button to complete the data input process.
+
+(INSERT IMAGES OF *input_data* SHINY APP HERE)
+
+*analyze_images* uses a simplistic Shiny UI which consists of two buttons, one to run the analysis and the other to close the UI interface.
+
+# Limitiations and warnings
+
+# Conclusions
+
+# Acknowledgements
+
+# Literature Cited
 
@@
     URL: https://cran.r-project.org/web/packages/beepr/index.html
+
+  - type: package
+    id: Hijmans2023
+    author:
+      - family: Hijmans
+        given: R.J
+      - family: van Etten
+        given: J.
+      - family: Sumner
+        given: M.
+      - family: Cheng
+        given: J.
+      - family: Baston
+        given: D.
+      - family: Bevan
+        given: A.
+      - family: Bivand
+        given: R.
+      - family: Busetto
+        given: L.
+      - family: Canty
+        given: M.
+      - family: Fasoli
+        given: B.
+      - family: Forrest
+        given: D.
+      - family: Ghosh
+        given: A.
+      - family: Golicher
+        given: D.
+      - family: Gray
+        given: J.
+      - family: Greenburg
+        given: J.A.
+      - family: Hiemstra
+        given: P.
+      - family: Hingee
+        given: K.
+      - family: Ilich
+        given: A.
+      - family: Karney
+        given: C.
+      - family: Mattiuzzi
+        given: M.
+      - family: Mosher
+        given: S.
+      - family: Naimi
+        given: B.
+      - family: Nowosad
+        given: J.
+      - family: Pebesma
+        given: E.
+      - family: Lamigueiro
+        given: O.P.
+      - family: Racine
+        given: E.B.
+      - family: Rowlingson
+        given: B.
+      - family: Shortridge
+        given: A.
+      - family: Venables
+        given: B.
+      - family: Wueest
+        given: R.
+    issued:
+      date-parts:
+      - - 2023
+        - 1
+        - 16
+    title: 'raster: Geographic Data Analysis and Modeling'
+    URL: https://cran.r-project.org/web/packages/raster/index.html
 ...

### 2023-02-24 R/input_data.R
Updated - changes to UI and additions of rectangular window width/height numerical values

@@ -18,6 +18,8 @@ input_data<-function(){
   tolerance1<<-textConnection('tolerance2','wr',local=FALSE)
   ext1<<-textConnection('ext2','wr',local=FALSE)
   areathresh1<<-textConnection('areathresh2','wr',local=FALSE)
+  rec_width1<<-textConnection('rec_width2','wr',local=FALSE)
+  rec_height1<<-textConnection('rec_height2','wr',local=FALSE)
   
   ui1 = fixedPage(
     useShinyjs(),
@@ -25,7 +27,8 @@ input_data<-function(){
     div(navlistPanel(
       "Cellcount - Image Analysis UI",
       " ",
-      tabPanel("Directories",
+      " ",
+      tabPanel(h4("Directories"),
                h3("   "),
                fixedRow(
                  column(2,offset=2,shinyDirButton('path1','CSV save directory','Please select a directory to save CSV files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
@@ -37,7 +40,7 @@ input_data<-function(){
                  column(2,offset=2,shinyDirButton('path3','Image analysis directory','Please select a directory containing image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
                )
       ),
-      tabPanel("Analysis Type",
+      tabPanel(h4("Analysis Type"),
                h3("   "),
                fixedRow(
                  column(8,offset=2,selectInput("selectData",
@@ -45,56 +48,67 @@ input_data<-function(){
                                                choices=list("Single-cell","Filamentous","Colonial"),selected="Single-cell"))
                )
       ),
-      tabPanel("Data inputs",
+      tabPanel(h4("Data inputs"),
+               h3("   "),
                fixedRow(
-                 column(2,offset=2,actionButton("run4", "Species name", class = "btn-success",style='height:30px;width:350px;font-size:120%',icon=icon("clipboard")))
+                 column(2,offset=2,actionButton("run4", "Species name", class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("clipboard")))
                ),
                fixedRow(
-                 column(2,offset=2,actionButton("run6", 'Threshold adjustment', class = "btn-success",style='height:30px;width:350px;font-size:120%',icon=icon("gear")))
[…]

### 2023-02-24 analyze_img.R

@@ -56,26 +56,52 @@ mapped <- function(x, threshold = 0.3) {
 }
 
 
-read_images <- lapply(images, readTIFF)
-img_transposed <- lapply(read_images,aperm,c(2,1,3))
-grey_images <- lapply(img_transposed, greyscale, contrast = contrast_adj2)
-imagesMapped <- lapply(grey_images, mapped, threshold = thresh_adj2)
-for (z in 1:length(images)) {
-  imagesConverted <- image_convert(imagesMapped[[z]], w = 30, h = 50, offset = 0.001, areathresh = areathresh2, tolerance = tolerance2, ext = ext2)
-  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
-  count <- countCells(imagesConverted)
-  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
-  analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
-  writeImage(final_img, files = paste0(image_file, analyzed_image),compression=c("LZW"))
+if(analysis_type == 'single-cell'){
+  read_images <- lapply(images, readTIFF)
+  img_transposed <- lapply(read_images,aperm,c(2,1,3))
+  grey_images <- lapply(img_transposed, greyscale, contrast = contrast_adj2)
+  imagesMapped <- lapply(grey_images, mapped, threshold = thresh_adj2)
+  for (z in 1:length(images)) {
+    imagesConverted <- image_convert(imagesMapped[[z]], w = rec_width2, h = rec_height2, offset = 0.001, areathresh = areathresh2, tolerance = tolerance2, ext = ext2)
+    final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+    count <- countCells(imagesConverted)
+    Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
+    analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
+    writeImage(final_img, files = paste0(image_file, analyzed_image),compression=c("LZW"))
+  }
+  cell.total <<- as.numeric(Cell.Count$Cell_Count)
+  cell.total <<- sum(cell.total)
[…]

### 2023-02-24 R/analyze_images.R
Updated.

@@ -7,6 +7,7 @@ analyze_images<-function(){
   library(shinyalert)
   library(shinyFiles)
   library(shinyjs)
+  library(lubridate)
   
   jscode <- "shinyjs.closeWindow = function() { window.close(); }"
   
@@ -32,6 +33,6 @@ analyze_images<-function(){
     })
   }
   runGadget(ui, server, viewer = dialogViewer("cellcount Image Analysis Interface",
-                                              width = 400, height = 100))
+                                              width = 400, height = 200))
   
 }

### 2023-02-24 analyze_img.R

@@ -49,6 +49,15 @@ image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, to
   image3 <- rmObjects(image2, nr)
   return(image3)
 }
+image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
+  image2 <- fillHull(image1)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
 mapped <- function(x, threshold = 0.3) {
   x <- as.matrix(x)
   x[x < threshold] <- 0

### 2023-02-24 draft_manuscript.Rmd
Updated.

@@
 Several commercially available products such as the FlowCam (Fluid Imaging Technologies Inc.) and Imaging FlowCytobot (McLane Research Laboratories Inc.) can be used to simplify enumeration methods (REF).
-These technologies can be accurate, demonstrate time-savings, and simplify data collection (REF).
-However, several limitations can be observed such as particle size, misidentification of cells, and overall grouping of particles within samples, which can require assistance from researchers (REF).
-In addition, these technologies are expensive to purchase and maintain, reducing the nature of these technologies being widely available [@Stauffer2019].
+**These technologies can be accurate, demonstrate time-savings, and simplify data collection (REF).
+However, several limitations can be observed such as particle size, misidentification of cells, and overall grouping of particles within samples, which can require assistance from researchers (REF).** In addition, these technologies are expensive to purchase and maintain, reducing the nature of these technologies being widely available [@Stauffer2019].
 A few studies have explored the process of automated cell enumeration methods by means of technology and software development, but promoting widespread availability.
-ImageJ is an open-source imaging analysis software and has available software for particle/cell enumeration [@Schulze2011].
+ImageJ is an open-source imaging analysis software and has available tools for particle/cell enumeration [@Schulze2011].
 Other software algorithms have been developed to accomplish automated cellular enumeration [@Jin2018].
@@
 [@Pokrzywinski2019] shows that while using fluorescent microscopy images, this enumeration script performs considerably well compared to standard microscopic enumeration methods.
-However, the previous presentation of this script can be considered extremely difficult to grasp for inexperienced R users.
+However, the previous presentation of this script can be considered extremely difficult to grasp for inexperienced R users and lacks organization.
 We utilized this script as a template for the development of an R package, involving the reduction of code into simple functions, expanding the ability for end-user modifications, development of a vignette for inexperienced users, and promoting open-source availability.
@@
 *analyze_images* uses a simplistic Shiny UI which consists of two buttons, one to run the analysis and the other to close the UI interface.
+Once the proper data has been entered from the *input_data* function, users simply click the "Run Image Analysis" button to begin.
+Sounds from the ***beepr*** package alert users of completion of the analysis, where the "Close window" button exits the UI window.
+Users can navigate to their specific directories to find analyzed images and related .csv files for their review.
+
+(INSERT IMAGE OF *analyze_images* SHINY APP HERE)
 
@@
 
+The current version of this package may be limited for some users regarding processing power.
+Analysis with this package uses local memory, which may be limiting if expected to analyze large amounts of images in batches.
+We recommend at least 16GB of localized RAM, the use of a cloud computing supercomputer, or organize images in easier-to-analyze batches to reduce local memory strain.
+
+Users must be aware that this package does not contain abilities to identify distinct cells.
+The development of this package utilized carefully cultured monoculture of various cyanobacteria species.
+Users may find difficulty in using this analysis tool if cultures contain heavy amounts of particulates, other cellular debris, and contamination of other species.
+It is advised to use this tool for single-species only.
+Future developments of this tool include identifying features and building image libraries of specific cellular species for use in environmental samples.
+
+Outside of package limitations, users must be aware of limitations that may arise from poor microscopy imaging; reduce lighting as much as possible, avoid heavy background fluorescence, and so forth.
+Good images produce good results - keep this in mind as you utilize this tool.
+
 # Conclusions
 
+***cellcount*** provides an open-source approach to cyanobacteria enumeration that is accurate and rapid compared to standardized microscopy methods.
+The open-source approach allows researchers to access this tool for free and to modify and make additions the existing code to fulfill their particular needs, whereas the Shiny GUI allows users of all coding backgrounds to utilize this analysis tool.
+
 # Acknowledgements
+
+We acknowledge...
 

### 2023-02-27 R/analyze_images.R
Updates made - changes to function documentation

@@ -1,6 +1,12 @@
 #' Analyze Images
 #'
-#'
+#' Brings in data entered from the "input_data" function and automatically runs a script to analyze images
+#' pertaining to cells from fluorescent microscopy images. This uses a Shiny user-interface, where two buttons
+#' are available to interact with. One button "Run Image Analysis" runs the analysis script while the "Close Window"
+#' button closes the Shiny interface.
+#' 
+#' @return Runs an automatic image analysis
+#' @export
 
 analyze_images<-function(){
   library(shiny)

### 2023-02-27 R/input_data.R

@@ -1,3 +1,13 @@
+#' Input Data
+#'
+#' This function provides a Shiny user-interface to easily input data needed for fluorescent microscopy image analysis
+#' outlined in *cellcount*. This Shiny app contains several tabs for data input, directory selection, etc. Users must close
+#' the app by selecting the 'Completion' tab followed by 'Submit data entries' for all entered data to be exported into 
+#' the local R environment.
+#' 
+#' @return Data input user-interface for *cellcount* image analysis
+#' @export
+
 input_data<-function(){
   library(shiny)
   library(shinyalert)

### 2023-02-27 EBImage_Markdown.Rmd
Updated
Updated - changes made to equations in vignette

@@
 author: "*T.E. Harman, M. Vandersea, D.E. Berthold, H.D. Laughinghouse, B. Boyd, K. Pokrzywinski*"
-date: "__/__/2022"
+date: "02/27/2023"
 abstract: >
@@
 - Width of image field of view (μm)
-- Median cell diameter (μm)
 - Total number of images analyzed
@@
 
-Field of view (FOV) is an important variable calculated from image height,
-image width, and cell diameter variables. The following equation is listed
-below:
-
-$FOV = HW - 2D(H+W)+4D^{2}$
-
-Where:
-
-$H = height$
-
-$W = width$
-
-$D = diameter$
-
-Image height and width are variables that are taken from dimensions 
+Field of view (FOV - mm<sup>2</sup>) is an important variable calculated from image height and
+image width. Image height and width are variables that are taken from dimensions 
 disseminated by the microscope camera's manufacturer. End-users must refer to 
@@
 ```
-
-#Literature cited:

### 2023-02-27 cellcount_intro.Rmd
Updates to equations and citations

@@
 author: "*T.E. Harman, M. Vandersea, D.E. Berthold, H.D. Laughinghouse, B. Boyd, K. Pokrzywinski*"
-date: "12/15/2022"
+date: "02/27/2023"
 abstract: >
@@
 - Width of image field of view (μm)
-- Median cell diameter (μm)
 - Total number of images analyzed
-- Filtration area (mm^2)
+- Filtration area (mm<sup>2</sup>)
 - Volume filtered per image (mL)
@@
 
-Field of view (FOV) is an important variable calculated from image height,
-image width, and cell diameter variables. The following equation is listed
-below:
-
-$FOV = HW - 2D(H+W)+4D^{2}$
-
-Where:
-
-$H = height$
-
-$W = width$
-
-$D = diameter$
-
-Image height and width are variables that are taken from dimensions 
+Field of view (FOV - mm<sup>2</sup>) is an important variable calculated from image height and
+image width. Image height and width are variables that are taken from dimensions 
 disseminated by your microscope camera's manufacturer. Please refer to their 
@@
 
-Gregoire, P, Fuchs, F, Sklyar, O, Boutros, M, and Huber, W.
-2010. EBImage - an R package for image processing with applications to cellular
-phenotypes. Bioinformatics, 26(7), pp. 979-981, 10.1093/bioinformatics/btq046
+Pau, G, Fuchs, F, Sklyar, O, Boutros, M, Huber, W.
+2010.
+EBImage—an R package for image processing with applications to cellular phenotypes, Bioinformatics, 26(7): 979–981, https://doi.org/10.1093/bioinformatics/btq046
 

### 2023-02-27 draft_manuscript.Rmd
Small updates to authorship

@@
     affiliation: National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
+    orcid_id: 0000-0003-1523-9930
 date: "`r Sys.Date()`"

### 2023-02-27 cellcount_UI_test.R
Removed - sandbox files no longer needed

@@ -1,348 +1,0 @@
-input_data<-function(){
-    library(shiny)
-    library(shinyalert)
-    library(shinyFiles)
-    library(shinyjs)
-    
-    jscode <- "shinyjs.closeWindow = function() { window.close(); }"
-    
-    image_names1<<-textConnection('image_names2','wr',local=FALSE)
-    thresh_adj1<<-textConnection('thresh_adj2','wr',local=FALSE)
-    FOV1<<-textConnection('FOV2','wr',local=FALSE)
-    image_num1<<-textConnection('image_num2','wr',local=FALSE)
-    filter_vol1<<-textConnection('filter_vol2','wr',local=FALSE)
-    total_vol1<<-textConnection('total_vol2','wr',local=FALSE)
-    contrast_adj1<<-textConnection('contrast_adj2','wr',local=FALSE)
-    filter_area1<<-textConnection('filter_area2','wr',local=FALSE)
-    csv_file1<<-textConnection('csv_file2','wr',local=FALSE)
-    tolerance1<<-textConnection('tolerance2','wr',local=FALSE)
-    ext1<<-textConnection('ext2','wr',local=FALSE)
-    areathresh1<<-textConnection('areathresh2','wr',local=FALSE)
-    
-    ui1 = fixedPage(
-      useShinyjs(),
-      extendShinyjs(text = jscode, functions = c("closeWindow")),
-      div(navlistPanel(
-        "Cellcount - Image Analysis UI",
-        " ",
-        tabPanel("Directories",
-                 h3("   "),
-                 fixedRow(
-                   column(2,offset=2,shinyDirButton('path1','CSV save directory','Please select a directory to save CSV files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
-                 ),
-                 fixedRow(
-                   column(2,offset=2,shinyDirButton('path2','Image save directory','Please select a directory to save Image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
-                 ),
-                 fixedRow(
-                   column(2,offset=2,shinyDirButton('path3','Image analysis directory','Please select a directory containing image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
-                 )
-        ),
-        tabPanel("Analysis Type",
-                 h3("   "),
-                 fixedRow(
-                   column(8,offset=2,selectInput("selectData",
[…]

### 2023-02-27 stats_code/BA_Code.R
Moved to new folder

@@ -1,148 +1,0 @@
-library(ggplot2)
-library(gridExtra)
-library(ggpubr)
-library(cowplot)
-
-setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data")
-
-BA<-read.csv("Boxplot_Analysis.csv",header=T)
-
-BA_F192<-subset(BA,Species=="F192")
-BA_F210<-subset(BA,Species=="F210")
-BA_Anabena<-subset(BA,Species=="Anabena")
-BA_Anabena$Concentration_f = factor(BA_Anabena$Concentration, levels=c('High','Medium','Low'))
-BA_Shewanella<-subset(BA,Species=="Shewanella")
-BA_E.coli<-subset(BA,Species=="E.coli")
-
-
-BA_F192$Output<-(BA_F192$Output/1000000)
-BA_F192$Output<-(sqrt(BA_F192$Output))
-
-F192<-ggplot(data=BA_F192, aes(x=Concentration,y=Output,fill=Method))+
-  scale_fill_grey(start=0.5,end=0.9)+
-  stat_boxplot(geom='errorbar')+
-  geom_boxplot(size=0.5)+
-  theme_bw()+
-  labs(x="Concentration")+
-  labs(y=expression(paste("Cells per mL ( ",x10^{6},") (sqrt transformed)")))+
-  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))+
-  theme(legend.text = element_text(size=13),
-        legend.title = element_text(size=13))+
-  theme(axis.title.x = element_text(size=15),
-        axis.title.y = element_text(size=13,vjust=2))+
-  theme(axis.text.x = element_text(size = 15),
-        axis.text.y = element_text(size = 15),
-        axis.ticks = element_line(size=1.5),
-        axis.ticks.length = unit(0.25,"cm"),
-        plot.margin=unit(c(1,0.25,1,1),"cm"),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.border = element_rect(size=1.5))
-F192  
-
-BA_F210$Output<-(BA_F210$Output/1000000)
-BA_F210$Output<-(sqrt(BA_F210$Output))
-
-F210<-ggplot(data=BA_F210, aes(x=Concentration,y=Output,fill=Method))+
-  scale_fill_grey(start=0.5,end=0.9)+
-  stat_boxplot(geom='errorbar')+
[…]

### 2023-02-27 stats_code/RA_IND_Code.R

@@ -1,169 +1,0 @@
-library(ggplot2)
-library(gridExtra)
-library(ggpubr)
-library(cowplot)
-
-setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data")
-
-RA<-read.csv("Regression_Analysis_IND.csv",header=T)
-
-RA_F192<-subset(RA,Species=="F192")
-RA_F210<-subset(RA,Species=="F210")
-RA_Anabena<-subset(RA,Species=="Anabena")
-RA_Shewanella<-subset(RA,Species=="Shewanella")
-RA_E.coli<-subset(RA,Species=="E.coli")
-
-data.lm1<-lm(Visual~Script,data=RA_F192)
-summary(data.lm1)
-
-F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
-  theme_bw()+
-  labs(x=expression(paste(italic("cellcount"), " results")))+
-  labs(y=expression(paste("Visual counts")))+
-  theme(axis.title.x = element_text(size=15),
-        axis.title.y = element_text(size=15,vjust=2))+
-  theme(axis.text.x = element_text(size = 15),
-        axis.text.y = element_text(size = 15),
-        axis.ticks = element_line(size=1.5),
-        axis.ticks.length = unit(0.25,"cm"),
-        plot.margin=unit(c(1,1,1,1),"cm"),
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.border = element_rect(size=2))+
-  scale_x_continuous(labels=function(x) format(x,scientific=FALSE))+
-  scale_y_continuous(labels=function(x) format(x,scientific=FALSE))+
-  xlim(0,500)+
-  ylim(0,500)+
-  geom_abline(slope=1,intercept=0,color="grey40",size=1.5,alpha=0.4)+
-  geom_smooth(method=lm,level=0.9,color=NA, fullrange=TRUE)+
-  geom_abline(slope=coef(data.lm1)[["Script"]],
-              intercept=coef(data.lm1)[["(Intercept)"]],
-              linetype=2,
-              size=1.5)+
-  geom_point(size=4,shape=21,colour="black",fill="grey60",stroke=1.5)
-F192
-
-data.lm2<-lm(Visual~Script,data=RA_F210)
-summary(data.lm2)
-
-F210<-ggplot(data=RA_F210, aes(x=Script,y=Visual))+
-  theme_bw()+
-  labs(x=expression(paste(italic("cellcount"), " results")))+
[…]

### 2023-02-27 stats_code/RA_Residuals.R

@@ -1,183 +1,0 @@
-library(ggplot2)
-library(gridExtra)
-library(ggpubr)
-
-setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/stat_code/")
-
-RA<-read.csv("Regression_Analysis_IND.csv",header=T)
-
-RA_F192<-subset(RA,Species=="F192")
-RA_F210<-subset(RA,Species=="F210")
-RA_Anabena<-subset(RA,Species=="Anabena")
-RA_Shewanella<-subset(RA,Species=="Shewanella")
-
-data.lm1<-lm(Visual~Script,data=RA_F192)
-RA_F192$predicted<-predict(data.lm1)
-RA_F192$residuals<-residuals(data.lm1)
-
-resid_F192<-ggplot(RA_F192,aes(x=Script,y=Visual))+
-  geom_smooth(method="lm",se=FALSE,color="lightgrey")+
-  geom_segment(aes(xend=Script,yend=predicted),alpha=0.2)+
-  geom_point(aes(color=abs(residuals),size=abs(residuals)))+
-  scale_color_continuous(low="green",high="red")+
-  guides(color="none",size="none")+
-  geom_point(aes(y=predicted),shape=1)+
-  theme_classic()+
-  labs(title="Microcystis F192")+
-  theme(
-    plot.title = element_text(face="bold.italic",size=17,hjust=0.5),
-    axis.text.x = element_text(face = "bold", 
-                               size = 13),
-    axis.text.y = element_text(face = "bold", 
-                               size = 13),
-    axis.title.x = element_text(size=17,vjust=-1),
-    axis.title.y = element_text(size=17,vjust=3),
-    plot.margin=unit(c(1,1,1,1),"cm")
-  )
-resid_F192
-
-data.lm2<-lm(Visual~Script,data=RA_F210)
-RA_F210$predicted<-predict(data.lm2)
-RA_F210$residuals<-residuals(data.lm2)
-
-resid_F210<-ggplot(RA_F210,aes(x=Script,y=Visual))+
-  geom_smooth(method="lm",se=FALSE,color="lightgrey")+
-  geom_segment(aes(xend=Script,yend=predicted),alpha=0.2)+
-  geom_point(aes(color=abs(residuals),size=abs(residuals)))+
-  scale_color_continuous(low="green",high="red")+
-  guides(color="none",size="none")+
-  geom_point(aes(y=predicted),shape=1)+
[…]

### 2023-02-27 manuscript_files/stats_code/BA_Code.R
New folder organization

@@ -1,0 +1,148 @@
+library(ggplot2)
+library(gridExtra)
+library(ggpubr)
+library(cowplot)
+
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data")
+
+BA<-read.csv("Boxplot_Analysis.csv",header=T)
+
+BA_F192<-subset(BA,Species=="F192")
+BA_F210<-subset(BA,Species=="F210")
+BA_Anabena<-subset(BA,Species=="Anabena")
+BA_Anabena$Concentration_f = factor(BA_Anabena$Concentration, levels=c('High','Medium','Low'))
+BA_Shewanella<-subset(BA,Species=="Shewanella")
+BA_E.coli<-subset(BA,Species=="E.coli")
+
+
+BA_F192$Output<-(BA_F192$Output/1000000)
+BA_F192$Output<-(sqrt(BA_F192$Output))
+
+F192<-ggplot(data=BA_F192, aes(x=Concentration,y=Output,fill=Method))+
+  scale_fill_grey(start=0.5,end=0.9)+
+  stat_boxplot(geom='errorbar')+
+  geom_boxplot(size=0.5)+
+  theme_bw()+
+  labs(x="Concentration")+
+  labs(y=expression(paste("Cells per mL ( ",x10^{6},") (sqrt transformed)")))+
+  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))+
+  theme(legend.text = element_text(size=13),
+        legend.title = element_text(size=13))+
+  theme(axis.title.x = element_text(size=15),
+        axis.title.y = element_text(size=13,vjust=2))+
+  theme(axis.text.x = element_text(size = 15),
+        axis.text.y = element_text(size = 15),
+        axis.ticks = element_line(size=1.5),
+        axis.ticks.length = unit(0.25,"cm"),
+        plot.margin=unit(c(1,0.25,1,1),"cm"),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.border = element_rect(size=1.5))
+F192  
+
+BA_F210$Output<-(BA_F210$Output/1000000)
+BA_F210$Output<-(sqrt(BA_F210$Output))
+
+F210<-ggplot(data=BA_F210, aes(x=Concentration,y=Output,fill=Method))+
+  scale_fill_grey(start=0.5,end=0.9)+
+  stat_boxplot(geom='errorbar')+
[…]

### 2023-02-27 manuscript_files/stats_code/RA_Code.R

@@ -1,0 +1,296 @@
+library(ggplot2)
+library(gridExtra)
+library(ggpubr)
+library(cowplot)
+
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data")
+
+RA<-read.csv("Regression_Analysis.csv",header=T)
+
+RA_F192<-subset(RA,Species=="F192")
+RA_F210<-subset(RA,Species=="F210")
+RA_Anabena<-subset(RA,Species=="Anabena")
+RA_Shewanella<-subset(RA,Species=="Shewanella")
+RA_E.coli<-subset(RA,Species=="E.coli")
+
+RA_F192[ nrow(RA_F192) + 1:6,] <- NA
+RA_F192$bot<-c(0,0,0,800000,800000,800000,1600000,1600000,1600000,2400000,2400000,2400000,3200000,3200000,3200000,4000000,4000000,4000000,4800000,4800000,4800000)
+RA_F192$top<-c(0,0,0,1200000,1200000,1200000,2400000,2400000,2400000,3600000,3600000,3600000,4800000,4800000,4800000,6000000,6000000,6000000,7200000,7200000,7200000)
+RA_F192$Script<-(RA_F192$Script/1000000)
+RA_F192$Visual<-(RA_F192$Visual/1000000)
+RA_F192$Microscope<-(RA_F192$Microscope/1000000)
+RA_F192$bot<-(RA_F192$bot/1000000)
+RA_F192$top<-(RA_F192$top/1000000)
+RA_F192$func1<-(sapply(RA_F192$bot, FUN = function(x){(x)}))
+RA_F192$func2<-(sapply(RA_F192$top, FUN = function(x){(x)}))
+RA_F192$num<-c(0,0,0,1,1,1,2,2,2,3,3,3,4,4,4,5,5,5,6,6,6)
+
+data.lm1<-lm(Visual~Script,data=RA_F192)
+summary(data.lm1)
+
+F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
+  theme_bw()+
+  labs(x=expression(paste(italic("cellcount"), " results (",x10^{6},")")))+
+  labs(y=expression(paste("Visual counts ( ",x10^{6},")")))+
+  theme(axis.title.x = element_text(size=15),
+        axis.title.y = element_text(size=15,vjust=2))+
+  theme(axis.text.x = element_text(size = 15),
+        axis.text.y = element_text(size = 15),
[…]

### 2023-02-27 manuscript_files/stats_code/RA_IND_Code.R

@@ -1,0 +1,169 @@
+library(ggplot2)
+library(gridExtra)
+library(ggpubr)
+library(cowplot)
+
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data")
+
+RA<-read.csv("Regression_Analysis_IND.csv",header=T)
+
+RA_F192<-subset(RA,Species=="F192")
+RA_F210<-subset(RA,Species=="F210")
+RA_Anabena<-subset(RA,Species=="Anabena")
+RA_Shewanella<-subset(RA,Species=="Shewanella")
+RA_E.coli<-subset(RA,Species=="E.coli")
+
+data.lm1<-lm(Visual~Script,data=RA_F192)
+summary(data.lm1)
+
+F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
+  theme_bw()+
+  labs(x=expression(paste(italic("cellcount"), " results")))+
+  labs(y=expression(paste("Visual counts")))+
+  theme(axis.title.x = element_text(size=15),
+        axis.title.y = element_text(size=15,vjust=2))+
+  theme(axis.text.x = element_text(size = 15),
+        axis.text.y = element_text(size = 15),
+        axis.ticks = element_line(size=1.5),
+        axis.ticks.length = unit(0.25,"cm"),
+        plot.margin=unit(c(1,1,1,1),"cm"),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.border = element_rect(size=2))+
+  scale_x_continuous(labels=function(x) format(x,scientific=FALSE))+
+  scale_y_continuous(labels=function(x) format(x,scientific=FALSE))+
+  xlim(0,500)+
+  ylim(0,500)+
+  geom_abline(slope=1,intercept=0,color="grey40",size=1.5,alpha=0.4)+
+  geom_smooth(method=lm,level=0.9,color=NA, fullrange=TRUE)+
+  geom_abline(slope=coef(data.lm1)[["Script"]],
+              intercept=coef(data.lm1)[["(Intercept)"]],
+              linetype=2,
+              size=1.5)+
+  geom_point(size=4,shape=21,colour="black",fill="grey60",stroke=1.5)
+F192
+
+data.lm2<-lm(Visual~Script,data=RA_F210)
+summary(data.lm2)
+
+F210<-ggplot(data=RA_F210, aes(x=Script,y=Visual))+
+  theme_bw()+
+  labs(x=expression(paste(italic("cellcount"), " results")))+
[…]

### 2023-02-27 manuscript_files/stats_code/RosnerTest.R

@@ -1,0 +1,94 @@
+library(outliers)
+library(EnvStats)
+
+setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data/")
+
+RA<-read.csv("Regression_Analysis_IND.csv",header=T)
+
+RA_F192<-subset(RA,Species=="F192")
+RA_F192A<-subset(RA_F192,Concentration=="A")
+RA_F192B<-subset(RA_F192,Concentration=="B")
+RA_F192C<-subset(RA_F192,Concentration=="C")
+RA_F192D<-subset(RA_F192,Concentration=="D")
+RA_F192E<-subset(RA_F192,Concentration=="E")
+
+rosnerTest(RA_F192A$Script)
+rosnerTest(RA_F192B$Script)
+rosnerTest(RA_F192C$Script)
+rosnerTest(RA_F192D$Script)
+rosnerTest(RA_F192E$Script)
+rosnerTest(RA_F192A$Visual)
+rosnerTest(RA_F192B$Visual)
+rosnerTest(RA_F192C$Visual)
+rosnerTest(RA_F192D$Visual)
+rosnerTest(RA_F192E$Visual)
+
+
+RA_F210<-subset(RA,Species=="F210")
+RA_F210A<-subset(RA_F210,Concentration=="A")
+RA_F210B<-subset(RA_F210,Concentration=="B")
+RA_F210C<-subset(RA_F210,Concentration=="C")
+RA_F210D<-subset(RA_F210,Concentration=="D")
+RA_F210E<-subset(RA_F210,Concentration=="E")
+
+rosnerTest(RA_F210A$Script)
+rosnerTest(RA_F210B$Script)
+rosnerTest(RA_F210C$Script)
+rosnerTest(RA_F210D$Script)
+rosnerTest(RA_F210E$Script)
+rosnerTest(RA_F210A$Visual)
+rosnerTest(RA_F210B$Visual)
+rosnerTest(RA_F210C$Visual)
+rosnerTest(RA_F210D$Visual)
+rosnerTest(RA_F210E$Visual)
+
+
+RA_Anabena<-subset(RA,Species=="Anabena")
+RA_AnabenaHigh<-subset(RA_Anabena,Concentration=="High")
+RA_AnabenaMed<-subset(RA_Anabena,Concentration=="Medium")
[…]

### 2023-02-27 cellcount_intro.Rmd
Updates to shiny-UI section

@@
 
+# Shiny app user-interface for ***cellcount*** analysis
+
+Several informal discussions with researchers have expressed interest in a easy-to-use analysis interface for inexperienced R users
+to utilize the ***cellcount*** tool. The following overview uses ***Shiny*** apps to enter input data and run image analyses.
+
+The *input_data* function generates a Shiny UI where users enter information regarding directories, numerical and categorical data, 
+and selecting analysis type. Tabs to the left of the UI separate particular categories for easy organization. Once all data has been
+provided, users complete the input of data by selecting the "Completion" tab and clicking the "Submit data entries" button to export 
+data to the local R environment.
+
+(SHOW UI IMAGES HERE)
+
+*analyze_images* is a simple UI that contains two interactive buttons.
+"Run Image Analysis" simply runs the image analysis previously outlined in this document and uses the inputted data from the previous function.
+Once the completion sound has been played, users can click the "Close window" button to exit the UI.
+Users can review the processed data in their respective directories.
+
+(SHOW UI IMAGE HERE)
+
 # Literature cited:

### 2023-03-03 manuscript_files/draft_manuscript.Rmd
Updates to files

--- 2023-02-27 draft_manuscript.Rmd
@@
 
-(INSERT IMAGES OF *input_data* SHINY APP HERE)
+![Graphical user interface of the 'input_data' function using Shiny](C:/Users/Tyler.Harman/Pictures/cellcount-UI_1.png)
 
@@
 
-(INSERT IMAGE OF *analyze_images* SHINY APP HERE)
+![Graphical user interface of the 'analyze_images' function using Shiny](C:/Users/Tyler.Harman/Pictures/cellcount-UI_2.png)
 

### 2023-03-03 manuscript_files/draft_manuscript.Rmd
Updates to authorship

@@
 
-  - name: David E. Berthold
-    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
+#  - name: David E. Berthold
+#    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
 
-  - name: H. Dail Laughinghouse
-    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
-    orcid_id: 0000-0003-1018-6948
-
-  - name: Brandon Boyd
-    affiliation: U.S. Army Corps of Engineers, Engineer Research and Development Center, Vicksburg, MS 39180, USA
+#  - name: H. Dail Laughinghouse
+#    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
+#    orcid_id: 0000-0003-1018-6948
 

### 2023-04-19 R/analyze_images_GUI.R
Minor changes to functions

@@ -8,7 +8,7 @@
 #' @return Runs an automatic image analysis
 #' @export
 
-analyze_images<-function(){
+analyze_images_GUI<-function(){
   library(shiny)
   library(shinyalert)
   library(shinyFiles)

### 2023-04-19 cellcount_intro.Rmd
Updates made

@@
 title: "Introduction to *cellcount* - Draft vignette"
-author: "*T.E. Harman, M. Vandersea, D.E. Berthold, H.D. Laughinghouse, B. Boyd, K. Pokrzywinski*"
-date: "02/27/2023"
+author: "*T.E. Harman, M. Vandersea, K. Pokrzywinski*"
+date: "04/19/2023"
 abstract: >
@@
 
-*cellcount* is an R package distributed as part of the [Bioconductor](http://bioconductor.org) project. To install the package, start R and enter:
+*cellcount* is an R package distributed freely by NOAA NCCOS HAB-F on [GitHub](https://github.com/orgs/HABF-SDI-NCCOS-NOAA/repositories) and uses components in the RStudio IDE program.
+Make sure to install both the latest version of R and RStudio.
 
-```{r, message=FALSE}
-if (!requireNamespace("BiocManager", quietly = TRUE)) {
-  install.packages("BiocManager")
-}
-if (!require("cellcount")) {
-  BiocManager::install("cellcount")
-}
+To get started with *cellcount*, users must make a local clone of the repository using programs like [GitHub Desktop](https://desktop.github.com/) or similar programs.
+Once cloned, open the R project file and install the package locally by selecting 'build' and 'install package'.
+You should now have the *cellcount* package installed.
+Use the following code to bring *cellcount* into the R environment:
+
+```{r}
 library(cellcount)
@@
 
-The following packages are required. Please install and bring them into the R
-environment: 
+In addition, the following packages are required to use functions within *cellcount*:
+
+- *EBImage*
+- *tiff*
+- *pixmap*
+- *raster*
+- *beepr*
+
+*EBImage* uses the [Bioconductor](http://bioconductor.org) development software system while the remaining packages use the CRAN server built within R.
+Use the following code to install and load the packages into the R environment before starting:
+
 ```{r, message=FALSE}
@@
 library(EBImage)
+
 library(tiff)
@@
 
-Next, generate a data frame for count data to be added into using the following
-code:
+Now that we have the basics installed, we can begin using the *cellcount* package. Start by generating a data frame for count data to be added into using the following code:
+
 ```{r}
@@
 color label images and another for CSV files. In this example, we are using 
-images of *Microcystis* sp. cells. Please note that your 
+images *Microcystis* sp. cells. Please note that your 
 directory addresses will be different. We recommend using a dedicated directory 
@@
 Next, we will use the "mapped" function to turn the 8-bit image we just
-converted into a data matrix using the following code. This also includes a
-pixel intensity threshold adjustment that end-users can change; this removes
-any pixel that is under the threshold value to easily separate the background
-from cells:
+converted into a data matrix. In addition, we use an additional function to 
+determine the median intensity of each image. From this, users can add a numerical
+buffer on top of the determined median intensity, giving a more accurate approach
+to separating the background intensity from the intensity of cells within each
+image:
 
@@
 }
+mapped_avg <- function(x) {
+  x <- as.matrix(x)
+  return(x)
+}
 ```
 ```{r}
-imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
+img1<-lapply(grey_images,mapped_avg)
+pre_adj<-median(img1[[5]])
+main_adj<-pre_adj+0.2
+imagesMapped <- lapply(grey_images, mapped, threshold = main_adj)
 ```
@@
 from the "mapped" function, the contrast from the "greyscale" function 
-(if used), or could involve making adjustments to the "image_convert" function,
-such as the width and height of the rectangular window, the pixel area
-threshold (i.e., removes cells or other objects under a certain pixel area), 
-or the tolerance and 'ext' variables. Observe your initial results as indicated
-below and make adjustments where necessary:
+(if used), or could involve making adjustments to the "single_cell_convert" or
+"filamentous_convert" functions, such as the width and height of the rectangular 
+window, the pixel area threshold (i.e., removes cells or other objects under a 
+certain pixel area),or the tolerance and 'ext' variables. Observe your initial 
+results as indicated below and make adjustments where necessary:
 
@@
 }
-countCells <- function(x) {
+count_cells <- function(x) {
   countAll <- max(x)
@@
 }
-countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+count_images <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
   if (removeEdgeCells) {
@@
 }
-image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+single_cell_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
   image <- thresh(x, w = w, h = h, offset = offset)
@@
 ```{r}
-imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
-final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+imagesConverted <- single_cell_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
+final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
 display(final_img)
@@
 Once you have made necessary adjustments, use the following for-loop to
-analyze your images in bulk, one at a time. The for-loop uses the 
-"image_convert" function to produce unique color labels on cells within
-an image and to produce a final image showcasing these labels using the 
-"countImages" function. Finally, "countCells" is used to count the number of 
-unique color labels (i.e., cells). We also have integrated code to add these 
-count results to the data frame made earlier, and also to save each unique 
-color label image to the "image_savdir" directory:
+analyze your images in bulk, one at a time. In this example using *Microcystis* sp.
+images, the for-loop uses the "single_cell_convert" function to produce unique 
+color labels on cells within an image and to produce a final image showcasing 
+these labels using the "count_images" function. Finally, "count_cells" is used 
+to count the number of unique color labels (i.e., cells). We also have 
+integrated code to add these count results to the data frame made earlier, and 
+also to save each unique color label image to the "image_savdir" directory:
 
@@
 for (z in 1:length(images)) {
-  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
-  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
-  count <- countCells(imagesConverted)
+  imagesConverted <- single_cell_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
+  final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
+  count <- count_cells(imagesConverted)
   Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
@@
 outliers from the produced data, within a range of two standard deviations 
-above and below the median. The "outlier" function tags which images are
+above and below the median. The "outlier_detect" function tags which images are
 out of this range and dictates a warning within our data frame, which is shown 
@@
 ```{r,include=FALSE}
-outlier<-function(){
+outlier_detect<-function(){
   mean<-median(Cell.Count$Cell_Count)
@@
 Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
-Cell.Count$SD_range<-outlier()
+Cell.Count$SD_range<-outlier_detect()
 Cell.Count
@@
 In order to calculate cell density of these images, we can use the
-"cell.density" function that's built into the *cellcount* package. We first
+"cell_density" function that's built into the *cellcount* package. We first
 need to generate the total amount of cells counted from these images and
@@
 
-# Output data from test images in the *cellcount* package
+# Output data from test images in the *cellcount* package:
 
@@
 Test data results for *Microcystis* strain F-192
+<br>
+<br>
+<br>
+<br>
 
@@
 }
-image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+single_cell_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- fillHull(image)
+  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+filamentous_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
   image <- thresh(x, w = w, h = h, offset = offset)
@@
   return(cyano_total)
+}
+outlier_detect<-function(){
+  mean<-median(Cell.Count$Cell_Count)
+  sd<-sd(Cell.Count$Cell_Count)*2
+  sd_neg<-(mean-sd)
+  sd_pos<-(mean+sd)
+  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
+    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
+  return(SD_range_detection)
 }
@@
 
-imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
+mapped_avg <- function(x) {
+  x <- as.matrix(x)
+  return(x)
+}
+img1<-lapply(grey_images,mapped_avg)
+pre_adj<-median(img1[[5]])
+main_adj<-pre_adj+0.2
+imagesMapped <- lapply(grey_images, mapped, threshold = main_adj)
 
-imagesConverted <- image_convert2(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
-final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+imagesConverted <- filamentous_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
+final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
 display(final_img)
@@
 for (z in 1:length(images)) {
-  imagesConverted <- image_convert2(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
-  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
-  count <- countCells(imagesConverted)
+  adj1<-median(img1[[z]])
+  adj2<-adj1+0.2
+  imagesMapped <- lapply(grey_images, mapped, threshold = adj2)
+  imagesConverted <- filamentous_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
+  final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
+  count <- count_cells(imagesConverted)
   Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
@@
 Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
-outlier<-function(){
-  mean<-median(Cell.Count$Cell_Count)
-  sd<-sd(Cell.Count$Cell_Count)*2
-  sd_neg<-(mean-sd)
-  sd_pos<-(mean+sd)
-  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
-    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
-  return(SD_range_detection)
-}
-Cell.Count$SD_range<-outlier()
+Cell.Count$SD_range<-outlier_detect()
 
@@
 Test data results for *Anabena* sp.
+<br>
+<br>
+<br>
+<br>
 
@@
   image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}
+filamentous_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
+  image2 <- fillHull(image1)
   nf <- computeFeatures.shape(image2)
@@
 
-imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
+mapped_avg <- function(x) {
+  x <- as.matrix(x)
+  return(x)
+}
+img1<-lapply(grey_images,mapped_avg)
+pre_adj<-median(img1[[5]])
+main_adj<-pre_adj+0.2
+imagesMapped <- lapply(grey_images, mapped, threshold = main_adj)
 
-imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
-final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+imagesConverted <- single_cell_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
 display(final_img)
@@
 for (z in 1:length(images)) {
-  imagesConverted <- image_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
-  final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
-  count <- countCells(imagesConverted)
+  adj1<-median(img1[[z]])
+  adj2<-adj1+0.2
+  imagesMapped <- lapply(grey_images, mapped, threshold = adj2)
+  imagesConverted <- single_cell_convert(imagesMapped[[z]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
+  final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
+  count <- count_cells(imagesConverted)
   Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
@@
 
-# Shiny app user-interface for ***cellcount*** analysis
+# Shiny app user-interface for ***cellcount*** analysis:
 
@@
 
-The *input_data* function generates a Shiny UI where users enter information regarding directories, numerical and categorical data, 
+The *input_data_GUI* function generates a Shiny UI where users enter information regarding directories, numerical and categorical data, 
 and selecting analysis type. Tabs to the left of the UI separate particular categories for easy organization. Once all data has been
@@
 
-(SHOW UI IMAGES HERE)
+![Visual demonstration of the Shiny GUI for inputting data.](C:/Users/Tyler.Harman/Pictures/cellcount-UI_1.png)
 
-*analyze_images* is a simple UI that contains two interactive buttons.
+<br>
+
+*analyze_images_GUI* is a simple UI that contains two interactive buttons.
 "Run Image Analysis" simply runs the image analysis previously outlined in this document and uses the inputted data from the previous function.
-Once the completion sound has been played, users can click the "Close window" button to exit the UI.
+Once the completion sound has been played (using the *beepr* package), users can click the "Close window" button to exit the UI.
 Users can review the processed data in their respective directories.
 
-(SHOW UI IMAGE HERE)
+![Visual demonstration of the Shiny GUI for running image analysis.](C:/Users/Tyler.Harman/Pictures/cellcount-UI_2.png)
+
+# Package limitations:
 

### 2023-04-20 cellcount_intro.Rmd
New additions/changes made

@@
 ```{r,include=FALSE}
-savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
+savdir <- ("./test_data/")
+image_savdir <- ("./test_results/")
 ```
@@
 ```{r,include=FALSE}
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+images <- list.files("./test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("./test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 ```
@@
 
+### Test data results for *Microcystis* strain F-192
 ```{r,echo=FALSE}
@@
 ```
-Test data results for *Microcystis* strain F-192
-<br>
-<br>
 <br>
@@
 
+### Test data results for *Anabena* sp.
 ```{r,include=FALSE}
@@
 
-savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)
+savdir <- ("./test_data/")
+image_savdir <- ("./test_results/")
+images <- list.files("./test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
+images_names <- list.files("./test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)
 
@@
 ```
-Test data results for *Anabena* sp.
-<br>
-<br>
 <br>
@@
 
+### Test data results for *Shewanella* IRI-160
 ```{r,include=FALSE}
@@
 
-savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)
+savdir <- ("./test_data/")
+image_savdir <- ("./test_results/")
+images <- list.files("./test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
+images_names <- list.files("./test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)
 
@@
 ```
-Test data results for *Shewanella* IRI-160
 
@@
 
+### Inputting data
 The *input_data_GUI* function generates a Shiny UI where users enter information regarding directories, numerical and categorical data, 
@@
 
-<br>
+### Running the image analysis
 

### 2023-04-20 cellcount_intro.Rmd
Updates made to file path convention

@@
 
-To get started with *cellcount*, users must make a local clone of the repository using programs like [GitHub Desktop](https://desktop.github.com/) or similar programs.
-Once cloned, open the R project file and install the package locally by selecting 'build' and 'install package'.
+To get started with *cellcount*, users can make a local clone of the repository using programs like [GitHub Desktop](https://desktop.github.com/) or similar programs.
+Users may also download the package from the provided release on GitHub that contains a zipped folder with the package material.
+Once cloned, open the R project file and install the package locally by selecting 'build' and 'install package', or use the shortcut Crtl+Shift+B (Cmd+Shift+B for MacOS users).
 You should now have the *cellcount* package installed.
@@
 ```{r,include=FALSE}
-savdir <- ("./test_data/")
-image_savdir <- ("./test_results/")
+savdir <- ("./inst/extdata/test_data/")
+image_savdir <- ("./inst/extdata/test_results/")
 ```
@@
 ```{r,include=FALSE}
-images <- list.files("./test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
-images_names <- list.files("./test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+images <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 ```
@@
 
-savdir <- ("./test_data/")
-image_savdir <- ("./test_results/")
-images <- list.files("./test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
-images_names <- list.files("./test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)
+savdir <- ("./inst/extdata/test_data/")
+image_savdir <- ("./inst/extdata/test_results/")
+images <- list.files("./inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
+images_names <- list.files("./inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)
 
@@
 
-savdir <- ("./test_data/")
-image_savdir <- ("./test_results/")
-images <- list.files("./test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
-images_names <- list.files("./test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)
+savdir <- ("./inst/extdata/test_data/")
+image_savdir <- ("./inst/extdata/test_results/")
+images <- list.files("./inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
+images_names <- list.files("./inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)
 
@@
 
+The current version of this package may be limiting for some users regarding processing power.
+Analysis with this package uses local memory, which may be limiting if expected to analyze large amounts of images in batches.
+We recommend at least 16GB of localized RAM, the use of a cloud computing supercomputer (i.e.
+Microsoft Azure), or organize images in easier-to-analyze batches to reduce local memory strain.
+
+Users must be aware that this package does not contain abilities to identify distinct cells.
+The development of this package utilized carefully cultured monoculture of various cyanobacteria species.
+Users may find difficulty in using this analysis tool if cultures contain heavy amounts of particulates, other cellular debris, and contamination of other species.
+It is advised to use this tool for single-species only, however future developments of this tool are planned include identifying cell features and building image libraries of specific cellular species for use in environmental samples.
+
+Outside of package limitations, users must be aware of limitations that may arise from poor microscopy imaging; reduce lighting as much as possible, avoid heavy background fluorescence, reduce debris/particulates, and so forth.
+Good images produce good *cellcount* results - keep this in mind as you utilize this tool.
+
 # Literature cited:

### 2023-04-20 analyze_img.R
Updates: removal of functions within the script, as well as correcting errors

@@ -4,65 +4,6 @@ library(pixmap)
 library(raster)
 library(beepr)
 
-border <- function(x, y) {
-  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
-}
-cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
-  MCF <- filtration.area / FOV
-  N <- x / images
-  cyano <- (MCF * N) / volume
-  cyano_total <- cyano * total.volume
-  return(cyano_total)
-}
-countCells <- function(x) {
-  countAll <- max(x)
-  dims <- dim(x)
-  border1 <- border(x, dims)
-  ids <- unique(border1[which(border1 != 0)])
-  countInner <- countAll - length(ids)
-  return(countInner)
-}
-countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
-  if (removeEdgeCells) {
-    dims <- dim(image)
-    border1 <- border(image, dims)
-    ids <- unique(border1[which(border1 != 0)])
-    inner <- rmObjects(image, ids)
-    EBImage::colorLabels(inner, normalize = normalize)
-  } else {
-    EBImage::colorLabels(image, normalize = normalize)
-  }
-}
-greyscale <- function(x, contrast = 2) {
-  x <- contrast * x
-  x <- x[, , 1] + x[, , 2] + x[, , 3]
-  x <- x / max(x)
-  x <- normalize(x, inputRange = c(0.1, 0.75))
-  return(x)
-}
-single_cell_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
-  image <- thresh(x, w = w, h = h, offset = offset)
-  image1 <- fillHull(image)
-  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
-  nf <- computeFeatures.shape(image2)
-  nr <- which(nf[, "s.area"] < areathresh)
-  image3 <- rmObjects(image2, nr)
-  return(image3)
-}
-filamentous_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
-  image <- thresh(x, w = w, h = h, offset = offset)
-  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
-  image2 <- fillHull(image1)
-  nf <- computeFeatures.shape(image2)
[…]

### 2023-04-20 R/analyze_images_GUI.R
Updates to GUI as well as documentation regarding the new changes

@@ -1,9 +1,10 @@
 #' Analyze Images GUI
 #'
 #' Brings in data entered from the "input_data_GUI" function and automatically runs a script to analyze images
-#' pertaining to cells from fluorescent microscopy images. This uses a Shiny user-interface, where two buttons
-#' are available to interact with. One button "Run Image Analysis" runs the analysis script while the "Close Window"
-#' button closes the Shiny interface.
+#' pertaining to cells from fluorescent microscopy images. This uses a Shiny user-interface, where three buttons
+#' are available to interact with, as well as an interactive table below. The "Run Image Analysis" button runs 
+#' the analysis script, the "Close Window" button closes the Shiny interface, and the "Refresh" button allows 
+#' users to view the analyzed data once the analysis has completed.
 #' 
 #' @return Runs an automatic image analysis
 #' @export
@@ -14,31 +15,49 @@ analyze_images_GUI<-function(){
   library(shinyFiles)
   library(shinyjs)
   library(lubridate)
+  library(DT)
   
   jscode <- "shinyjs.closeWindow = function() { window.close(); }"
   
   ui = fluidPage(
     useShinyjs(),
     h1("  "),
-    fluidRow(
-      column(4, offset=2,actionButton("buttonId", "Run Image Analysis",class = "btn-success",style='height:130px;width:375px;font-size:120%',icon=icon("wand-magic-sparkles")))),
-    fluidRow(
-      column(4, offset=2,actionButton("close", "Close window",class = "btn-danger",style='height:130px;width:375px;font-size:120%',icon=icon("check"))))
+    h1("  "),
+    actionButton("buttonId", "Run Image Analysis",class = "btn-success",style='height:75px;width:252px;font-size:140%',icon=icon("wand-magic-sparkles"),style="display:center-align"),
+    actionButton("close", "Close window",class = "btn-danger",style='height:75px;width:252px;font-size:140%',icon=icon("check"),style="display:center-align"),
+    actionButton("BRefresh","Refresh",class = "btn-success",style='height:75px;width:252px;font-size:140%',icon=icon("arrows-rotate"),style="display:center-align"),
+    h1("  "),
+    DT::dataTableOutput('data')
   )
   
   
   server = function(input, output, session) {
+    rv <- reactiveVal(Cell.Count)
+    
+    output$data <- DT::renderDataTable ({
+      DT::datatable(rv(), editable = TRUE)
+    })
     observeEvent(input$buttonId, {
[…]

### 2023-04-20 cellcount_intro.Rmd
Minor updates to text regarding GUI functions

@@
 
-*analyze_images_GUI* is a simple UI that contains two interactive buttons.
-"Run Image Analysis" simply runs the image analysis previously outlined in this document and uses the inputted data from the previous function.
-Once the completion sound has been played (using the *beepr* package), users can click the "Close window" button to exit the UI.
-Users can review the processed data in their respective directories.
+*analyze_images_GUI* is a simple UI that contains three interactive buttons as well as a data table below.
+"Run Image Analysis" simply runs the image analysis previously outlined in this document and uses the inputted data from the *input_data_GUI* function.
+Once the completion sound has been played (using the *beepr* package), users can click the "Refresh" button to update the table below with the analyzed data and review.
+Once finished with analysis, the "Close window" button to exit the UI.
 

### 2023-04-20 manuscript_files/draft_manuscript.Rmd
Updated function names, manuscript text, added citations, updated authorship

--- 2023-03-03 manuscript_files/draft_manuscript.Rmd
@@
 
-  - name: Brandon Boyd
-    affiliation: U.S. Army Corps of Engineers, Engineer Research and Development Center, Vicksburg, MS 39180, USA
-
   - name: Kaytee Pokrzywinski
@@
 In addition, additional tools are presented in this package using [R Shiny](https://cran.r-project.org/package=shiny) [@Chang2022] interfaces to demonstrate ease-of-use for researchers not familiar with the R environment.
-All materials outlined in this article, along with tutorials, are found in [GitHub.](https://github.com/teharman/cellcount_main)
+All materials outlined in this article, along with tutorials, are found on NOAA's NCCOS HAB-F [GitHub.](https://github.com/orgs/HABF-SDI-NCCOS-NOAA/repositories)
 
@@
 Technology has become significant in advancing scientific methods, and it is no less significant in its role in advancing enumeration processes.
-Several commercially available products such as the FlowCam (Fluid Imaging Technologies Inc.) and Imaging FlowCytobot (McLane Research Laboratories Inc.) can be used to simplify enumeration methods (REF).
-**These technologies can be accurate, demonstrate time-savings, and simplify data collection (REF).
+**Several commercially available products such as the FlowCam (Fluid Imaging Technologies Inc.) and Imaging FlowCytobot (McLane Research Laboratories Inc.) can be used to simplify enumeration methods (REF).
+These technologies can be accurate, demonstrate time-savings, and simplify data collection (REF).
 However, several limitations can be observed such as particle size, misidentification of cells, and overall grouping of particles within samples, which can require assistance from researchers (REF).** In addition, these technologies are expensive to purchase and maintain, reducing the nature of these technologies being widely available [@Stauffer2019].
@@
 Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
-savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_results/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_results/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 file_name<-paste0("F192_")
@@
 The *border* function is built into several other functions within ***cellcount*** (i.e.
-*countImages*, *image_convert*, etc.), where its primarily goal is to predict cells along a predetermined border and to remove these cells from the final cell count estimate from each image.
+*count_images*, *single_cell_convert*, etc.), where its primarily goal is to predict cells along a predetermined border and to remove these cells from the final cell count estimate from each image.
 (WHY IS THAT IMPORTANT-OUTLINE IT HERE).
@@
 
-The *image_convert* functions take the data matrices corresponding to each image and counts individual cells within an image, omitting the cells indentified from the *border* function.
+The *single_cell_convert* functions take the data matrices corresponding to each image and counts individual cells within an image, omitting the cells indentified from the *border* function.
 Fluorescence microscopy images analyzed by this function are of a single fluorescent dimension (i.e.
 auto-fluorescence, DAPI, etc.).
-End-users should note that this function is useful only for analyzing images containing single-cell cyanobacteria.
+End-users should note that this function, like the name suggests, is useful only for analyzing images containing single-cell cyanobacteria.
 *fillHull* is placed first in this function before *watershed* in order to disseminate single-cells with more accuracy.
-A parallel function in ***cellcount***, *image_convert2*, is similarly organized but meant for filamentous or colonial-type cyanobacteria.
-The difference with this is largely where *watershed* is placed before *fillHull*, opposite of *image_convert*, which improves accuracy.
+A parallel function in ***cellcount***, *filamentous_convert*, is similarly organized but meant for filamentous or colonial-type cyanobacteria.
+The difference with this is largely where *watershed* is placed before *fillHull*, opposite of *single_cell_convert*, which improves accuracy.
 Within these two functions, several variables can be adjusted by end-users, such as the width and height of the rectangular window (see ***EBImage***), the pixel area threshold (i.e., removes cells or other objects under a certain pixel area), and the tolerance and ‘ext’ variables (see ***EBImage***).
@@
 ```{r}
-image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+single_cell_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
   image <- thresh(x, w = w, h = h, offset = offset)
@@
 ```{r}
-image_convert2 <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+filamentous_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
   image <- thresh(x, w = w, h = h, offset = offset)
@@
 
-Once the conversion occurs, the *countImages* function generates a color-label image, where each separated cell is given a unique color-label.
+Once the conversion occurs, the *count_images* function generates a color-label image, where each separated cell is given a unique color-label.
 This function includes removing cells along a predetermined border using the *border* function in order to improve count accuracy.
@@
 ```{r}
-countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+count_images <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
   if (removeEdgeCells) {
@@
 ```{r,echo=FALSE,fig.cap="A converted color-label image of *Microcystis* sp."}
-imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
-final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
+imagesConverted <- single_cell_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
+final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
 display(final_img)
@@
 
-The final function, *countCells*, counts the number of cells resulting from the image generated by the *countImages* function.
+The final function, *count_cells*, counts the number of cells resulting from the image generated by the *count_images* function.
 Once again, the *border* function is integrated here to omit counting cells within its predetermined border.
@@
 ```{r}
-countCells <- function(x) {
+count_cells <- function(x) {
   countAll <- max(x)
@@
 Informal discussions with researchers ignited the development of a graphical user interface (GUI) using Shiny apps, paired with functions from the ***cellcount*** package.
-This results in the use of two functions, *input_data* and *analyze images*, to run the entire analysis process as outlined in the *sandbox_analysis_test.R* script.
-In addition to packages required by ***cellcount***, these GUIs require the ***shiny*** [@Chang2022], ***shinyalert*** [@1Attali2021], ***shinyFiles*** [@Pedersen2022], ***shinyjs*** [@2Attali2021], and ***beepr*** [@Baath2018] libraries.
+This results in the use of two functions, *input_data_GUI* and *analyze images_GUI*, to run the entire analysis process as outlined in the *sandbox_analysis_test.R* script.
+In addition to packages required by ***cellcount***, these GUIs require the ***shiny*** [@Chang2022], ***shinyalert*** [@1Attali2021], ***shinyFiles*** [@Pedersen2022], ***shinyjs*** [@2Attali2021], ***DT*** [@Xie2023], and ***beepr*** [@Baath2018] libraries.
 
-*input_data* uses a Shiny app UI for end-users to easily fill out the needed information in order to correctly label analyzed images, provide numerical inputs for variables, select image analysis type, and easily select directories containing images and saving data.
+*input_data_GUI* uses a Shiny app UI for end-users to easily fill out the needed information in order to correctly label analyzed images, provide numerical inputs for variables, select image analysis type, and easily select directories containing images and saving data.
 The UI format is a simple approach, with tabs outlined on the left-hand side of the interface for its respective purpose (i.e., directory selection, analysis type selection, etc.).
@@
 
-*analyze_images* uses a simplistic Shiny UI which consists of two buttons, one to run the analysis and the other to close the UI interface.
-Once the proper data has been entered from the *input_data* function, users simply click the "Run Image Analysis" button to begin.
-Sounds from the ***beepr*** package alert users of completion of the analysis, where the "Close window" button exits the UI window.
+*analyze_images_GUI* uses a simplistic Shiny UI which consists of three buttons, one to run the analysis, another to close the UI interface, and one to refresh the available data table below once the analysis has completed.
+Once the proper data has been entered from the *input_data_GUI* function, users simply click the "Run Image Analysis" button to begin.
+Sounds from the ***beepr*** package alert users of completion of the analysis, allowing users to refresh the data table to see their results.
+Once satisified, users can click the "Close window" button to exit the UI window.
 Users can navigate to their specific directories to find analyzed images and related .csv files for their review.
@@
     URL: https://cran.r-project.org/web/packages/raster/index.html
-...
+
+  - type: package
+    id: Xie2023
+    author:
+      - family: Xie
+        given: Y.
+      - family: Cheng
+        given: J.
+      - family: Tan
+        given: X.
+      - family: Allaire
+        given: J.J.
+      - family: Girlich
+        given: M.
+      - family: Ellis
+        given: G.F.
+      - family: Rauh
+        given: J.
+      - family: Reavis
+        given: B.
+      - family: Gersen
+        given: L.
+      - family: Szopka
+        given: B.
+      - family: Pickering
+        given: A.
+      - family: Holmes
+        given: W.
+      - family: Marttila
+        given: M.
+      - family: Quintero
+        given: A.
+      - family: Laurent
+        given: S.
+    issued:
+      date-parts:
+      - - 2023
+        - 1
+        - 17
+    title: 'DT: A Wrapper of the JavaScript Library DataTables'
+    URL: https://cran.r-project.org/web/packages/DT/index.html
+---

### 2023-04-20 R/input_data_GUI.R
Changes made to function - button interactions and styling

@@ -13,6 +13,7 @@ input_data_GUI<-function(){
   library(shinyalert)
   library(shinyFiles)
   library(shinyjs)
+  library(shinyBS)
   
   jscode <- "shinyjs.closeWindow = function() { window.close(); }"
   
@@ -43,9 +44,11 @@ input_data_GUI<-function(){
                fixedRow(
                  column(2,offset=2,shinyDirButton('path1','CSV save directory','Please select a directory to save CSV files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
                ),
+               h4("  "),
                fixedRow(
                  column(2,offset=2,shinyDirButton('path2','Image save directory','Please select a directory to save Image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
                ),
+               h4("  "),
                fixedRow(
                  column(2,offset=2,shinyDirButton('path3','Image analysis directory','Please select a directory containing image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
                )
@@ -60,47 +63,42 @@ input_data_GUI<-function(){
       ),
       tabPanel(h4("Data inputs"),
                h3("   "),
-               fixedRow(
-                 column(2,offset=2,actionButton("run4", "Species name", class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("clipboard")))
+               fluidRow(
+                 column(1,bsButton("run4", align="center", "Species name", class = "btn-danger",size='large',icon=icon("clipboard"))),
+                 column(6,offset=5,bsButton("run6", align="center", 'Threshold adjustment', class = "btn-danger",size='large',icon=icon("gear")))
                ),
-               fixedRow(
-                 column(2,offset=2,actionButton("run6", 'Threshold adjustment', class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("gear")))
+               h4("  "),
+               fluidRow(
+                 column(1,bsButton("run7", 'Field of View', class = "btn-danger",size='large',icon=icon("magnifying-glass"))),
+                 column(6,offset=5,bsButton("run8", 'Number of images', class = "btn-danger",size='large',icon=icon("image")))
                ),
-               fixedRow(
[…]

### 2023-04-20 demo_analysis_test.R
Demo analysis script - initial upload

@@ -1,0 +1,78 @@
+library(EBImage)
+library(tiff)
+library(pixmap)
+library(raster)
+library(beepr)
+library(outliers)
+library(cellcount)
+
+mapped_avg <- function(x) {
+  x <- as.matrix(x)
+  return(x)
+}
+
+# clear environment and free unused memory
+gc()
+
+Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
+
+file_name<-paste0("F192_") #Change the name of species analyzed here
+
+#Change directories here
+savdir <- ("./inst/extdata/test_data/")
+image_savdir <- ("./inst/extdata/test_results/")
+images <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+
+imgNames <- paste0(file_name, images_names)
+read_images <- lapply(images, readTIFF)
+img_transposed <- lapply(read_images,aperm,c(2,1,3))
+names(images) <- imgNames
+
+grey_images <- lapply(img_transposed, greyscale, contrast = 4)
+display(grey_images[[2]]) #visualize contrast/brightness adjustment here
+
+img1<-lapply(grey_images,mapped_avg)
+
+for (j in 1:length(images)) {
+  adj1<-median(img1[[j]])
+  adj2<-adj1+0.2
+  imagesMapped <- lapply(grey_images, mapped, threshold = adj2) #background intensity threshold adjustment
+  imagesConverted <- single_cell_convert(imagesMapped[[j]], w = 17, h = 17, offset = 0.001, areathresh = 250, tolerance = 1, ext = 1)
+  final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
+  count <- count_cells(imagesConverted)
+  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[j]], count)
+  analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[j]), "_analyzed.tiff")
+  writeImage(final_img, files = paste0(image_savdir, analyzed_image),compression=c("LZW"))
+}
+
+Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
[…]

### 2023-04-26 cellcount_intro.Rmd
Updates made - initial changes from author review and overall changes to third person narration.

@@
 author: "*T.E. Harman, M. Vandersea, K. Pokrzywinski*"
-date: "04/19/2023"
+date: "`r Sys.Date()`"
 abstract: >
-  The *cellcount* package provides wrapper functions for  [*EBImage*](http://bioconductor.org/packages/devel/bioc/vignettes/EBImage/inst/doc/EBImage-introduction.html) to simplfy quantifying bacteria and cyanobacteria via flourescence microscopy.
-Cyanobacteria enumeration is an important tool to incorporate data results into mathematical models to assess bloom dynamics.
-However, standard manual cyanobacteria enumeration is considered time-consuming and can impact overall random error due to different inaccuracies from researcher to researcher.
-Here, we outline a new open-source tool, *cellcount*, a package designed for the computing language R to assist with cyanobacteria enumeration.
-Fluorescent microscopy images are analyzed by this package to generate cyanobacteria counts to assess overall cell density from samples.
+  The ***cellcount*** package provides a simple approach to quantifying bacteria and cyanobacteria *via* flourescence microscopy.
+Microbial enumeration is an important tool for validating new methods and critical to assess potential ecological changes.
+However, standard manual microbial enumeration is time-consuming and notoriously difficult due to large colonies, filaments, and dense biomats.
+Conventional counting methods require specialized expertise and are error prone due to variations between researchers.
+Additionally, other methods of counting cells or particles (i.e.
+coulter counter, flow cytometer, etc.) are not suitable for filamentous or colonial morphotypes that can clog flow cells and apertures.
+Here, we outline a new open-source package, ***cellcount***, designed in the R computing language, to quickly and accurately count various microbial morphotypes.
+In this open access package, fluorescent microscope images are analyzed to generate particle counts and quantify overall cell density from samples at comparable concentrations as conventional counting methods.
 output: html_document
@@
 
-*cellcount* is an R package distributed freely by NOAA NCCOS HAB-F on [GitHub](https://github.com/orgs/HABF-SDI-NCCOS-NOAA/repositories) and uses components in the RStudio IDE program.
-Make sure to install both the latest version of R and RStudio.
+***cellcount*** is an open-source R package distributed by NOAA NCCOS HAB-F on [GitHub](https://github.com/orgs/HABF-SDI-NCCOS-NOAA/repositories) and uses components in the RStudio IDE program.
+It is advised to use both the latest version of R and RStudio before beginning.
 
-To get started with *cellcount*, users can make a local clone of the repository using programs like [GitHub Desktop](https://desktop.github.com/) or similar programs.
+To get started with ***cellcount***, users can make a local clone of the repository using programs like [GitHub Desktop](https://desktop.github.com/) or similar programs.
 Users may also download the package from the provided release on GitHub that contains a zipped folder with the package material.
 Once cloned, open the R project file and install the package locally by selecting 'build' and 'install package', or use the shortcut Crtl+Shift+B (Cmd+Shift+B for MacOS users).
-You should now have the *cellcount* package installed.
-Use the following code to bring *cellcount* into the R environment:
+Once the ***cellcount*** package is installed, the following code can be used to bring ***cellcount*** into the R environment:
 
@@
 
-In addition, the following packages are required to use functions within *cellcount*:
+In addition, the following packages are required to use functions within ***cellcount***:
 
-- *EBImage*
-- *tiff*
-- *pixmap*
-- *raster*
-- *beepr*
+- [***EBImage***](http://bioconductor.org/packages/devel/bioc/vignettes/EBImage/inst/doc/EBImage-introduction.html)
+- [***tiff***](https://www.rforge.net/tiff/)
+- [***pixmap***](https://cran.r-project.org/package=pixmap)
+- [***raster***](https://cran.r-project.org/web/packages/raster/index.html)
+- [***beepr***](https://cran.r-project.org/web/packages/beepr/index.html)
 
-*EBImage* uses the [Bioconductor](http://bioconductor.org) development software system while the remaining packages use the CRAN server built within R.
+***EBImage*** uses the [Bioconductor](http://bioconductor.org) development software system while the remaining packages use the CRAN server built within R.
 Use the following code to install and load the packages into the R environment before starting:
@@
 
-Now that we have the basics installed, we can begin using the *cellcount* package. Start by generating a data frame for count data to be added into using the following code:
+After the basic supporting packages have been installed, an empty data frame can be generated to record resulting count data from the ***cellcount*** package using the following code:
 
@@
 
-You'll need to generate a couple of directories; one for saving the unique
-color label images and another for CSV files. In this example, we are using 
-images *Microcystis* sp. cells. Please note that your 
-directory addresses will be different. We recommend using a dedicated directory 
-for *cellcount* results:
+In order to appropriately save files from the ***cellcount*** analysis, some directories will be required; one for saving the unique color label images and another for CSV files.
+Please note that directory addresses will be different based on operating system and location.
+It is recommended to use a dedicated directory for ***cellcount*** results:
 
@@
 
-Next, you'll need to read in the images you are interested in processing using
-the following code. Once again, your directories will be different and must be
-updated to the correct file address. However, in this example we will be using
-the provided test images from this package:
+Next, images that are of interest will need to read into the R environment using
+the following code. Once again, directories will be different and must be
+updated to the correct file address. This example will use 'tif' images of *Microcystis* sp. that are provided within the ***cellcount*** package:
 
-```{r,message=FALSE}
-images <- list.files("C:/file path of images to analyze", pattern = "tif", full.name = T)
-images_names <- list.files("C:/file path of images to analyze", pattern = "tif", full.name = F)
-```
-```{r,include=FALSE}
+```{r}
 images <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
@@
 
-In addition, you'll need to assign the file names to a vector along with adding
-in characterization to each file for easy organization:
+In addition, users can assign additional identifications to each file for easy organization, as shown here:
 
@@
 
-Now we will "read" in the images that we are interested in analyzing. The
-following code outlines this step using the "readTIFF" function from the 
-*tiff* package. In addition, we'll need to re-orient the image as it's uploaded
-into the environment. The "readTIFF" function uploads images in a portrait 
-format - the "aperm" function reorients the uploaded image to a landscape format:
+Images from the assigned directory can now be "read" into the environment for analysis. The
+following code outlines this step using the *readTIFF* function from the 
+***tiff*** package. In addition, images will need to be re-oriented as the *readTIFF* function uploads images in a portrait 
+format. The *aperm* function shown below re-orients the uploaded image to a landscape format:
 
@@
 
-Now that our images are within the R environment, we can begin to convert these
-images into useful data that R can use to process images. First, we begin by
-converting 24-bit images into 8-bit images using the "greyscale" function (see 
+Once images are within the R environment, the process to convert these
+images into useful data can begin. First, begin by
+converting 24-bit images into 8-bit images using the *greyscale* function (see 
 photo below). This function also has a built-in numerical contrast adjustment 
-that end-users can adjust:
+that end-users can change:
 
@@
 
-Next, we will use the "mapped" function to turn the 8-bit image we just
-converted into a data matrix. In addition, we use an additional function to 
-determine the median intensity of each image. From this, users can add a numerical
+The *mapped* function turns the converted 8-bit image into a data matrix. An additional function to 
+determine the median intensity of each image is also used. From this, users can add a numerical
 buffer on top of the determined median intensity, giving a more accurate approach
 to separating the background intensity from the intensity of cells within each
-image:
+image. For example, if an image has a median intensity of 0.05, with a numerical buffer of 0.2,
+the threshold is equal to 0.25. Any pixel within the image below 0.25 will be omitted from the matrix and assigned a value of zero.
 
@@
 
-Before we analyze our images in bulk, set aside some code space to visually 
-analyze your images results. This could include adjusting the threshold
+Before images are analyzed in bulk, it is encouraged to visually 
+analyze image results and determine if changes need to be made. This could include adjusting the threshold
 from the "mapped" function, the contrast from the "greyscale" function 
@@
 window, the pixel area threshold (i.e., removes cells or other objects under a 
-certain pixel area),or the tolerance and 'ext' variables. Observe your initial 
+certain pixel area), or the tolerance and 'ext' variables. Observe initial 
 results as indicated below and make adjustments where necessary:
@@
 
-Once you have made necessary adjustments, use the following for-loop to
-analyze your images in bulk, one at a time. In this example using *Microcystis* sp.
+Once necessary adjustments have been made, the following for-loop is used to
+analyze images of interested in bulk, one at a time. In this example, using the provided demo *Microcystis* sp.
 images, the for-loop uses the "single_cell_convert" function to produce unique 
@@
 these labels using the "count_images" function. Finally, "count_cells" is used 
-to count the number of unique color labels (i.e., cells). We also have 
-integrated code to add these count results to the data frame made earlier, and 
-also to save each unique color label image to the "image_savdir" directory:
+to count the number of unique color labels (i.e., cells). Code has also been integrated in the for-loop
+to add these count results to the 'Cell.Count' data frame made earlier, and 
+to save each unique color label image to the "image_savdir" directory:
 
@@
 
-Data from the image analysis has been generated in our data frame "Cell.Count"
-that we previously made. We can use the following code to view our results:
+Data from the image analysis has been written to the data frame "Cell.Count", where
+calling the data frame can view these results:
 
@@
 
-Some images that were initially taken may be too poor of quality or contain
-high background fluorescence, among other errors. The code below analyzes 
+In same cases, some images that were initially taken may be too poor of quality or contain
+high background fluorescence, among other errors that may arise. The following code analyzes 
 outliers from the produced data, within a range of two standard deviations 
@@
 below. This alerts the end-user to perhaps re-image a particular sample, or
-to investigate the error further.
+to investigate the error further:
 
@@
 
-In order to calculate cell density of these images, we can use the
-"cell_density" function that's built into the *cellcount* package. We first
-need to generate the total amount of cells counted from these images and
-add it to our existing data frame using the following code:
+In order to calculate cell density of these images, the
+"cell_density" function can be implemented, using equations from Wetzel and Likens, 1991. Generating the total amount of 
+cells counted from these images is needed, where it can be added to the 'Cell.Count' data frame using the following code:
 
@@
 *follow) fills in the blanks from the previous data produced by the* 
-*"outlier" function.*
+***outlier_detect*** *function.*
 
-Calculating cell density, taken from Wetzel and Likens 1991, requires multiple
+Calculating cell density requires multiple
 values. End-users will need the following information to precisely 
@@
 With the variables calculated, the following code uses these variables to 
-calculate cell density and adds the data into the "Cell.Count" data frame:
+calculate cell density and adds the data into the 'Cell.Count' data frame:
 
@@
 
-We can also calculate the average number of cells per image, as well as cells
+The average number of cells per image can also be calculated, as well as cells
 per mL using the following code. Note that 'cell.av' must be divided by the
@@
 
-We can now view our data frame to double check for any calculation errors:
+Calling upon the 'Cell.Count' data frame can show the new generated data and to double check for any calculation errors:
 
@@
 
-Finally, we write the following code to export our finalized .csv file to our
+Finally, use the following code to export the finalized data frame to a .csv file to the
 pre-determined save directory:
@@
 
-# Output data from test images in the *cellcount* package:
+# Output data from test images in the ***cellcount*** package:
 
@@
 
-Several informal discussions with researchers have expressed interest in a easy-to-use analysis interface for inexperienced R users
-to utilize the ***cellcount*** tool. The following overview uses ***Shiny*** apps to enter input data and run image analyses.
+Informal discussions with researchers have expressed interest in a easy-to-use analysis interface for inexperienced R users
+to utilize the ***cellcount*** tool. The following overview uses ***shiny*** apps to enter input data and run image analyses. Several packages are required to utilize these UI tools:
+
+- [***shiny***](https://cran.r-project.org/web/packages/shiny/index.html)
+- [***shinyalert***](https://cran.r-project.org/web/packages/shinyalert/index.html)
+- [***shinyFiles***](https://cran.r-project.org/web/packages/shinyFiles/index.html)
+- [***shinyjs***](https://cran.r-project.org/web/packages/shinyjs/index.html)
+- [***DT***](https://cran.r-project.org/web/packages/DT/index.html)
+- [***shinyBS***](https://cran.r-project.org/web/packages/shinyBS/index.html)
+
+Users can use the following code to bring packages into the R environment once the packages have been installed using CRAN:
+
+```{r,message=FALSE,warning=FALSE}
+library(shiny)
+library(shinyalert)
+library(shinyFiles)
+library(shinyjs)
+library(DT)
+library(shinyBS)
+```
 
@@
 The *input_data_GUI* function generates a Shiny UI where users enter information regarding directories, numerical and categorical data, 
-and selecting analysis type. Tabs to the left of the UI separate particular categories for easy organization. Once all data has been
-provided, users complete the input of data by selecting the "Completion" tab and clicking the "Submit data entries" button to export 
-data to the local R environment.
+and selecting analysis type.
+Tabs to the left of the UI separate particular categories for easy organization and contain interactive buttons or drop-down menus to enter and select appropriate data.
+Some buttons change from red to green to provide an organized data entry approach.
+Once all data has been provided, users complete the input of data by selecting the "Completion" tab and clicking the "Submit data entries" button to export data to the local R environment.
 
@@
 ### Running the image analysis
-
-*analyze_images_GUI* is a simple UI that contains three interactive buttons as well as a data table below.
+*analyze_images_GUI* is a simple UI that contains three interactive buttons as well as an interactive data table.
 "Run Image Analysis" simply runs the image analysis previously outlined in this document and uses the inputted data from the *input_data_GUI* function.
-Once the completion sound has been played (using the *beepr* package), users can click the "Refresh" button to update the table below with the analyzed data and review.
+Once the completion sound has been played (using the ***beepr*** package), users can click the "Refresh" button to update the table below with the analyzed data and review.
 Once finished with analysis, the "Close window" button to exit the UI.
@@
 Analysis with this package uses local memory, which may be limiting if expected to analyze large amounts of images in batches.
-We recommend at least 16GB of localized RAM, the use of a cloud computing supercomputer (i.e.
+It is recommended to have at least 16GB of localized RAM, the ability to use cloud computing (i.e.
 Microsoft Azure), or organize images in easier-to-analyze batches to reduce local memory strain.
@@
 Users must be aware that this package does not contain abilities to identify distinct cells.
-The development of this package utilized carefully cultured monoculture of various cyanobacteria species.
+The development of this package utilized monoculture of various cyanobacteria species.
 Users may find difficulty in using this analysis tool if cultures contain heavy amounts of particulates, other cellular debris, and contamination of other species.
-It is advised to use this tool for single-species only, however future developments of this tool are planned include identifying cell features and building image libraries of specific cellular species for use in environmental samples.
+It is advised to use this tool for single-species only, however future developments of this tool are planned to include identifying cell features and deep-learning mechanics that use image libraries of specific cellular species.
 
 Outside of package limitations, users must be aware of limitations that may arise from poor microscopy imaging; reduce lighting as much as possible, avoid heavy background fluorescence, reduce debris/particulates, and so forth.
-Good images produce good *cellcount* results - keep this in mind as you utilize this tool.
+Good images produce good ***cellcount*** results - keep this in mind as you utilize this tool.
+
+# Conclusions:
 

### 2023-05-01 cellcount_intro.Rmd
Re-formatting of document to 'distill' article, minor changes to documentation outline, and changes to text to reflect in-depth explanation of package materials

@@
 title: "Introduction to *cellcount* - Draft vignette"
-author: "*T.E. Harman, M. Vandersea, K. Pokrzywinski*"
+author: 
+  - name: Tyler E. Harman 
+    affiliation: Consolidated Safety Services, Fairfax, VA 22031, USA
+      National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
+    affiliation_email: tyler.harman@noaa.gov
+    orcid_id: 0000-0001-5007-8170
+
+  - name: Mark W. Vandersea
+    affiliation: National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
+
+  - name: David E. Berthold
+    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
+
+  - name: H. Dail Laughinghouse
+    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
+    orcid_id: 0000-0003-1018-6948
+
+  - name: Kaytee Pokrzywinski
+    affiliation: National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
+    orcid_id: 0000-0003-1523-9930
 date: "`r Sys.Date()`"
-abstract: >
+output: distill::distill_article
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+# Summary:
+
 The ***cellcount*** package provides a simple approach to quantifying bacteria and cyanobacteria *via* flourescence microscopy.
@@
 coulter counter, flow cytometer, etc.) are not suitable for filamentous or colonial morphotypes that can clog flow cells and apertures.
-Here, we outline a new open-source package, ***cellcount***, designed in the R computing language, to quickly and accurately count various microbial morphotypes.
+Here, we outline a new open-source package, ***cellcount***, a continuation of the report outlined in [@Pokrzywinski2019] designed in the R computing language, to quickly and accurately count various microbial morphotypes.
 In this open access package, fluorescent microscope images are analyzed to generate particle counts and quantify overall cell density from samples at comparable concentrations as conventional counting methods.
-output: html_document
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+In addition, additional tools are presented in this package using [R Shiny](https://cran.r-project.org/package=shiny) [@Chang2022] interfaces to demonstrate ease-of-use for researchers not familiar with the R environment.
+All materials outlined in this article, along with tutorials, are found on NOAA's NCCOS HAB-F [GitHub.](https://github.com/orgs/HABF-SDI-NCCOS-NOAA/repositories)
 
@@
 
-***cellcount*** is an open-source R package distributed by NOAA NCCOS HAB-F on [GitHub](https://github.com/orgs/HABF-SDI-NCCOS-NOAA/repositories) and uses components in the RStudio IDE program.
+***cellcount*** is an open-source R package distributed by NOAA NCCOS HAB-F on GitHub and uses components in the RStudio IDE program.
 It is advised to use both the latest version of R and RStudio before beginning.
@@
 
-- [***EBImage***](http://bioconductor.org/packages/devel/bioc/vignettes/EBImage/inst/doc/EBImage-introduction.html)
-- [***tiff***](https://www.rforge.net/tiff/)
-- [***pixmap***](https://cran.r-project.org/package=pixmap)
-- [***raster***](https://cran.r-project.org/web/packages/raster/index.html)
-- [***beepr***](https://cran.r-project.org/web/packages/beepr/index.html)
+- ***EBImage*** [@Pau2010]
+- ***tiff*** [@Urbanek2022]
+- ***pixmap*** [@Bivand2021]
+- ***raster*** [@Hijmans2023]
+- ***beepr*** [@Baath2018]
 
@@
 ```
-```{r}
+```{r,fig.cap="\\label Converted greyscale image of *Microcystis* sp."}
 grey_images <- lapply(img_transposed, greyscale, contrast = 1)
@@
 analyze image results and determine if changes need to be made. This could include adjusting the threshold
-from the "mapped" function, the contrast from the "greyscale" function 
-(if used), or could involve making adjustments to the "single_cell_convert" or
-"filamentous_convert" functions, such as the width and height of the rectangular 
+from the *mapped* function, the contrast from the *greyscale* function 
+(if used), or could involve making adjustments to the *single_cell_convert* or
+*filamentous_convert* functions, such as the width and height of the rectangular 
 window, the pixel area threshold (i.e., removes cells or other objects under a 
@@
 ```
-```{r}
+```{r,fig.cap="A converted color-label image of *Microcystis* sp."}
 imagesConverted <- single_cell_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 250, tolerance = 0.8, ext = 1)
@@
 analyze images of interested in bulk, one at a time. In this example, using the provided demo *Microcystis* sp.
-images, the for-loop uses the "single_cell_convert" function to produce unique 
+images, the for-loop uses the *single_cell_convert* function to produce unique 
 color labels on cells within an image and to produce a final image showcasing 
-these labels using the "count_images" function. Finally, "count_cells" is used 
+these labels using the *count_images* function. Finally, *count_cells* is used 
 to count the number of unique color labels (i.e., cells). Code has also been integrated in the for-loop
@@
 outliers from the produced data, within a range of two standard deviations 
-above and below the median. The "outlier_detect" function tags which images are
+above and below the median. The *outlier_detect* function tags which images are
 out of this range and dictates a warning within our data frame, which is shown 
@@
 In order to calculate cell density of these images, the
-"cell_density" function can be implemented, using equations from Wetzel and Likens, 1991. Generating the total amount of 
+*cell_density* function can be implemented, using equations from [@Wetzel2000]. Generating the total amount of 
 cells counted from these images is needed, where it can be added to the 'Cell.Count' data frame using the following code:
@@
 The average number of cells per image can also be calculated, as well as cells
-per mL using the following code. Note that 'cell.av' must be divided by the
+per mL using the following code. Note that 'cell.av' calculation must contain the
 number of images the end-user has analyzed, and the 'cell.mL' must be divided
@@
 Informal discussions with researchers have expressed interest in a easy-to-use analysis interface for inexperienced R users
-to utilize the ***cellcount*** tool. The following overview uses ***shiny*** apps to enter input data and run image analyses. Several packages are required to utilize these UI tools:
+to utilize the ***cellcount*** tool.
+The following overview uses 'Shiny' apps to enter input data and run image analyses.
+This results in the use of two functions, *input_data_GUI* and *analyze images_GUI*, to run the entire analysis process as outlined previously, and in the **sandbox_analysis_test.R** script.
+Several packages are required to utilize these UI tools:
 
-- [***shiny***](https://cran.r-project.org/web/packages/shiny/index.html)
-- [***shinyalert***](https://cran.r-project.org/web/packages/shinyalert/index.html)
-- [***shinyFiles***](https://cran.r-project.org/web/packages/shinyFiles/index.html)
-- [***shinyjs***](https://cran.r-project.org/web/packages/shinyjs/index.html)
-- [***DT***](https://cran.r-project.org/web/packages/DT/index.html)
-- [***shinyBS***](https://cran.r-project.org/web/packages/shinyBS/index.html)
+- ***shiny*** [@Chang2022]
+- ***shinyalert*** [@1Attali2021]
+- ***shinyFiles*** [@Pedersen2022]
+- ***shinyjs*** [@2Attali2021]
+- ***DT*** [@Xie2023]
+- ***shinyBS*** [@Bailey2022]
 
@@
 
+Cells with a uniform chloroplast (auto-fluorescence) or fluorescent dyes illuminating the entire cell are the most useful when using this version of ***cellcount*** for enumeration purposes.
+Other cells such as dinoflagellates (i.e.
+*Karenia* sp.) that do not contain uniform chloroplasts may be more difficult to enumerate as the ***cellcount*** approach may separate these lobed chloroplasts into separate sections, causing enumeration outputs to be unreliable.
+In-situ hybridization dyes that target specific species or genera may be helpful if other cellular features are not useful or rather hinder in providing accurate enumeration.
+
 Users must be aware that this package does not contain abilities to identify distinct cells.
@@
 
-# Literature cited:
+***cellcount*** provides an open-source approach to cyanobacteria enumeration that is accurate and rapid compared to standardized microscopy methods.
+The open-source approach allows researchers to access this tool for free and to modify and make additions the existing code to fulfill their particular needs, whereas the Shiny GUI allows users of all coding backgrounds to utilize this analysis tool.
 
-Wetzel, RG and Likens, GE. 1991. Limnological Analyses, Second Edition. Springer-
-Verlag. 391 pp.
+# Literature Cited
 
-Pau, G, Fuchs, F, Sklyar, O, Boutros, M, Huber, W.
-2010.
-EBImage—an R package for image processing with applications to cellular phenotypes, Bioinformatics, 26(7): 979–981, https://doi.org/10.1093/bioinformatics/btq046
+---
+references:
+  - type: technical-report
+    id: Pokrzywinski2019
+    author:
+      - family: Pokrzywinski
+        given: K.
+      - family: Boyd
+        given: B.M.
+      - family: Smith
+        given: S.J.
+    issued:
+      date-parts:
+      - - 2019
+        - 8
+    title: 'A high-throughput method for counting cyanobacteria using fluorescence microscopy'
+    DOI: 10.21079/11681/33870
+    URL: https://erdc-library.erdc.dren.mil/jspui/handle/11681/33870
 
-Pokrzywinski, K, Boyd, B, Smith, J. 2019. A High-throughput Method for Counting 
-Cyanobacteria Using Fluorescence Microscopy. ERDC TR-19-21.
+  - type: article-journal
+    id: Pau2010
+    author:
+      - family: Pau
+        given: G.
+      - family: Fuchs
+        given: F.
+      - family: Skylar
+        given: O.
+      - family: Boutros
+        given: M.
+      - family: Huber
+        given: W.
+    issued:
+      date-parts:
+      - - 2010
+        - 3
+        - 23
+    title: 'EBImage—an R package for image processing with applications to cellular phenotypes'
+    container-title: Bioinformatics
+    volume: 26
+    issue: 7
+    page: 979-981
+    DOI: 10.1093/bioinformatics/btq046
+    URL: https://academic.oup.com/bioinformatics/article/26/7/979/211835
+
+  - type: package
+    id: Urbanek2022
+    author:
+      - family: Urbanek
+        given: S.
+      - family: Johnson
+        given: K.
+    issued:
+      date-parts:
+      - - 2022
+        - 1
+        - 31
+    title: 'tiff: Read and Write TIFF Images. R package version 0.1-11.'
+    URL: https://www.rforge.net/tiff/
+
+  - type: package
+    id: Bivand2021
+    author:
+      - family: Bivand
+        given: R.
+      - family: Leisch
+        given: F.
+      - family: Maechler
+        given: Martin
+    issued:
+      date-parts:
+      - - 2021
+        - 1
+        - 29
+    title: 'Bitmap Images/Pixel Maps, R Package Version 0.4.12.'
+    URL: https://cran.r-project.org/package=pixmap
+
+  - type: book
+    id: Wetzel2000
+    author:
+      - family: Wetzel
+        given: R.G.
+      - family: Likens
+        given: G.E.
+    issued:
+      date-parts:
+      - - 2000
+        - 3
+        - 3
+    title: 'Limnological Analyses - 3rd Edition'
+    page: 271-274
+    DOI: 10.1007/978-1-4757-3250-4
+    URL: https://link.springer.com/book/10.1007/978-1-4757-3250-4
+
+  - type: package
+    id: Baath2018
+    author:
+      - family: Baath
+        given: R.
+      - family: Dobbyn
+        given: A.
+    issued:
+      date-parts:
+      - - 2018
+        - 6
+        - 4
+    title: 'beepr: Easily Play Notification Sounds on any Platform'
+    URL: https://cran.r-project.org/web/packages/beepr/index.html
+
+  - type: package
+    id: Hijmans2023
+    author:
+      - family: Hijmans
+        given: R.J
+      - family: van Etten
+        given: J.
+      - family: Sumner
+        given: M.
+      - family: Cheng
+        given: J.
+      - family: Baston
+        given: D.
+      - family: Bevan
+        given: A.
+      - family: Bivand
+        given: R.
+      - family: Busetto
+        given: L.
+      - family: Canty
+        given: M.
+      - family: Fasoli
+        given: B.
+      - family: Forrest
+        given: D.
+      - family: Ghosh
+        given: A.
+      - family: Golicher
+        given: D.
+      - family: Gray
+        given: J.
+      - family: Greenburg
+        given: J.A.
+      - family: Hiemstra
+        given: P.
+      - family: Hingee
+        given: K.
+      - family: Ilich
+        given: A.
+      - family: Karney
+        given: C.
+      - family: Mattiuzzi
+        given: M.
+      - family: Mosher
+        given: S.
+      - family: Naimi
+        given: B.
+      - family: Nowosad
+        given: J.
+      - family: Pebesma
+        given: E.
+      - family: Lamigueiro
+        given: O.P.
+      - family: Racine
+        given: E.B.
+      - family: Rowlingson
+        given: B.
+      - family: Shortridge
+        given: A.
+      - family: Venables
+        given: B.
+      - family: Wueest
+        given: R.
+    issued:
+      date-parts:
+      - - 2023
+        - 1
+        - 16
+    title: 'raster: Geographic Data Analysis and Modeling'
+    URL: https://cran.r-project.org/web/packages/raster/index.html
+
+  - type: package
+    id: Bailey2022
+    author:
+      - family: Bailey
+        given: E.
+    issued:
+      date-parts:
+      - - 2022
+        - 4
+        - 17
+    title: 'shinyBS: Twitter Bootstrap Components for Shiny'
+    URL: https://cran.r-project.org/web/packages/shinyBS/index.html
+
+  - type: package
+    id: Chang2022
+    author:
+      - family: Chang
+        given: W.
+      - family: Cheng
+        given: J.
+      - family: Allaire
+        given: J.
+      - family: Boutros
+        given: M.
+      - family: Huber
+        given: W.
+    issued:
+      date-parts:
+      - - 2022
+        - 12
+        - 15
+    title: 'Shiny: Web application framework for R'
+    URL: https://cran.r-project.org/web/packages/shiny/index.html
+
+  - type: package
+    id: Pedersen2022
+    author:
+      - family: Pedersen
+        given: T.L.
+      - family: Nijs
+        given: V.
+      - family: Schaffner
+        given: T.
+      - family: Nantz
+        given: E.
+    issued:
+      date-parts:
+      - - 2022
+        - 8
+        - 19
+    title: 'shinyFiles: A Server-Side File System Viewer for Shiny'
+    URL: https://cran.r-project.org/web/packages/shinyFiles/index.html
+
+  - type: package
+    id: 1Attali2021
+    author:
+      - family: Attali
+        given: D.
+      - family: Edwards
+        given: T.
+      - family: Wang
+        given: Z.
+    issued:
+      date-parts:
+      - - 2021
+        - 12
+        - 20
+    title: 'shinyalert: Easily Create Pretty Popup Messages (Modals) in Shiny'
+    URL: https://cran.r-project.org/web/packages/shinyalert/index.html
+
+  - type: package
+    id: 2Attali2021
+    author:
+      - family: Attali
+        given: D.
+    issued:
+      date-parts:
+      - - 2021
+        - 12
+        - 23
+    title: 'shinyjs: Easily Improve the User Experience of Your Shiny Apps in Seconds'
+    URL: https://cran.r-project.org/web/packages/shinyjs/index.html
+
+  - type: package
+    id: Xie2023
+    author:
+      - family: Xie
+        given: Y.
+      - family: Cheng
+        given: J.
+      - family: Tan
+        given: X.
+      - family: Allaire
+        given: J.J.
+      - family: Girlich
+        given: M.
+      - family: Ellis
+        given: G.F.
+      - family: Rauh
+        given: J.
+      - family: Reavis
+        given: B.
+      - family: Gersen
+        given: L.
+      - family: Szopka
+        given: B.
+      - family: Pickering
+        given: A.
+      - family: Holmes
+        given: W.
+      - family: Marttila
+        given: M.
+      - family: Quintero
+        given: A.
+      - family: Laurent
+        given: S.
+    issued:
+      date-parts:
+      - - 2023
+        - 1
+        - 17
+    title: 'DT: A Wrapper of the JavaScript Library DataTables'
+    URL: https://cran.r-project.org/web/packages/DT/index.html
+---

### 2023-05-01 manuscript_files/stats_code/RA_Code.R
Removed from local repository

@@ -1,296 +1,0 @@
-library(ggplot2)
-library(gridExtra)
-library(ggpubr)
-library(cowplot)
-
-setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data")
-
-RA<-read.csv("Regression_Analysis.csv",header=T)
-
-RA_F192<-subset(RA,Species=="F192")
-RA_F210<-subset(RA,Species=="F210")
-RA_Anabena<-subset(RA,Species=="Anabena")
-RA_Shewanella<-subset(RA,Species=="Shewanella")
-RA_E.coli<-subset(RA,Species=="E.coli")
-
-RA_F192[ nrow(RA_F192) + 1:6,] <- NA
-RA_F192$bot<-c(0,0,0,800000,800000,800000,1600000,1600000,1600000,2400000,2400000,2400000,3200000,3200000,3200000,4000000,4000000,4000000,4800000,4800000,4800000)
-RA_F192$top<-c(0,0,0,1200000,1200000,1200000,2400000,2400000,2400000,3600000,3600000,3600000,4800000,4800000,4800000,6000000,6000000,6000000,7200000,7200000,7200000)
-RA_F192$Script<-(RA_F192$Script/1000000)
-RA_F192$Visual<-(RA_F192$Visual/1000000)
-RA_F192$Microscope<-(RA_F192$Microscope/1000000)
-RA_F192$bot<-(RA_F192$bot/1000000)
-RA_F192$top<-(RA_F192$top/1000000)
-RA_F192$func1<-(sapply(RA_F192$bot, FUN = function(x){(x)}))
-RA_F192$func2<-(sapply(RA_F192$top, FUN = function(x){(x)}))
-RA_F192$num<-c(0,0,0,1,1,1,2,2,2,3,3,3,4,4,4,5,5,5,6,6,6)
-
-data.lm1<-lm(Visual~Script,data=RA_F192)
-summary(data.lm1)
-
-F192<-ggplot(data=RA_F192, aes(x=Script,y=Visual))+
-  theme_bw()+
-  labs(x=expression(paste(italic("cellcount"), " results (",x10^{6},")")))+
-  labs(y=expression(paste("Visual counts ( ",x10^{6},")")))+
-  theme(axis.title.x = element_text(size=15),
-        axis.title.y = element_text(size=15,vjust=2))+
-  theme(axis.text.x = element_text(size = 15),
-        axis.text.y = element_text(size = 15),
[…]

### 2023-05-01 manuscript_files/stats_code/RosnerTest.R

@@ -1,94 +1,0 @@
-library(outliers)
-library(EnvStats)
-
-setwd("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/data/")
-
-RA<-read.csv("Regression_Analysis_IND.csv",header=T)
-
-RA_F192<-subset(RA,Species=="F192")
-RA_F192A<-subset(RA_F192,Concentration=="A")
-RA_F192B<-subset(RA_F192,Concentration=="B")
-RA_F192C<-subset(RA_F192,Concentration=="C")
-RA_F192D<-subset(RA_F192,Concentration=="D")
-RA_F192E<-subset(RA_F192,Concentration=="E")
-
-rosnerTest(RA_F192A$Script)
-rosnerTest(RA_F192B$Script)
-rosnerTest(RA_F192C$Script)
-rosnerTest(RA_F192D$Script)
-rosnerTest(RA_F192E$Script)
-rosnerTest(RA_F192A$Visual)
-rosnerTest(RA_F192B$Visual)
-rosnerTest(RA_F192C$Visual)
-rosnerTest(RA_F192D$Visual)
-rosnerTest(RA_F192E$Visual)
-
-
-RA_F210<-subset(RA,Species=="F210")
-RA_F210A<-subset(RA_F210,Concentration=="A")
-RA_F210B<-subset(RA_F210,Concentration=="B")
-RA_F210C<-subset(RA_F210,Concentration=="C")
-RA_F210D<-subset(RA_F210,Concentration=="D")
-RA_F210E<-subset(RA_F210,Concentration=="E")
-
-rosnerTest(RA_F210A$Script)
-rosnerTest(RA_F210B$Script)
-rosnerTest(RA_F210C$Script)
-rosnerTest(RA_F210D$Script)
-rosnerTest(RA_F210E$Script)
-rosnerTest(RA_F210A$Visual)
-rosnerTest(RA_F210B$Visual)
-rosnerTest(RA_F210C$Visual)
-rosnerTest(RA_F210D$Visual)
-rosnerTest(RA_F210E$Visual)
-
-
-RA_Anabena<-subset(RA,Species=="Anabena")
-RA_AnabenaHigh<-subset(RA_Anabena,Concentration=="High")
-RA_AnabenaMed<-subset(RA_Anabena,Concentration=="Medium")
[…]

### 2023-08-30 cellcount_intro.Rmd
Minor changes to vignette file
Minor updates to vignette from co-author comments

@@
 However, standard manual microbial enumeration is time-consuming and notoriously difficult due to large colonies, filaments, and dense biomats.
-Conventional counting methods require specialized expertise and are error prone due to variations between researchers.
+Conventional counting methods like microscopy require specialized expertise and are error prone due to variations between microscopists.
 Additionally, other methods of counting cells or particles (i.e.
 coulter counter, flow cytometer, etc.) are not suitable for filamentous or colonial morphotypes that can clog flow cells and apertures.
-Here, we outline a new open-source package, ***cellcount***, a continuation of the report outlined in [@Pokrzywinski2019] designed in the R computing language, to quickly and accurately count various microbial morphotypes.
-In this open access package, fluorescent microscope images are analyzed to generate particle counts and quantify overall cell density from samples at comparable concentrations as conventional counting methods.
-In addition, additional tools are presented in this package using [R Shiny](https://cran.r-project.org/package=shiny) [@Chang2022] interfaces to demonstrate ease-of-use for researchers not familiar with the R environment.
-All materials outlined in this article, along with tutorials, are found on NOAA's NCCOS HAB-F [GitHub.](https://github.com/orgs/HABF-SDI-NCCOS-NOAA/repositories)
+Here, we outline a new open-source package, ***cellcount***, a continuation of the report outlined in [@Pokrzywinski2019] and designed in the R computing language, to quickly and accurately count various microbial morphotypes.
+With this open source package, fluorescent microscope images can be analyzed to generate particle counts and quantify overall cell density from samples at a rapid speed with the accuracy of comparable conventional counting methods.
+Furthermore, additional tools are presented in this package using [R Shiny](https://cran.r-project.org/package=shiny) [@Chang2022] interfaces to demonstrate ease-of-use for researchers not familiar with the R environment.
+All materials outlined in this article, along with tutorials, can be found on NOAA's National Center for Coastal Ocean Science Harmful Algal Bloom Forecasting (NOAA NCCOS HAB-F) [GitHub.](https://github.com/orgs/HABF-SDI-NCCOS-NOAA/repositories)
 
@@
 
-***cellcount*** is an open-source R package distributed by NOAA NCCOS HAB-F on GitHub and uses components in the RStudio IDE program.
+***cellcount*** is an open-source R package for enumerating particles in fluorescent microscopy images distributed by NOAA NCCOS HAB-F on GitHub and uses components in the RStudio IDE program.
 It is advised to use both the latest version of R and RStudio before beginning.
 
-To get started with ***cellcount***, users can make a local clone of the repository using programs like [GitHub Desktop](https://desktop.github.com/) or similar programs.
-Users may also download the package from the provided release on GitHub that contains a zipped folder with the package material.
+To get started with ***cellcount***, make a local clone of the repository using programs like [GitHub Desktop](https://desktop.github.com/).
 Once cloned, open the R project file and install the package locally by selecting 'build' and 'install package', or use the shortcut Crtl+Shift+B (Cmd+Shift+B for MacOS users).
@@
 Next, images that are of interest will need to read into the R environment using
-the following code. Once again, directories will be different and must be
-updated to the correct file address. This example will use 'tif' images of *Microcystis* sp. that are provided within the ***cellcount*** package:
+the following code. This example uses 'tif' images of *Microcystis* sp. that are provided within the ***cellcount*** package:
 
@@
 
-Once images are within the R environment, the process to convert these
-images into useful data can begin. First, begin by
+Once images are within the R environment, the process to quantify particles within an image can begin. First, begin by
 converting 24-bit images into 8-bit images using the *greyscale* function (see 
@@
 
-Before images are analyzed in bulk, it is encouraged to visually 
-analyze image results and determine if changes need to be made. This could include adjusting the threshold
+Before images can be analyzed with batch processing, it is encouraged to visually 
+assess image results and determine if fine adjustments need to be made. This could include adjusting the threshold
 from the *mapped* function, the contrast from the *greyscale* function 
@@
 Once necessary adjustments have been made, the following for-loop is used to
-analyze images of interested in bulk, one at a time. In this example, using the provided demo *Microcystis* sp.
+interatively process batch images of interest. In this example, using the provided demo *Microcystis* sp.
 images, the for-loop uses the *single_cell_convert* function to produce unique 
@@
 these labels using the *count_images* function. Finally, *count_cells* is used 
-to count the number of unique color labels (i.e., cells). Code has also been integrated in the for-loop
+to count the number of unique color labels (i.e., cells or particles). Code has also been integrated in the for-loop
 to add these count results to the 'Cell.Count' data frame made earlier, and 
@@
 
-Data from the image analysis has been written to the data frame "Cell.Count", where
-calling the data frame can view these results:
+Data from the image analysis will be written to the data frame "Cell.Count", where
+results can be viewed by calling the data frame:
 
@@
 out of this range and dictates a warning within our data frame, which is shown 
-below. This alerts the end-user to perhaps re-image a particular sample, or
-to investigate the error further:
+below. This alerts the end-user to investigate the error further where re-imaging a sample
+may be necessary:
 
@@
 - Total counts (generated previously)
-- Height of image field of view (μm)
-- Width of image field of view (μm)
+- Height of camera field of view (μm)
+- Width of camera field of view (μm)
 - Total number of images analyzed
@@
 
-Field of view (FOV - mm<sup>2</sup>) is an important variable calculated from image height and
+Camera field of view (FOV - mm<sup>2</sup>) is an important variable calculated from image height and
 image width. Image height and width are variables that are taken from dimensions 
@@
 
-Finally, use the following code to export the finalized data frame to a .csv file to the
+Finally, the following code can be used to export the finalized data frame to a .csv file to the
 pre-determined save directory:
@@
 ```
-<br>
 <br>
@@
 <br>
-<br>
 
@@
 
-Informal discussions with researchers have expressed interest in a easy-to-use analysis interface for inexperienced R users
-to utilize the ***cellcount*** tool.
-The following overview uses 'Shiny' apps to enter input data and run image analyses.
+Informal discussions with researchers have expressed interest in an easy-to-use analysis interface for inexperienced R users
+to utilize the ***cellcount*** package.
+The following overview uses 'Shiny' apps to enter input data and process images.
 This results in the use of two functions, *input_data_GUI* and *analyze images_GUI*, to run the entire analysis process as outlined previously, and in the **sandbox_analysis_test.R** script.
@@
 
-Users can use the following code to bring packages into the R environment once the packages have been installed using CRAN:
+The following code can be used to bring packages into the R environment once the packages have been installed using CRAN:
 
@@
 ### Inputting data
-The *input_data_GUI* function generates a Shiny UI where users enter information regarding directories, numerical and categorical data, 
+The *input_data_GUI* function generates a Shiny UI where users can enter information regarding directories, numerical and categorical data, 
 and selecting analysis type.
@@
 Other cells such as dinoflagellates (i.e.
-*Karenia* sp.) that do not contain uniform chloroplasts may be more difficult to enumerate as the ***cellcount*** approach may separate these lobed chloroplasts into separate sections, causing enumeration outputs to be unreliable.
-In-situ hybridization dyes that target specific species or genera may be helpful if other cellular features are not useful or rather hinder in providing accurate enumeration.
-
-Users must be aware that this package does not contain abilities to identify distinct cells.
+*Karenia* sp.) that contain multiple discriminate chloroplasts are difficult to enumerate as the ***cellcount*** approach may separate these lobed chloroplasts into unique cells, causing enumeration outputs to be unreliable.
+*In-situ* hybridization dyes that target specific species or genera may be helpful if other cellular features are not acceptable in providing accurate enumeration.
+
+Users must be aware that this package does not contain abilities to classify distinct cell types (i.e., species).
 The development of this package utilized monoculture of various cyanobacteria species.
 Users may find difficulty in using this analysis tool if cultures contain heavy amounts of particulates, other cellular debris, and contamination of other species.
-It is advised to use this tool for single-species only, however future developments of this tool are planned to include identifying cell features and deep-learning mechanics that use image libraries of specific cellular species.
-
-Outside of package limitations, users must be aware of limitations that may arise from poor microscopy imaging; reduce lighting as much as possible, avoid heavy background fluorescence, reduce debris/particulates, and so forth.
-Good images produce good ***cellcount*** results - keep this in mind as you utilize this tool.
+It is advised to use this tool for single-species only, however future developments of this tool are planned to include identifying cell features and deep-learning mechanics that use image libraries of unique cellular species.
+
+Outside of package limitations, users must be aware of limitations that may arise from poor microscope image quality; reduce room lighting for best results, avoid heavy background fluorescence, reduce debris/particulates, and so forth.
+Good images produce good ***cellcount*** results.
 
@@
 
-***cellcount*** provides an open-source approach to cyanobacteria enumeration that is accurate and rapid compared to standardized microscopy methods.
-The open-source approach allows researchers to access this tool for free and to modify and make additions the existing code to fulfill their particular needs, whereas the Shiny GUI allows users of all coding backgrounds to utilize this analysis tool.
+***cellcount*** provides an open-source approach to bacteria and cyanobacteria enumeration that is accurate and rapid compared to standardized microscopy methods.
+The open-source approach allows researchers to access this tool for free and to modify and make additions to the existing code to fulfill particular needs, whereas the Shiny GUI allows users of all coding backgrounds to utilize this analysis tool.
 

### 2023-08-30 sandbox_analysis_test.R
No major changes - updates/changes from test data

@@ -16,14 +16,14 @@ gc()
 
 Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
 
-file_name<-paste0("E.coli_") #Change the name of species analyzed here
+file_name<-paste0("Anabena_") #Change the name of species analyzed here
 
 #Change directories here
-savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/CSV_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/Convert_Images/E.coli_100x_DAPI/1C/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/quantitative_images/E.coli_100x_DAPI/1C/"
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_data")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_results/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_images/NZ_Anabena_60x/"
                      , pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/quantitative_images/E.coli_100x_DAPI/1C/"
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_images/NZ_Anabena_60x/"
                            , pattern = "tif", full.name = F)
 
 imgNames <- paste0(file_name, images_names)
@@ -38,9 +38,9 @@ img1<-lapply(grey_images,mapped_avg)
 
 for (j in 1:length(images)) {
   adj1<-median(img1[[j]])
-  adj2<-adj1+0.2
+  adj2<-adj1+0.15
   imagesMapped <- lapply(grey_images, mapped, threshold = adj2) #background intensity threshold adjustment
-  imagesConverted <- single_cell_convert(imagesMapped[[j]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
[…]

### 2023-09-06 cellcount_intro.Rmd
Updates to vingette from comments

@@
 
-The ***cellcount*** package provides a simple approach to quantifying bacteria and cyanobacteria *via* flourescence microscopy.
+The ***cellcount*** package provides a simple approach to quantifying bacteria and cyanobacteria *via* fluorescence microscopy.
 Microbial enumeration is an important tool for validating new methods and critical to assess potential ecological changes.
@@
 Additionally, other methods of counting cells or particles (i.e.
-coulter counter, flow cytometer, etc.) are not suitable for filamentous or colonial morphotypes that can clog flow cells and apertures.
+Coulter counter, flow cytometer, etc.) are not suitable for filamentous or colonial morphotypes that can clog flow cells and apertures.
 Here, we outline a new open-source package, ***cellcount***, a continuation of the report outlined in [@Pokrzywinski2019] and designed in the R computing language, to quickly and accurately count various microbial morphotypes.
@@
 To get started with ***cellcount***, make a local clone of the repository using programs like [GitHub Desktop](https://desktop.github.com/).
-Once cloned, open the R project file and install the package locally by selecting 'build' and 'install package', or use the shortcut Crtl+Shift+B (Cmd+Shift+B for MacOS users).
+Once cloned, open the R project file and install the package locally by selecting 'build' and 'install package', or use the shortcut Ctrl+Shift+B (Cmd+Shift+B for MacOS users).
 Once the ***cellcount*** package is installed, the following code can be used to bring ***cellcount*** into the R environment:
@@
 ```{r}
-Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
+Demo.Results <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
 ```
@@
 Once necessary adjustments have been made, the following for-loop is used to
-interatively process batch images of interest. In this example, using the provided demo *Microcystis* sp.
+interactively process batch images of interest. In this example, using the provided demo *Microcystis* sp.
 images, the for-loop uses the *single_cell_convert* function to produce unique 
@@
 to count the number of unique color labels (i.e., cells or particles). Code has also been integrated in the for-loop
-to add these count results to the 'Cell.Count' data frame made earlier, and 
+to add these count results to the 'Demo.Results' data frame made earlier, and 
 to save each unique color label image to the "image_savdir" directory:
@@
   count <- count_cells(imagesConverted)
-  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
+  Demo.Results[nrow(Demo.Results) + 1, ] <- c(imgNames[[z]], count)
   analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
@@
 
-Data from the image analysis will be written to the data frame "Cell.Count", where
+Data from the image analysis will be written to the data frame 'Demo.Results, where
 results can be viewed by calling the data frame:
 
-```{r}
-Cell.Count
+```{r,include=FALSE}
+Demo.Results
 ```
@@
 outlier_detect<-function(){
-  mean<-median(Cell.Count$Cell_Count)
-  sd<-sd(Cell.Count$Cell_Count)*2
+  mean<-median(Demo.Results$Cell_Count)
+  sd<-sd(Demo.Results$Cell_Count)*2
   sd_neg<-(mean-sd)
   sd_pos<-(mean+sd)
-  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
-    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
+  SD_range_detection<-ifelse(sd_pos>Demo.Results$Cell_Count,Demo.Results$Cell_Count,NA)&
+    ifelse(sd_neg<Demo.Results$Cell_Count,Demo.Results$Cell_Count,NA)
   return(SD_range_detection)
@@
 ```{r}
-Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
-Cell.Count$SD_range<-outlier_detect()
-Cell.Count
+Demo.Results$Cell_Count<-as.numeric(Demo.Results$Cell_Count)
+Demo.Results$SD_range<-outlier_detect()
+Demo.Results
 ```
@@
 In order to calculate cell density of these images, the
-*cell_density* function can be implemented, using equations from [@Wetzel2000]. Generating the total amount of 
-cells counted from these images is needed, where it can be added to the 'Cell.Count' data frame using the following code:
+*cell_density* function can be implemented, using equations from [@Wetzel2000]. The output results need to then be converted
+to a cell density variable, where it can be added to the 'Demo.Results' data frame using the following code:
 
 ```{r}
-cell.total <- as.numeric(Cell.Count$Cell_Count)
+cell.total <- as.numeric(Demo.Results$Cell_Count)
 cell.total <- sum(cell.total)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total,"null")
-```
-
-*Important note* - *the "null" data produced in the code above (and others that*
-*follow) fills in the blanks from the previous data produced by the* 
-***outlier_detect*** *function.*
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Cell Total", cell.total," ")
+```
 
@@
 
-- Total counts (generated previously)
+- Total counts (generated previously using this package)
 - Height of camera field of view (μm)
@@
 With the variables calculated, the following code uses these variables to 
-calculate cell density and adds the data into the 'Cell.Count' data frame:
+calculate cell density and adds the data into the 'Demo.Results' data frame:
 
@@
 cell.den <- cell_density(cell.total, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
+cell.den <- round(cell.den,2)
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Cell Density", cell.den," ")
 ```
@@
 per mL using the following code. Note that 'cell.av' calculation must contain the
-number of images the end-user has analyzed, and the 'cell.mL' must be divided
-by the total volume from the original sample:
+number of images the end-user has analyzed (excluding poor quality images and outliers), 
+and the 'cell.mL' must be divided by the total volume from the original sample where the
+filtered subsample was aliquotted from:
 
@@
 cell.av<-(cell.total/10) #change this via total number of images
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Imaging Average",cell.av," ")
 
 cell.mL<-(cell.den/5) #change this via total volume
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")
-```
-
-Calling upon the 'Cell.Count' data frame can show the new generated data and to double check for any calculation errors:
-
-```{r}
-Cell.Count
+cell.mL <- round(cell.mL,2)
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Total Volume Cell per mL",cell.mL," ")
+```
+
+Calling upon the 'Demo.Results' data frame can show the new generated data and to double check for any calculation errors:
+
+```{r,include=FALSE}
+Demo.Results
 ```
@@
 ```{r}
-write.csv(Cell.Count, paste0(savdir, "/F192_Test counts.csv"))
-```
-
-# Output data from test images in the ***cellcount*** package:
+write.csv(Demo.Results, paste0(savdir, "/F192_Test counts.csv"))
+```
+
+# Output data from test images in 
+#the ***cellcount*** package:
 
@@
 display(final_img)
-Cell.Count
+Demo.Results
 ```
@@
 
-### Test data results for *Anabena* sp.
+### Test data results for *Anabaena* sp.
 ```{r,include=FALSE}
@@
 ```{r,echo=FALSE}
-Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
+Demo.Results <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
 
@@
 
-Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
-Cell.Count$SD_range<-outlier_detect()
-
-cell.total <- as.numeric(Cell.Count$Cell_Count)
+Demo.Results$Cell_Count<-as.numeric(Demo.Results$Cell_Count)
+Demo.Results$SD_range<-outlier_detect()
+
+cell.total <- as.numeric(Demo.Results$Cell_Count)
 cell.total <- sum(cell.total)
 
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total,"null")
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Cell Total", cell.total," ")
 
 cell.den <- cell_density(cell.total, FOV = 0.027770534, images = 10, filtration.area = 213.8, volume = 0.5, total.volume =15)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
+cell.den <- round(cell.den,2)
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Cell Density", cell.den," ")
 
 cell.av<-(cell.total/10)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Imaging Average",cell.av," ")
 
 cell.mL<-(cell.den/15)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")
-
-write.csv(Cell.Count, paste0(savdir, "/Anabena_test counts.csv"))
-
-Cell.Count
+cell.mL <- round(cell.mL,2)
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Total Volume Cell per mL",cell.mL," ")
+
+write.csv(Demo.Results, paste0(savdir, "/Anabena_test counts.csv"))
+
+Demo.Results
 ```
@@
 
-Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
-Cell.Count$SD_range<-outlier_detect()
-
-cell.total <- as.numeric(Cell.Count$Cell_Count)
+Demo.Results$Cell_Count<-as.numeric(Demo.Results$Cell_Count)
+Demo.Results$SD_range<-outlier_detect()
+
+cell.total <- as.numeric(Demo.Results$Cell_Count)
 cell.total <- sum(cell.total)
 
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total,"null")
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Cell Total", cell.total," ")
 
 cell.den <- cell_density(cell.total, FOV = 0.00042, images = 10, filtration.area = 213.8, volume = 0.02, total.volume =2)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
+cell.den <- round(cell.den,2)
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Cell Density", cell.den," ")
 
 cell.av<-(cell.total/10)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av,"null")
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Imaging Average",cell.av," ")
 
 cell.mL<-(cell.den/2)
-Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL,"null")
-
-write.csv(Cell.Count, paste0(savdir, "/Shewanella_test counts.csv"))
-
-Cell.Count
+cell.mL <- round(cell.mL,2)
+Demo.Results[nrow(Demo.Results) + 1, ] <- c("Total Volume Cell per mL",cell.mL," ")
+
+write.csv(Demo.Results, paste0(savdir, "/Shewanella_test counts.csv"))
+
+Demo.Results
 ```
@@
 
-# Package limitations:
+# Package limitations/constraints:
 
@@
 Users must be aware that this package does not contain abilities to classify distinct cell types (i.e., species).
-The development of this package utilized monoculture of various cyanobacteria species.
+The development of this package utilized mono-culture of various cyanobacteria species.
 Users may find difficulty in using this analysis tool if cultures contain heavy amounts of particulates, other cellular debris, and contamination of other species.

### 2024-03-14 R/analyze_images_GUI.R
Updated - change file location of 'analyze_img.R'

@@ -39,7 +39,7 @@ analyze_images_GUI<-function(){
     })
     observeEvent(input$buttonId, {
       message("running analyze_img.R")
-      source("analyze_img.R")
+      source("cellcount_files_HABF/analyze_img.R")
     })
     onclick("BRefresh",{
       proxy=dataTableProxy("data")

### 2024-03-14 cellcount_files_HABF/analyze_img.R
Create new folder containing project files, functions, etc.

@@ -1,0 +1,67 @@
+library(EBImage)
+library(tiff)
+library(pixmap)
+library(raster)
+library(beepr)
+
+mapped_avg <- function(x) {
+  x <- as.matrix(x)
+  return(x)
+}
+
+
+if(analysis_type == 'single-cell'){
+  read_images <- lapply(images, readTIFF)
+  img_transposed <- lapply(read_images,aperm,c(2,1,3))
+  grey_images <- lapply(img_transposed, greyscale, contrast = contrast_adj2)
+  img1<-lapply(grey_images,mapped_avg)
+  for (z in 1:length(images)) {
+    adj1<-median(img1[[z]])
+    adj2<-adj1+thresh_adj2
+    imagesMapped <- lapply(grey_images, mapped, threshold = adj2) #background intensity threshold adjustment
+    imagesConverted <- single_cell_convert(imagesMapped[[z]], w = rec_width2, h = rec_height2, offset = 0.001, areathresh = areathresh2, tolerance = tolerance2, ext = ext2)
+    final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
+    count <- count_cells(imagesConverted)
+    Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
+    analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
+    writeImage(final_img, files = paste0(image_file, analyzed_image),compression=c("LZW"))
+  }
+  cell.total <<- as.numeric(Cell.Count$Cell_Count)
+  cell.total <<- sum(cell.total)
+  Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total)
+  cell.den <<- cell_density(cell.total, FOV = FOV2, images = image_num2, filtration.area = filter_area2, volume = filter_vol2, total.volume = total_vol2)
+  Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den)
+  cell.av<<-(cell.total/image_num2)
+  Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av)
+  cell.mL<<-(cell.den/total_vol2)
+  Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL)
+  write.csv(Cell.Count, paste0(csv_file, csv_file2))
[…]

### 2024-03-14 cellcount_files_HABF/cellcount_sandbox.R

@@ -1,0 +1,82 @@
+library(EBImage)
+library(tiff)
+library(pixmap)
+library(raster)
+library(beepr)
+library(outliers)
+library(cellcount)
+
+mapped_avg <- function(x) {
+  x <- as.matrix(x)
+  return(x)
+}
+
+# clear environment and free unused memory
+gc()
+
+Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
+
+file_name<-paste0("E.coli_") #Change the name of species analyzed here
+
+#Change directories here
+savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/cellcount_data/CSV_data/")
+image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/Convert_Images/E.coli_100x_qPCR/Rep 3/Con_E/")
+images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/quantitative_images/E.coli_100x_qPCR/Rep 3/Con_E/"
+                     , pattern = "tif", full.name = T)
+images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/quantitative_images/E.coli_100x_qPCR/Rep 3/Con_E/"
+                           , pattern = "tif", full.name = F)
+
+imgNames <- paste0(file_name, images_names)
+read_images <- lapply(images, readTIFF)
+img_transposed <- lapply(read_images,aperm,c(2,1,3))
+names(images) <- imgNames
+
+grey_images <- lapply(img_transposed, greyscale, contrast = 4)
+display(grey_images[[2]]) #visualize contrast/brightness adjustment here
+
+img1<-lapply(grey_images,mapped_avg)
+
+for (j in 1:length(images)) {
+  adj1<-median(img1[[j]])
+  adj2<-adj1+0.15
+  imagesMapped <- lapply(grey_images, mapped, threshold = adj2) #background intensity threshold adjustment
+  imagesConverted <- filamentous_convert(imagesMapped[[j]], w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance = 4, ext = 2)
+  final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
[…]

### 2024-03-14 sandbox_analysis_test.R
File renamed

@@ -1,80 +1,0 @@
-library(EBImage)
-library(tiff)
-library(pixmap)
-library(raster)
-library(beepr)
-library(outliers)
-library(cellcount)
-
-mapped_avg <- function(x) {
-  x <- as.matrix(x)
-  return(x)
-}
-
-# clear environment and free unused memory
-gc()
-
-Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
-
-file_name<-paste0("Anabena_") #Change the name of species analyzed here
-
-#Change directories here
-savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_data")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_results/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_images/NZ_Anabena_60x/"
-                     , pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/inst/extdata/test_images/NZ_Anabena_60x/"
-                           , pattern = "tif", full.name = F)
-
-imgNames <- paste0(file_name, images_names)
-read_images <- lapply(images, readTIFF)
-img_transposed <- lapply(read_images,aperm,c(2,1,3))
-names(images) <- imgNames
-
-grey_images <- lapply(img_transposed, greyscale, contrast = 4)
-display(grey_images[[2]]) #visualize contrast/brightness adjustment here
-
-img1<-lapply(grey_images,mapped_avg)
-
-for (j in 1:length(images)) {
-  adj1<-median(img1[[j]])
-  adj2<-adj1+0.15
-  imagesMapped <- lapply(grey_images, mapped, threshold = adj2) #background intensity threshold adjustment
-  imagesConverted <- filamentous_convert(imagesMapped[[j]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
-  final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
-  count <- count_cells(imagesConverted)
[…]

### 2024-03-14 analyze_img.R
File moved

@@ -1,67 +1,0 @@
-library(EBImage)
-library(tiff)
-library(pixmap)
-library(raster)
-library(beepr)
-
-mapped_avg <- function(x) {
-  x <- as.matrix(x)
-  return(x)
-}
-
-
-if(analysis_type == 'single-cell'){
-  read_images <- lapply(images, readTIFF)
-  img_transposed <- lapply(read_images,aperm,c(2,1,3))
-  grey_images <- lapply(img_transposed, greyscale, contrast = contrast_adj2)
-  img1<-lapply(grey_images,mapped_avg)
-  for (z in 1:length(images)) {
-    adj1<-median(img1[[z]])
-    adj2<-adj1+thresh_adj2
-    imagesMapped <- lapply(grey_images, mapped, threshold = adj2) #background intensity threshold adjustment
-    imagesConverted <- single_cell_convert(imagesMapped[[z]], w = rec_width2, h = rec_height2, offset = 0.001, areathresh = areathresh2, tolerance = tolerance2, ext = ext2)
-    final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
-    count <- count_cells(imagesConverted)
-    Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
-    analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
-    writeImage(final_img, files = paste0(image_file, analyzed_image),compression=c("LZW"))
-  }
-  cell.total <<- as.numeric(Cell.Count$Cell_Count)
-  cell.total <<- sum(cell.total)
-  Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total)
-  cell.den <<- cell_density(cell.total, FOV = FOV2, images = image_num2, filtration.area = filter_area2, volume = filter_vol2, total.volume = total_vol2)
-  Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den)
-  cell.av<<-(cell.total/image_num2)
-  Cell.Count[nrow(Cell.Count) + 1, ] <- c("Imaging Average",cell.av)
-  cell.mL<<-(cell.den/total_vol2)
-  Cell.Count[nrow(Cell.Count) + 1, ] <- c("Total Volume Cell per mL",cell.mL)
-  write.csv(Cell.Count, paste0(csv_file, csv_file2))
[…]

### 2024-03-14 demo_analysis_test.R

@@ -1,78 +1,0 @@
-library(EBImage)
-library(tiff)
-library(pixmap)
-library(raster)
-library(beepr)
-library(outliers)
-library(cellcount)
-
-mapped_avg <- function(x) {
-  x <- as.matrix(x)
-  return(x)
-}
-
-# clear environment and free unused memory
-gc()
-
-Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
-
-file_name<-paste0("F192_") #Change the name of species analyzed here
-
-#Change directories here
-savdir <- ("./inst/extdata/test_data/")
-image_savdir <- ("./inst/extdata/test_results/")
-images <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
-images_names <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
-
-imgNames <- paste0(file_name, images_names)
-read_images <- lapply(images, readTIFF)
-img_transposed <- lapply(read_images,aperm,c(2,1,3))
-names(images) <- imgNames
-
-grey_images <- lapply(img_transposed, greyscale, contrast = 4)
-display(grey_images[[2]]) #visualize contrast/brightness adjustment here
-
-img1<-lapply(grey_images,mapped_avg)
-
-for (j in 1:length(images)) {
-  adj1<-median(img1[[j]])
-  adj2<-adj1+0.2
-  imagesMapped <- lapply(grey_images, mapped, threshold = adj2) #background intensity threshold adjustment
-  imagesConverted <- single_cell_convert(imagesMapped[[j]], w = 17, h = 17, offset = 0.001, areathresh = 250, tolerance = 1, ext = 1)
-  final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
-  count <- count_cells(imagesConverted)
-  Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[j]], count)
-  analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[j]), "_analyzed.tiff")
-  writeImage(final_img, files = paste0(image_savdir, analyzed_image),compression=c("LZW"))
-}
-
-Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
[…]

### 2024-03-14 R/analyze_images_GUI.R
File moved to cellount_files_HABF folder

@@ -1,63 +1,0 @@
-#' Analyze Images GUI
-#'
-#' Brings in data entered from the "input_data_GUI" function and automatically runs a script to analyze images
-#' pertaining to cells from fluorescent microscopy images. This uses a Shiny user-interface, where three buttons
-#' are available to interact with, as well as an interactive table below. The "Run Image Analysis" button runs 
-#' the analysis script, the "Close Window" button closes the Shiny interface, and the "Refresh" button allows 
-#' users to view the analyzed data once the analysis has completed.
-#' 
-#' @return Runs an automatic image analysis
-#' @export
-
-analyze_images_GUI<-function(){
-  library(shiny)
-  library(shinyalert)
-  library(shinyFiles)
-  library(shinyjs)
-  library(lubridate)
-  library(DT)
-  
-  jscode <- "shinyjs.closeWindow = function() { window.close(); }"
-  
-  ui = fluidPage(
-    useShinyjs(),
-    h1("  "),
-    h1("  "),
-    actionButton("buttonId", "Run Image Analysis",class = "btn-success",style='height:75px;width:252px;font-size:140%',icon=icon("wand-magic-sparkles"),style="display:center-align"),
-    actionButton("close", "Close window",class = "btn-danger",style='height:75px;width:252px;font-size:140%',icon=icon("check"),style="display:center-align"),
-    actionButton("BRefresh","Refresh",class = "btn-success",style='height:75px;width:252px;font-size:140%',icon=icon("arrows-rotate"),style="display:center-align"),
-    h1("  "),
-    DT::dataTableOutput('data')
-  )
-  
-  
-  server = function(input, output, session) {
-    rv <- reactiveVal(Cell.Count)
-    
-    output$data <- DT::renderDataTable ({
-      DT::datatable(rv(), editable = TRUE)
-    })
-    observeEvent(input$buttonId, {
-      message("running analyze_img.R")
-      source("cellcount_files_HABF/analyze_img.R")
-    })
-    onclick("BRefresh",{
-      proxy=dataTableProxy("data")
-      replaceData(proxy,Cell.Count)
-    })
-    observeEvent(input$data_cell_edit, {
-      info <- input$data_cell_edit
-      newdf <- rv()
-      newdf[info$row, info$col] <- info$value
-      rv(newdf)
-      Cell.Count <<- rv()
-    })
-    observeEvent(input$close, {
-      js$closeWindow()
-      stopApp()
-    })
-  }
[…]

### 2024-03-14 R/border.R

@@ -1,13 +1,0 @@
-#' Border conversion
-#'
-#' Takes images to predict cells along a predetermined border in order
-#' to count cells along the predetermined border. This will be used to help
-#' determine those cells to "remove" them from the count function "countCells".
-#'
-#' @param x An image file
-#' @param y Dimming variable (using the 'dim' function from EBImage)
-#' @return Analyzing the cells on the border of \code{x}
-
-border <- function(x, y) {
-  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
-}

### 2024-03-14 R/cell_density.R

@@ -1,37 +1,0 @@
-#' Cell density
-#'
-#' Takes count results from processed images to calculate cell density based
-#' on camera field of view settings (inputted by end-user), number of images, filter area, volumes,
-#' and amount of cells filtered that the image is visualizing.
-#'
-#' @param x count results from processed image
-#' @param FOV field of view from camera specs in square millimeters
-#' @param images total number of fields
-#' @param filtration.area filtration area in square millimeters
-#' @param volume volume filtered per field in mL
-#' @param total.volume total sample volume in mL
-#'
-#' @return Calculates the density of cells within your original sample, based on the inputted data
-#' @export
-#'
-#' @examples Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
-#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
-#' @examples images_names <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = F)
-#' @examples grey_images <- greyscale(images,contrast=2)
-#' @examples images_mapped <- mapped(grey_images,threshold=0.3)
-#' @examples imagesConverted <- image_convert(imagesMapped[[1]],w=10,h=10,offset=0.001,areathresh=50,tolerance=0.8,ext=1)
-#' @examples count <- countCells(imagesConverted)
-#' @examples Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[1]], count)
-#' @examples Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
-#' @examples cell.total <- as.numeric(Cell.Count$Cell_Count)
-#' @examples cell.total <- sum(cell.total)
-#' @examples cell_density(cell.total,FOV = 0.00084,images = 10,filtration.area = 213.8,volume = 1,total.volume = 2)
-#' 
-
-cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
-  MCF <- filtration.area / FOV
-  N <- x / images
-  cyano <- (MCF * N) / volume
-  cyano_total <- cyano * total.volume
-  return(cyano_total)
-}

### 2024-03-14 R/count_cells.R

@@ -1,25 +1,0 @@
-#' Image count
-#'
-#' Takes information from a converted image and counts cells within the image.
-#' This remove cells along the image border and only count cells within the image
-#' in order to produce a more accurate cell count.
-#'
-#' @param x converted image variable using the "image_convert" function
-#' 
-#' @return Counts cells within an processed image
-#' @export
-#'
-#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
-#' @examples grey_images <- greyscale(images,contrast=2)
-#' @examples images_mapped <- mapped(grey_images,threshold=0.3)
-#' @examples imagesConverted <- image_convert(imagesMapped[[1]],w=10,h=10,offset=0.001,areathresh=50,tolerance=0.8,ext=1)
-#' @examples count_cells(imagesConverted)
-
-count_cells <- function(x) {
-  countAll <- max(x)
-  dims <- dim(x)
-  border1 <- c(x[1:dims[1], 1], x[1:dims[1], dims[2]], x[1, 1:dims[2]], x[dims[1], 1:dims[2]])
-  ids <- unique(border1[which(border1 != 0)])
-  countInner <- countAll - length(ids)
-  return(countInner)
-}

### 2024-03-14 R/filamentous_convert.R

@@ -1,33 +1,0 @@
-#' Filamentous convert
-#'
-#' Takes a mapped image (i.e., data matrix) and counts individual cells within an image.
-#' This is to be used with images that contains a single
-#' fluorescent dimension of cells. Best used for filamentous/colonial-cell type images.
-#' images.
-#'
-#' @param x an image variable that has been mapped using the "mapped" function
-#' @param w width of the moving rectangular window (see EBImage)
-#' @param h heigth of the moving rectangular window (see EBImage)
-#' @param offset a thresholding offset
-#' @param areathresh a numerical value regarding the individual pixel areas of cells within each image - this sets a threshold for what items are removed from images
-#' @param tolerance The minimum height of the object in the units of image intensity between its highest point and the point where it contacts another object (see EBImage)
-#' @param ext Radius of the neighborhood in pixels for the detection of neighboring objects. Higher values smooth out small objects (see EBImage for more details)
-#'
-#' @return Produces the proper data format for the function "countImages" to produce color-label images.
-#' @export 
-#'
-#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
-#' @examples grey_images <- greyscale(images,contrast=2)
-#' @examples images_mapped <- mapped(grey_images,threshold=0.3)
-#' @examples filamentous_convert(imagesMapped[[1]],w=10,h=10,offset=0.001,areathresh=50,tolerance=0.8,ext=1)
-
-
-filamentous_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
-  image <- thresh(x, w = w, h = h, offset = offset)
-  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
-  image2 <- fillHull(image1)
-  nf <- computeFeatures.shape(image2)
-  nr <- which(nf[, "s.area"] < areathresh)
-  image3 <- rmObjects(image2, nr)
-  return(image3)
-}

### 2024-03-14 R/greyscale.R

@@ -1,22 +1,0 @@
-#' Grey-scale conversion
-#'
-#' Takes images to adjust the contrast and converts them to grey-scale/8-bit.
-#' This function is needed to use the "mapping" function to properly
-#' convert images to data-frames.
-#'
-#' @param x an image variable
-#' @param contrast a numerical value for contrast adjustment
-#' 
-#' @return An image variable gets converted from RGB scale to grey scale
-#' @export
-#'
-#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
-#' @examples greyscale(images,contrast=2)
-
-greyscale <- function(x, contrast = 2) {
-  x <- contrast * x
-  x <- x[, , 1] + x[, , 2] + x[, , 3]
-  x <- x / max(x)
-  x <- normalize(x, inputRange = c(0.1, 0.75))
-  return(x)
-}

### 2024-03-14 R/input_data_GUI.R

@@ -1,407 +1,0 @@
-#' Input Data GUI
-#'
-#' This function provides a Shiny user-interface to easily input data needed for fluorescent microscopy image analysis
-#' outlined in *cellcount*. This Shiny app contains several tabs for data input, directory selection, etc. Users must close
-#' the app by selecting the 'Completion' tab followed by 'Submit data entries' for all entered data to be exported into 
-#' the local R environment, where users can then utilize the 'analyze_images_GUI' function to run the *cellcount* image analysis.
-#' 
-#' @return Data input user-interface for *cellcount* image analysis
-#' @export
-
-input_data_GUI<-function(){
-  library(shiny)
-  library(shinyalert)
-  library(shinyFiles)
-  library(shinyjs)
-  library(shinyBS)
-  
-  jscode <- "shinyjs.closeWindow = function() { window.close(); }"
-  
-  image_names1<<-textConnection('image_names2','wr',local=FALSE)
-  thresh_adj1<<-textConnection('thresh_adj2','wr',local=FALSE)
-  FOV1<<-textConnection('FOV2','wr',local=FALSE)
-  image_num1<<-textConnection('image_num2','wr',local=FALSE)
-  filter_vol1<<-textConnection('filter_vol2','wr',local=FALSE)
-  total_vol1<<-textConnection('total_vol2','wr',local=FALSE)
-  contrast_adj1<<-textConnection('contrast_adj2','wr',local=FALSE)
-  filter_area1<<-textConnection('filter_area2','wr',local=FALSE)
-  csv_file1<<-textConnection('csv_file2','wr',local=FALSE)
-  tolerance1<<-textConnection('tolerance2','wr',local=FALSE)
-  ext1<<-textConnection('ext2','wr',local=FALSE)
-  areathresh1<<-textConnection('areathresh2','wr',local=FALSE)
-  rec_width1<<-textConnection('rec_width2','wr',local=FALSE)
-  rec_height1<<-textConnection('rec_height2','wr',local=FALSE)
-  
-  ui1 = fixedPage(
-    useShinyjs(),
-    extendShinyjs(text = jscode, functions = c("closeWindow")),
-    div(navlistPanel(
-      "Cellcount - Image Analysis UI",
-      " ",
-      " ",
-      tabPanel(h4("Directories"),
-               h3("   "),
-               fixedRow(
[…]

### 2024-03-14 R/mapped.R

@@ -1,22 +1,0 @@
-#' Matrix mapping conversion
-#'
-#' Takes a grey-scale/8-bit image to convert to a data matrix,
-#' where data under the numerical pixel intensity threshold is removed.
-#' This function is needed for the use of the "image_convert" function.
-#'
-#' @param x a grey scale image variable
-#' @param threshold a numerical value of pixel intensity
-#'
-#' @return Converts an grey scale image to a data matrix, in a pixel by
-#' @return pixel format
-#' @export
-#' 
-#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
-#' @examples grey_images <- greyscale(images,contrast=2)
-#' @examples mapped(grey_images,threshold=0.3)
-
-mapped <- function(x, threshold = 0.3) {
-  x <- as.matrix(x)
-  x[x < threshold] <- 0
-  return(x)
-}

### 2024-03-14 R/single_cell_convert.R

@@ -1,31 +1,0 @@
-#' Single-cell convert
-#'
-#' Takes a mapped image (i.e., data matrix) and counts individual cells within an image.
-#' This is to be used with images that contains a single
-#' fluorescent dimension of cells. Best used for single-cell type images. 
-#'
-#' @param x an image variable that has been mapped using the "mapped" function
-#' @param w width of the moving rectangular window (see EBImage)
-#' @param h heigth of the moving rectangular window (see EBImage)
-#' @param offset a thresholding offset
-#' @param areathresh a numerical value regarding the individual pixel areas of cells within each image - this sets a threshold for what items are removed from images
-#' @param tolerance The minimum height of the object in the units of image intensity between its highest point and the point where it contacts another object (see EBImage for more details)
-#' @param ext Radius of the neighborhood in pixels for the detection of neighboring objects. Higher values smooth out small objects (see EBImage for more details)
-#'
-#' @return Produces the proper data format for the function "countImages" to produce color-label images.
-#' @export 
-#'
-#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
-#' @examples grey_images <- greyscale(images,contrast=2)
-#' @examples images_mapped <- mapped(grey_images,threshold=0.3)
-#' @examples single_cell_convert(imagesMapped[[1]],w=10,h=10,offset=0.001,areathresh=50,tolerance=0.8,ext=1)
-
-single_cell_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
-  image <- thresh(x, w = w, h = h, offset = offset)
-  image1 <- fillHull(image)
-  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
-  nf <- computeFeatures.shape(image2)
-  nr <- which(nf[, "s.area"] < areathresh)
-  image3 <- rmObjects(image2, nr)
-  return(image3)
-}

### 2024-03-14 cellcount_files_HABF/R/analyze_images_GUI.R
Folders moved to cellcount_files_HABF

@@ -1,0 +1,63 @@
+#' Analyze Images GUI
+#'
+#' Brings in data entered from the "input_data_GUI" function and automatically runs a script to analyze images
+#' pertaining to cells from fluorescent microscopy images. This uses a Shiny user-interface, where three buttons
+#' are available to interact with, as well as an interactive table below. The "Run Image Analysis" button runs 
+#' the analysis script, the "Close Window" button closes the Shiny interface, and the "Refresh" button allows 
+#' users to view the analyzed data once the analysis has completed.
+#' 
+#' @return Runs an automatic image analysis
+#' @export
+
+analyze_images_GUI<-function(){
+  library(shiny)
+  library(shinyalert)
+  library(shinyFiles)
+  library(shinyjs)
+  library(lubridate)
+  library(DT)
+  
+  jscode <- "shinyjs.closeWindow = function() { window.close(); }"
+  
+  ui = fluidPage(
+    useShinyjs(),
+    h1("  "),
+    h1("  "),
+    actionButton("buttonId", "Run Image Analysis",class = "btn-success",style='height:75px;width:252px;font-size:140%',icon=icon("wand-magic-sparkles"),style="display:center-align"),
+    actionButton("close", "Close window",class = "btn-danger",style='height:75px;width:252px;font-size:140%',icon=icon("check"),style="display:center-align"),
+    actionButton("BRefresh","Refresh",class = "btn-success",style='height:75px;width:252px;font-size:140%',icon=icon("arrows-rotate"),style="display:center-align"),
+    h1("  "),
+    DT::dataTableOutput('data')
+  )
+  
+  
+  server = function(input, output, session) {
+    rv <- reactiveVal(Cell.Count)
+    
+    output$data <- DT::renderDataTable ({
+      DT::datatable(rv(), editable = TRUE)
+    })
+    observeEvent(input$buttonId, {
+      message("running analyze_img.R")
+      source("cellcount_files_HABF/analyze_img.R")
+    })
+    onclick("BRefresh",{
+      proxy=dataTableProxy("data")
+      replaceData(proxy,Cell.Count)
+    })
+    observeEvent(input$data_cell_edit, {
+      info <- input$data_cell_edit
+      newdf <- rv()
+      newdf[info$row, info$col] <- info$value
+      rv(newdf)
+      Cell.Count <<- rv()
+    })
+    observeEvent(input$close, {
+      js$closeWindow()
+      stopApp()
+    })
+  }
[…]

### 2024-03-14 cellcount_files_HABF/R/border.R

@@ -1,0 +1,13 @@
+#' Border conversion
+#'
+#' Takes images to predict cells along a predetermined border in order
+#' to count cells along the predetermined border. This will be used to help
+#' determine those cells to "remove" them from the count function "countCells".
+#'
+#' @param x An image file
+#' @param y Dimming variable (using the 'dim' function from EBImage)
+#' @return Analyzing the cells on the border of \code{x}
+
+border <- function(x, y) {
+  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
+}

### 2024-03-14 cellcount_files_HABF/R/cell_density.R

@@ -1,0 +1,37 @@
+#' Cell density
+#'
+#' Takes count results from processed images to calculate cell density based
+#' on camera field of view settings (inputted by end-user), number of images, filter area, volumes,
+#' and amount of cells filtered that the image is visualizing.
+#'
+#' @param x count results from processed image
+#' @param FOV field of view from camera specs in square millimeters
+#' @param images total number of fields
+#' @param filtration.area filtration area in square millimeters
+#' @param volume volume filtered per field in mL
+#' @param total.volume total sample volume in mL
+#'
+#' @return Calculates the density of cells within your original sample, based on the inputted data
+#' @export
+#'
+#' @examples Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
+#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
+#' @examples images_names <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = F)
+#' @examples grey_images <- greyscale(images,contrast=2)
+#' @examples images_mapped <- mapped(grey_images,threshold=0.3)
+#' @examples imagesConverted <- image_convert(imagesMapped[[1]],w=10,h=10,offset=0.001,areathresh=50,tolerance=0.8,ext=1)
+#' @examples count <- countCells(imagesConverted)
+#' @examples Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[1]], count)
+#' @examples Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
+#' @examples cell.total <- as.numeric(Cell.Count$Cell_Count)
+#' @examples cell.total <- sum(cell.total)
+#' @examples cell_density(cell.total,FOV = 0.00084,images = 10,filtration.area = 213.8,volume = 1,total.volume = 2)
+#' 
+
+cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
+  MCF <- filtration.area / FOV
+  N <- x / images
+  cyano <- (MCF * N) / volume
+  cyano_total <- cyano * total.volume
+  return(cyano_total)
+}

### 2024-03-14 cellcount_files_HABF/R/count_images.R

@@ -1,0 +1,29 @@
+#' Count Images
+#'
+#' Displays converted microscopy images of distinct cell counts subsequently from the "image_convert" function.
+#'
+#' @param image A processed image variable from the "image_convert" function
+#' @param normalize if TRUE, normalizes the resulting color image
+#' @param removeEdgeCells automatically removes cells from along the image
+#' border using a true (T) or false (F) input
+#' 
+#' @return Produces a unique color-label image. Users can use the "display" function to view the image (see EBImage for more details) as well as save the image using the "writeImage" function (see readTIFF for more details).
+#' @export
+#'
+#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
+#' @examples grey_images <- greyscale(images,contrast=2)
+#' @examples images_mapped <- mapped(grey_images,threshold=0.3)
+#' @examples imagesConverted <- image_convert(imagesMapped[[1]],w=10,h=10,offset=0.001,areathresh=50,tolerance=0.8,ext=1)
+#' @examples count_images(imagesConverted, normalize = T, removeEdgeCells = T)
+
+count_images <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
+  if (removeEdgeCells) {
+    dims <- dim(image)
+    border1 <- c(image[1:dims[1], 1], image[1:dims[1], dims[2]], image[1, 1:dims[2]], image[dims[1], 1:dims[2]])
+    ids <- unique(border1[which(border1 != 0)])
+    inner <- rmObjects(image, ids)
+    EBImage::colorLabels(inner, normalize = normalize)
+  } else {
+    EBImage::colorLabels(image, normalize = normalize)
+  }
+}

### 2024-03-14 cellcount_files_HABF/R/filamentous_convert.R

@@ -1,0 +1,33 @@
+#' Filamentous convert
+#'
+#' Takes a mapped image (i.e., data matrix) and counts individual cells within an image.
+#' This is to be used with images that contains a single
+#' fluorescent dimension of cells. Best used for filamentous/colonial-cell type images.
+#' images.
+#'
+#' @param x an image variable that has been mapped using the "mapped" function
+#' @param w width of the moving rectangular window (see EBImage)
+#' @param h heigth of the moving rectangular window (see EBImage)
+#' @param offset a thresholding offset
+#' @param areathresh a numerical value regarding the individual pixel areas of cells within each image - this sets a threshold for what items are removed from images
+#' @param tolerance The minimum height of the object in the units of image intensity between its highest point and the point where it contacts another object (see EBImage)
+#' @param ext Radius of the neighborhood in pixels for the detection of neighboring objects. Higher values smooth out small objects (see EBImage for more details)
+#'
+#' @return Produces the proper data format for the function "countImages" to produce color-label images.
+#' @export 
+#'
+#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
+#' @examples grey_images <- greyscale(images,contrast=2)
+#' @examples images_mapped <- mapped(grey_images,threshold=0.3)
+#' @examples filamentous_convert(imagesMapped[[1]],w=10,h=10,offset=0.001,areathresh=50,tolerance=0.8,ext=1)
+
+
+filamentous_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
+  image <- thresh(x, w = w, h = h, offset = offset)
+  image1 <- watershed(distmap(image), tolerance = tolerance, ext = ext)
+  image2 <- fillHull(image1)
+  nf <- computeFeatures.shape(image2)
+  nr <- which(nf[, "s.area"] < areathresh)
+  image3 <- rmObjects(image2, nr)
+  return(image3)
+}

### 2024-03-14 cellcount_files_HABF/R/greyscale.R

@@ -1,0 +1,22 @@
+#' Grey-scale conversion
+#'
+#' Takes images to adjust the contrast and converts them to grey-scale/8-bit.
+#' This function is needed to use the "mapping" function to properly
+#' convert images to data-frames.
+#'
+#' @param x an image variable
+#' @param contrast a numerical value for contrast adjustment
+#' 
+#' @return An image variable gets converted from RGB scale to grey scale
+#' @export
+#'
+#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
+#' @examples greyscale(images,contrast=2)
+
+greyscale <- function(x, contrast = 2) {
+  x <- contrast * x
+  x <- x[, , 1] + x[, , 2] + x[, , 3]
+  x <- x / max(x)
+  x <- normalize(x, inputRange = c(0.1, 0.75))
+  return(x)
+}

### 2024-03-14 cellcount_files_HABF/R/input_data_GUI.R

@@ -1,0 +1,407 @@
+#' Input Data GUI
+#'
+#' This function provides a Shiny user-interface to easily input data needed for fluorescent microscopy image analysis
+#' outlined in *cellcount*. This Shiny app contains several tabs for data input, directory selection, etc. Users must close
+#' the app by selecting the 'Completion' tab followed by 'Submit data entries' for all entered data to be exported into 
+#' the local R environment, where users can then utilize the 'analyze_images_GUI' function to run the *cellcount* image analysis.
+#' 
+#' @return Data input user-interface for *cellcount* image analysis
+#' @export
+
+input_data_GUI<-function(){
+  library(shiny)
+  library(shinyalert)
+  library(shinyFiles)
+  library(shinyjs)
+  library(shinyBS)
+  
+  jscode <- "shinyjs.closeWindow = function() { window.close(); }"
+  
+  image_names1<<-textConnection('image_names2','wr',local=FALSE)
+  thresh_adj1<<-textConnection('thresh_adj2','wr',local=FALSE)
+  FOV1<<-textConnection('FOV2','wr',local=FALSE)
+  image_num1<<-textConnection('image_num2','wr',local=FALSE)
+  filter_vol1<<-textConnection('filter_vol2','wr',local=FALSE)
+  total_vol1<<-textConnection('total_vol2','wr',local=FALSE)
+  contrast_adj1<<-textConnection('contrast_adj2','wr',local=FALSE)
+  filter_area1<<-textConnection('filter_area2','wr',local=FALSE)
+  csv_file1<<-textConnection('csv_file2','wr',local=FALSE)
+  tolerance1<<-textConnection('tolerance2','wr',local=FALSE)
+  ext1<<-textConnection('ext2','wr',local=FALSE)
+  areathresh1<<-textConnection('areathresh2','wr',local=FALSE)
+  rec_width1<<-textConnection('rec_width2','wr',local=FALSE)
+  rec_height1<<-textConnection('rec_height2','wr',local=FALSE)
+  
+  ui1 = fixedPage(
+    useShinyjs(),
+    extendShinyjs(text = jscode, functions = c("closeWindow")),
+    div(navlistPanel(
+      "Cellcount - Image Analysis UI",
+      " ",
+      " ",
+      tabPanel(h4("Directories"),
+               h3("   "),
+               fixedRow(
[…]

### 2024-03-14 cellcount_files_HABF/R/mapped.R

@@ -1,0 +1,22 @@
+#' Matrix mapping conversion
+#'
+#' Takes a grey-scale/8-bit image to convert to a data matrix,
+#' where data under the numerical pixel intensity threshold is removed.
+#' This function is needed for the use of the "image_convert" function.
+#'
+#' @param x a grey scale image variable
+#' @param threshold a numerical value of pixel intensity
+#'
+#' @return Converts an grey scale image to a data matrix, in a pixel by
+#' @return pixel format
+#' @export
+#' 
+#' @examples images <- list.files("C:/Users/Desktop/Test_Images/",pattern = "tif", full.name = T)
+#' @examples grey_images <- greyscale(images,contrast=2)
+#' @examples mapped(grey_images,threshold=0.3)
+
+mapped <- function(x, threshold = 0.3) {
+  x <- as.matrix(x)
+  x[x < threshold] <- 0
+  return(x)
+}

### 2024-03-14 cellcount_intro.Rmd
Updated - small changes to text

@@
   - name: Tyler E. Harman 
-    affiliation: Consolidated Safety Services, Fairfax, VA 22031, USA
-      National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
+    affiliation: Consolidated Safety Services (Contracted to NOAA NCCOS), Fairfax, VA 22031, USA
     affiliation_email: tyler.harman@noaa.gov
@@
 
-  - name: David E. Berthold
-    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
-
-  - name: H. Dail Laughinghouse
-    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
-    orcid_id: 0000-0003-1018-6948
-
-  - name: Kaytee Pokrzywinski
+  - name: Kaytee Pokrzywinski, PhD
     affiliation: National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
@@
 ```{r,include=FALSE}
-savdir <- ("./inst/extdata/test_data/")
-image_savdir <- ("./inst/extdata/test_results/")
+savdir <- ("./cellcount_files_HABF/inst/extdata/test_data/")
+image_savdir <- ("./cellcount_files_HABF/inst/extdata/test_results/")
 ```
@@
 ```{r}
-images <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
-images_names <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+images <- list.files("./cellcount_files_HABF/inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("./cellcount_files_HABF/inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 ```
@@
 
-savdir <- ("./inst/extdata/test_data/")
-image_savdir <- ("./inst/extdata/test_results/")
-images <- list.files("./inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
-images_names <- list.files("./inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)
+savdir <- ("./cellcount_files_HABF/inst/extdata/test_data/")
+image_savdir <- ("./cellcount_files_HABF/inst/extdata/test_results/")
+images <- list.files("./cellcount_files_HABF/inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
+images_names <- list.files("./cellcount_files_HABF/inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)
 
@@
 
-savdir <- ("./inst/extdata/test_data/")
-image_savdir <- ("./inst/extdata/test_results/")
-images <- list.files("./inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
-images_names <- list.files("./inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)
+savdir <- ("./cellcount_files_HABF/inst/extdata/test_data/")
+image_savdir <- ("./cellcount_files_HABF/inst/extdata/test_results/")
+images <- list.files("./cellcount_files_HABF/inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
+images_names <- list.files("./cellcount_files_HABF/inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)
 
@@
 The following overview uses 'Shiny' apps to enter input data and process images.
-This results in the use of two functions, *input_data_GUI* and *analyze images_GUI*, to run the entire analysis process as outlined previously, and in the **sandbox_analysis_test.R** script.
-Several packages are required to utilize these UI tools:
+This results in the use of two functions, *input_data_GUI* and *analyze images_GUI*, to run the entire analysis process as outlined previously, and in the **cellcount_sandbox.R** script under the
+<span style="text-decoration:underline">*cellcount_files_HABF*</span> folder. Several packages are required to utilize these UI tools:
 

### 2024-03-18 cellcount_files_HABF/cellcount_sandbox.R
Minor updates - changing file paths within code to remove local information

@@ -19,12 +19,13 @@ Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0)
 file_name<-paste0("E.coli_") #Change the name of species analyzed here
 
 #Change directories here
-savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/cellcount_data/CSV_data/")
-image_savdir <- ("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/Convert_Images/E.coli_100x_qPCR/Rep 3/Con_E/")
-images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/quantitative_images/E.coli_100x_qPCR/Rep 3/Con_E/"
-                     , pattern = "tif", full.name = T)
-images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_data/quantitative_images/E.coli_100x_qPCR/Rep 3/Con_E/"
-                           , pattern = "tif", full.name = F)
+savdir <- ("./CSV_data/") #Change save directory for .csv file data here
+image_savdir <- ("./Convert_Images/E.coli_100x_qPCR/Rep 3/Con_E/") #Change the save directory for analyzed images
+
+images <- list.files("./quantitative_images/E.coli_100x_qPCR/Rep 3/Con_E/"
+                     , pattern = "tif", full.name = T) #List the location of the raw images here
+images_names <- list.files("./quantitative_images/E.coli_100x_qPCR/Rep 3/Con_E/"
+                           , pattern = "tif", full.name = F) #List the location of the raw images here
 
 imgNames <- paste0(file_name, images_names)
 read_images <- lapply(images, readTIFF)

### 2024-03-18 cellcount_intro.Rmd
Minor updates - changed data.frame column header

@@
 Demo.Results$Cell_Count<-as.numeric(Demo.Results$Cell_Count)
-Demo.Results$SD_range<-outlier_detect()
+Demo.Results$within_outlier_range<-outlier_detect()
 Demo.Results
@@
 Demo.Results$Cell_Count<-as.numeric(Demo.Results$Cell_Count)
-Demo.Results$SD_range<-outlier_detect()
+Demo.Results$within_outlier_range<-outlier_detect()
 

### 2024-03-18 cellcount_files_HABF/cellcount_sandbox.R
Minor change - updated data.frame column header

@@ -61,7 +61,7 @@ outlier<-function(){
   return(SD_range_detection)
 }
 
-Cell.Count$SD_range<-outlier()
+Cell.Count$within_outlier_range<-outlier()
 
 cell.total <- as.numeric(Cell.Count$Cell_Count)
 cell.total <- sum(cell.total)

### 2024-04-10 pkgdown/logo.R
Initial creation of hexsticker logo for pkgdown site

@@ -1,0 +1,3 @@
+library(ggplot2)
+library(hexSticker)
+

### 2024-04-10 pkgdown/logo.R
Initial uploads of hexsticker file

@@ -1,3 +1,32 @@
 library(ggplot2)
 library(hexSticker)
+library(dplyr)
+library(cowplot)
 
+noaa <- magick::image_read("https://www.noaa.gov/sites/default/files/2022-03/noaa_digital_logo-2022.png") %>%
+  magick::image_colorize(65,"white")
+
+test_image1 <- ggdraw() +
+  draw_image(noaa)
+
+microscope <- magick::image_read("https://static.vecteezy.com/system/resources/previews/007/165/321/original/microscope-icon-microscope-icon-simple-sign-microscope-icon-free-free-vector.jpg") %>%
+  magick::image_resize("400x400")
+
+test_image2 <- ggdraw() +
+  draw_image(microscope)
+
+hexSticker::sticker(
+  filename = "inst/figures/logo.png",
+  subplot = test_image2,
+  s_width = 1.25, s_height = 1.5,
+  s_x = 1, s_y = .8,
+  package = "cellcount",
+  p_size = 24,
+  p_color = "orange",
+  h_size = 1,
+  h_fill = "white",
+  h_color = "black",
+  url = "HABF-SDI-NCCOS-NOAA/HABF_cellcount",
+  u_size = 4,
+  u_color = "black"
+) %>% plot()

### 2024-04-11 cellcount_intro.Rmd
Moved vingette file back to main folder
Minor updates to vignette files

@@
 ```{r,include=FALSE}
-savdir <- ("./cellcount_files_HABF/inst/extdata/test_data/")
-image_savdir <- ("./cellcount_files_HABF/inst/extdata/test_results/")
+savdir <- ("./inst/extdata/test_data/")
+image_savdir <- ("./inst/extdata/test_results/")
 ```
@@
 ```{r}
-images <- list.files("./cellcount_files_HABF/inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
-images_names <- list.files("./cellcount_files_HABF/inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
+images <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
+images_names <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 ```
@@
 
-savdir <- ("./cellcount_files_HABF/inst/extdata/test_data/")
-image_savdir <- ("./cellcount_files_HABF/inst/extdata/test_results/")
-images <- list.files("./cellcount_files_HABF/inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
-images_names <- list.files("./cellcount_files_HABF/inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)
+savdir <- ("./inst/extdata/test_data/")
+image_savdir <- ("./inst/extdata/test_results/")
+images <- list.files("./inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = T)
+images_names <- list.files("./inst/extdata/test_images/NZ_Anabena_60x/", pattern = "tif", full.name = F)
 
@@
 
-savdir <- ("./cellcount_files_HABF/inst/extdata/test_data/")
-image_savdir <- ("./cellcount_files_HABF/inst/extdata/test_results/")
-images <- list.files("./cellcount_files_HABF/inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
-images_names <- list.files("./cellcount_files_HABF/inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)
+savdir <- ("./inst/extdata/test_data/")
+image_savdir <- ("./inst/extdata/test_results/")
+images <- list.files("./inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = T)
+images_names <- list.files("./inst/extdata/test_images/Shewanella_100x_DAPI/", pattern = "tif", full.name = F)
 
@@
 
-![Visual demonstration of the Shiny GUI for inputting data.](C:/Users/Tyler.Harman/Pictures/cellcount-UI_1.png)
+![Visual demonstration of the Shiny GUI for inputting data.](./inst/figures/cellcount-UI_1.png)
 
@@
 
-![Visual demonstration of the Shiny GUI for running image analysis.](C:/Users/Tyler.Harman/Pictures/cellcount-UI_2.png)
+![Visual demonstration of the Shiny GUI for running image analysis.](./inst/figures/cellcount-UI_2.png)
 

### 2024-04-11 pkgdown/logo.R
Updated pkgdown site hexsticker

@@ -9,7 +9,9 @@ noaa <- magick::image_read("https://www.noaa.gov/sites/default/files/2022-03/noa
 test_image1 <- ggdraw() +
   draw_image(noaa)
 
-microscope <- magick::image_read("https://static.vecteezy.com/system/resources/previews/007/165/321/original/microscope-icon-microscope-icon-simple-sign-microscope-icon-free-free-vector.jpg") %>%
+test_image1
+
+microscope <- magick::image_read("https://cdn.icon-icons.com/icons2/1465/PNG/512/708microscope_100683.png") %>%
   magick::image_resize("400x400")
 
 test_image2 <- ggdraw() +
@@ -18,15 +20,15 @@ test_image2 <- ggdraw() +
 hexSticker::sticker(
   filename = "inst/figures/logo.png",
   subplot = test_image2,
-  s_width = 1.25, s_height = 1.5,
-  s_x = 1, s_y = .8,
+  s_width = .95, s_height = 1.5,
+  s_x = 1, s_y = .75,
   package = "cellcount",
-  p_size = 24,
-  p_color = "orange",
+  p_size = 26,
+  p_color = "royalblue",
   h_size = 1,
-  h_fill = "white",
+  h_fill = "grey95",
   h_color = "black",
   url = "HABF-SDI-NCCOS-NOAA/HABF_cellcount",
   u_size = 4,
-  u_color = "black"
+  u_color = "royalblue"
 ) %>% plot()

### 2024-07-31 cellcount_files_HABF/demo_analysis_test.R
Updates made from VLab comments - improvement to overall function

@@ -6,21 +6,26 @@ library(beepr)
 library(outliers)
 library(cellcount)
 
-mapped_avg <- function(x) {
-  x <- as.matrix(x)
-  return(x)
-}
-
 # clear environment and free unused memory
 gc()
 
 Cell.Count <- data.frame(Image_File_Name = character(0), Cell_Count = numeric(0))
-
 file_name<-paste0("F192_") #Change the name of species analyzed here
 
+getwd() #Location of current working directory and where folders will be generated/files stored
+
 #Change directories here
-savdir <- ("./inst/extdata/test_data/")
-image_savdir <- ("./inst/extdata/test_results/")
+main_directory <- paste0("C:/Users/Tyler.Harman/Desktop") #Set working directory here to store outcoming data
+savdir <- paste0(main_directory,"/test_data/")
+if(!dir.exists(savdir)){
+  dir.create(savdir)
+}
+
+image_savdir <- paste0(main_directory,"/test_results/")
+if(!dir.exists(image_savdir)){
+  dir.create(image_savdir)
+}
+
 images <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
 images_names <- list.files("./inst/extdata/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
 
@@ -30,7 +35,12 @@ img_transposed <- lapply(read_images,aperm,c(2,1,3))
 names(images) <- imgNames
 
 grey_images <- lapply(img_transposed, greyscale, contrast = 4)
-display(grey_images[[2]]) #visualize contrast/brightness adjustment here
+EBImage::display(grey_images[[2]]) #visualize contrast/brightness adjustment here
+
+mapped_avg <- function(x) {
+  x <- as.matrix(x)
+  return(x)
+}
 
 img1<-lapply(grey_images,mapped_avg)
 
@@ -56,7 +66,7 @@ outlier<-function(){
     ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
   return(SD_range_detection)
 }
-Cell.Count$SD_range<-outlier()
+Cell.Count$within_outlier_range<-outlier()
 
 cell.total <- as.numeric(Cell.Count$Cell_Count)
 cell.total <- sum(cell.total)

### 2024-08-28 cellcount_files_HABF/demo_analysis_test.R
Minor changes from VLab commits - using pacman code to summarize loading target packages.

@@ -1,10 +1,6 @@
-library(EBImage)
-library(tiff)
-library(pixmap)
-library(raster)
-library(beepr)
-library(outliers)
-library(cellcount)
+library(pacman)
+
+pacman::p_load(EBImage,tiff,pixmap,raster,beepr,outliers,cellcount)
 
 # clear environment and free unused memory
 gc()

### 2025-02-19 cellcount_files_HABF/demo_analysis_test.R
Removing sensetive data

@@ -11,7 +11,7 @@ file_name<-paste0("F192_") #Change the name of species analyzed here
 getwd() #Location of current working directory and where folders will be generated/files stored
 
 #Change directories here
-main_directory <- paste0("C:/Users/Tyler.Harman/Desktop") #Set working directory here to store outcoming data
+main_directory <- paste0("./") #Set working directory here to store outcoming data
 savdir <- paste0(main_directory,"/test_data/")
 if(!dir.exists(savdir)){
   dir.create(savdir)

### 2025-02-19 cellcount_files_HABF/analyze_img.R
Improvements to code suggested by reviewers - including error documentation metrics

@@ -10,58 +10,98 @@ mapped_avg <- function(x) {
 }
 
 
-if(analysis_type == 'single-cell'){
+if(analysis_type == 'single-cell'){ # Check if the analysis type is 'single-cell'
+  # Attempt to read TIFF images into a list
   read_images <- lapply(images, readTIFF)
-  img_transposed <- lapply(read_images,aperm,c(2,1,3))
-  grey_images <- lapply(img_transposed, greyscale, contrast = contrast_adj2)
-  img1<-lapply(grey_images,mapped_avg)
-  for (z in 1:length(images)) {
-    adj1<-median(img1[[z]])
-    adj2<-adj1+thresh_adj2
-    imagesMapped <- lapply(grey_images, mapped, threshold = adj2) #background intensity threshold adjustment
-    imagesConverted <- single_cell_convert(imagesMapped[[z]], w = rec_width2, h = rec_height2, offset = 0.001, areathresh = areathresh2, tolerance = tolerance2, ext = ext2)
-    final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T)
-    count <- count_cells(imagesConverted)
-    Cell.Count[nrow(Cell.Count) + 1, ] <- c(imgNames[[z]], count)
-    analyzed_image <- paste0(sub(".tif", replacement = "", x = imgNames[z]), "_analyzed.tiff")
-    writeImage(final_img, files = paste0(image_file, analyzed_image),compression=c("LZW"))
+  
+  # Check if images were successfully read
+  if (length(read_images) > 0 && all(sapply(read_images, is.array))) {
+    # If condition is true, images were successfully read
+    
+    # Print success message to console
+    cat("Images successfully read into the environment.\n")
+    
+    # Print the number of images read
+    print(paste("Number of images read:", length(read_images)))
+  } else {
+    # If condition is false, there was an error reading the images
+    
+    # Print error message to console
+    cat("Error: Failed to read one or more images.\n")
+    
+    # Stop script execution and display error message
+    stop("Image reading failed. Please check your image files and path.")
   }
-  cell.total <<- as.numeric(Cell.Count$Cell_Count)
-  cell.total <<- sum(cell.total)
-  Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Total", cell.total)
[…]

### 2025-02-19 R/analyze_images_GUI.R
Improving code organization - suggestions made by reviewers

@@ -38,8 +38,105 @@ analyze_images_GUI<-function(){
       DT::datatable(rv(), editable = TRUE)
     })
     observeEvent(input$buttonId, {
-      message("running analyze_img.R")
-      source("cellcount_files_HABF/analyze_img.R")
+      #message("running analyze_img.R")
+      #source("cellcount_files_HABF/analyze_img.R")
+      
+      if(analysis_type == 'single-cell'){ # Check if the analysis type is 'single-cell'
+        # Attempt to read TIFF images into a list
+        read_images <- lapply(images, readTIFF)
+        
+        # Check if images were successfully read
+        if (length(read_images) > 0 && all(sapply(read_images, is.array))) {
+          # If condition is true, images were successfully read
+          
+          # Print success message to console
+          cat("Images successfully read into the environment.\n")
+          
+          # Print the number of images read
+          print(paste("Number of images read:", length(read_images)))
+        } else {
+          # If condition is false, there was an error reading the images
+          
+          # Print error message to console
+          cat("Error: Failed to read one or more images.\n")
+          
+          # Stop script execution and display error message
+          stop("Image reading failed. Please check your image files and path.")
+        }
+        img_transposed <- lapply(read_images,aperm,c(2,1,3)) # Transpose images
+        grey_images <- lapply(img_transposed, greyscale, contrast = contrast_adj2) # Convert to greyscale with contrast adjustment
+        img1<-lapply(grey_images,mapped_avg) # Apply mapped_avg function to grey images
+        for (z in 1:length(images)) { # Loop through each image
+          adj1<-median(img1[[z]]) # Calculate median of mapped average
+          adj2<-adj1+thresh_adj2 # Adjust threshold
+          imagesMapped <- lapply(grey_images, mapped, threshold = adj2) # Apply mapping with adjusted threshold
+          imagesConverted <- single_cell_convert(imagesMapped[[z]], w = rec_width2, h = rec_height2, offset = 0.001, areathresh = areathresh2, tolerance = tolerance2, ext = ext2) # Convert single cell images
+          final_img <- count_images(imagesConverted, normalize = T, removeEdgeCells = T) # Count and process images
+          count <- count_cells(imagesConverted) # Count cells in converted images
[…]

### 2025-02-19 R/input_data_GUI.R

@@ -42,70 +42,218 @@ input_data_GUI<-function(){
       tabPanel(h4("Directories"),
                h3("   "),
                fixedRow(
-                 column(2,offset=2,shinyDirButton('path1','CSV save directory','Please select a directory to save CSV files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
+                 column(
+                   2,
+                   offset=2,
+                   shinyDirButton(
+                     'path1',
+                     'CSV save directory',
+                     'Please select a directory to save CSV files',
+                     FALSE,
+                     class = "btn-success",
+                     style='height:35px;width:350px;font-size:120%',
+                     icon=icon("folder-open")))
                ),
                h4("  "),
                fixedRow(
-                 column(2,offset=2,shinyDirButton('path2','Image save directory','Please select a directory to save Image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
+                 column(
+                   2,
+                   offset=2,
+                   shinyDirButton(
+                     'path2',
+                     'Image save directory',
+                     'Please select a directory to save Image files',
+                     FALSE,
+                     class = "btn-success",
+                     style='height:35px;width:350px;font-size:120%',
+                     icon=icon("folder-open")))
                ),
                h4("  "),
                fixedRow(
-                 column(2,offset=2,shinyDirButton('path3','Image analysis directory','Please select a directory containing image files',FALSE,class = "btn-success",style='height:35px;width:350px;font-size:120%',icon=icon("folder-open")))
+                 column(
+                   2,
+                   offset=2,
+                   shinyDirButton(
+                     'path3',
+                     'Image analysis directory',
+                     'Please select a directory containing image files',
+                     FALSE,
+                     class = "btn-success",
+                     style='height:35px;width:350px;font-size:120%',
+                     icon=icon("folder-open")))
                )
       ),
       tabPanel(h4("Analysis Type"),
                h3("   "),
                fixedRow(
-                 column(8,offset=2,selectInput("selectData",
-                                               h5(strong("Please select one of the following options regarding imaging analysis type:")),
-                                               choices=list("Single-cell","Filamentous","Colonial"),selected="Single-cell"))
+                 column(
+                   8,
+                   offset=2,
+                   selectInput(
+                     "selectData",
[…]
